### 2015-01-24 src/tex/main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphicx}
\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{hyperref}
\usepackage{textcomp}

\title{\LARGE \bf
	Object tracking with the Universal Robots\texttrademark \ UR5
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Ayla Shamineh Nawaz and Dawid Horacy Golebiewski}% <-this % stops a space
%\thanks{*This work was not supported by any organization}% <-this % stops a space
%\thanks{$^{1}$H. Kwakernaak is with Faculty of Electrical Engineering, Mathematics %and Computer Science,
%        University of Twente, 7500 AE Enschede, The Netherlands
%        {\tt\small h.kwakernaak at papercept.net}}%
%\thanks{$^{2}$D. Golebiewski is with the Department of Electrical Engineering, Hamburg University of Technology,
%        Dayton, OH 45435, USA
%        {\tt\small d.golebiewski at tuhh.de}}%
%}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
This paper concerns a general tracking problem in which a camera is mounted on the end effector of an articulated robot (UR5). Using image processing techniques and basic kinematic ideas, a method for tracking an object on a plane is introduced and explained. The limitations and alternative approaches to this problem are outlined and treated in the discussion section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}
Object tracking is an important part of many robotic applications. Especially in medical applicaitions the advantage of a robot being able to identify a certain object from the rest of the body and being able to not loose track of that object is high. For example surgical procedures can be augmented by using vision capabilities and sensor capabilities of robotic systems that a human surgeon by himself cannot have. To better understand the different aspects of tracking an object with a robot a simple tracking problem was adressed. A red object was placed on the same table on which the used robot arm was mounted. With the help of a camera mounted on the end effector of the robot arm, a method was implemented with which the robot arm is able to track the movement of a red test object on the table. Tracking was used in the sense of centering the object in the image plane. It is important at this point to realize that there were no other restrictions on how the robot arm should in general move through it's workspace while tracking the object. There are therefore multiple possibilities for realizing such a tracking problem. This paper provides a method for tracking in which the z-axis of the end effector is at any time oriented in the negative-z direction of the world coordinate frame. The world coordinate frame is defined as the origin of the first joint of the robot, with the z-axis being perpendicular to the table. At the same time the distance between table and the end effector is to be held constant. Since this will correspond to a motion of the robot in which the end effector is parallel to the table at all times, it is called "parallel tracking". This means that the X,Y position of the end effector is solely definded by the position of the red object. Doing so, no time has to be spent on defining possible end effector positions. This can lead to smoother and faster tracking. In the following we will present the steps necessary for parallel object tracking. This includes the kinematic properties of the robot and a way to define them for tracking purposes. As well as image processing concerning object recognition. It will be clear that most of the discussed methods also apply if a different form of tracking is chosen.

% \begin{itemize}
% \item problem discription
% \item looked at spherical tracking
% \item looked at parallel tracking
% \item red object recogintion any red object, just a dot?
% \item red object position defined
% \item position of the camera to the red object defined. e.g. our choice of tracking

% \end{itemize}

\section{Procedure and methods}

\subsection{General Setup}
The Universal Robot's UR5 model is a general-purpose articulated robot built using 6 serially connected rotary joints. It has a TCP/IP network interface and the MTEC institute at TUHH provided us with a MATLAB interface to connect with the robot's internal server, send commands and receive all information about the current state of the UR5. Attached to UR5's end effector was a Logitech QuickCam Pro 9000, which we chose to use at the maximum resolution of 1600x1200 pixels. The connection was direct over USB 2.0. 
To have tracking capability as close to real-time as possible the entire processing chain should be as fast as possible - ideally, the position of the camera center would only lag behind the object location by the image processing time $t_{img}$, the delay to calculate the target pose $t_{pos}$, the delay of getting the current state of the robot $t_{sta}$ and sending the next position $t_{snd}$ to the UR5. The sum of these variables is the variable reaction time $t_{chain}$. Knowing that the camera has a maximum frame rate of only 5 fps at its maximum resolution, we can deduce that the update frequency of our tracking setup is $5 $ Hz, since the image processing is parallel to the kinematics of the robot and a new pose is computed (if necessary) each time a frame from he video stream is acquired. This means that for large changes in target position, that result in $t_{move} > t_{FPS} $ the movement exectued by the UR5 can be interrupted by an updated target pose, which is an important feature for our intended use. %The total time-to-target position would be $ t_{FPS}+t_{chain}+t_{move} $, if the notion of a continuously moving target is abandoned.
\subsection{Kinematics and Inverse Kinematics}
Kinematic as well as Inverse Kinematic principals are used to relate the pose of the end effector to the world coordinate system and vice versa. Provided with the joint parameters of the robot, the relationships are described through simple transformation matrices. The transformation matrix from end effector of the robot to the base is described by a matrix filled with the information of the Denavit Hartenberg parameters and is only dependent on the joint parameters. Since the base of the robot and the base of the world coordinate system coincide this transformation matrix is just the pose of the end effector for a given set of joint parameters. Inverse kinematics is useful when the target pose of the end effector is computed and the joint parameters that are sent to the robot have to be determined. For each pose there are more than one possible solution set of joint parameters. The relationship between camera and end effector, also represented by a transformation matrix $H_{cam2endeff}$ will be computed once and will not change during tracking. This relationship can be determined through hand to eye calibration as explained in the following.
\subsection{Camera calibration}
Camera calibration is a method to estimate the intrinsic  parameters of the camera such as focal point, image sensor format, and principal point. This information is collected in a matrix (intrinsicparam). It also includes the estimation of the extrinsic parameters such as the transformation matrix from world coordinates to camera coordinates. The hand-eye calibration relating the position of the camera to the position of the end effector (Hcam2endeff) is also needed. Hand eye calibration as well as the intrinsic camera parameters can be computed using the proposed fully automatic camera and hand to eye calibration with a MATLAB add-on package. This package requires a certain amount of pictures taken by the camera of a grid that can be also be downloaded from \url{https://www.vision.ee.ethz.ch/software/calibration_toolbox/calibration_toolbox.php}. The grid is placed on the table and the pictures are taken from different angles, providing a unique pose for each picture. Both the pictures and the poses are requirements for the toolbox to compute the camera and the hand to eye calibration simultaneously.

\subsection{Target localisation in the Image Plane}
%We assume a pinhole camera model, we know that every scene in the world results in an inverted image on the camera sensor, which defines the image plane.
The image is aquired directly from the camera so that the image data can be used directly in the MATLAB environment. The focus of the solution lies in detecting circular objects for tracking as the test object used was of circular shape. To extend the capability of the system, a more general method has been implemented: to allow for the tracking of multiple objects of non-circular shape, the center of mass of all red shapes in the image is computed. This step is illustrated in Fig.1. In order to be able to find the the red objects in the scene in the field of vision, the acquired image is processed in the following steps:
\begin{enumerate}
\item Subtract red channel from grayscale image of the acquired frame
\item Use median filter $medfilt2(scene,[3~3])$ to filter out noise
\item Convert the resulting grayscale image into a binary image using $im2bw$
\item Remove small connected areas (fewer than 100px) from the binary image using $bwareaopen(scene,100)$ (see Fig.1)
\item Label all the connected areas in the scene using $bwlabel(scene,8)$
\item Use the $regionprops$ function to do image blob analysis using Eccentricity and the Centroid of the labeled areas
\item Case distinction: 
{\setlength\itemindent{25pt} \begin{enumerate}\item{If circular shape is found, find its centroid's coordinates}\item{Else If more than one connected area present, compute coordinates of the \textbf{center of mass} of all objects adn return $x_{pixel},y_{pixel}$} \item{Return coordinates on the image plane as $x_{pixel},y_{pixel}$ in pixel units}\end{enumerate}} 
\end{enumerate}
\begin{figure}
\includegraphics[width=0.5\textwidth]{daniel_suksuk.jpg}
\caption*{\label{fig:center of mass}Last step in image processing. In this case, the center of mass of all white connected areas would be computed.}
\end{figure}

%#####################################################
\subsection{Center of object in World coordinates}

The position of the red object in world coordinates is defined by the ray of the object intersecting with the principal point of the camera(see Fig.2). The direction of the negative part of the ray is given by the pixel coordinates $x_{pixel},y_{pixel}$ that were identified as the localization of the objects' centroid in the image plane. The distance of the principal point to the camera center, shown also in Fig.2. This length is the value of the focal length and can easily be extracted from the intrinsic camera parameters obtained though camera calibration. The pose of the principal point in world coordinates can be determined by the current pose of the end effector - given by the current joint parameters - and the matrix of the hand-to-eye calibration (Hcam2endeff). It is defined by: $pose*(Hcam2endeff)^{⁻1}$. Taking the principal point as the origin and going in the defined direction the $x_{world}$ and $y_{world}$ coordinates of the object can be computed with the condition that the object will always be at $z=0$.  
\begin{figure}
\includegraphics[width=0.5\textwidth]{DaniLLL.jpg}
\caption*{\label{fig:finding target position}ray of object on image plane}
\end{figure}

%######################################################
\subsection{Constructing the target pose of the end effector}
The target pose of the end effector is defined by the position of the object in world coordinates $x_{world}$ and $y_{world}$. Considering that the total arm length (from base to end effector) of the robot arm  is about $0.8m$, the z-coordinate can be chosen arbitrarily as long as a collision with the table is avoided. It should however be considered that when choosing the z-coordinate of the target pose too high the area which the robot can track will become smaller. On the other hand, choosing the z-coordinate too low will make it harder for the robot to track fast moving objects, since the field of vision will reduce. A height of $z=0.3m$ offered a good compromise between the field of vision an d reachable workspace in parallel mode.\\The rotation of the end effector is constructed such that the z-coordinate of the end effector will always point in the negative z direction of the world coordinate system. The remaining components of the rotation matrix were taken from a pose that was set up manually. We chose this pose in a way that the camera was aligned with the longer side of the table. Of course any other pose of the end effector is also acceptable.

%#######################################################

\subsection{Using the target pose to determine the joint parameters}

The joint parameters, as was mentioned earlier, can be found by using indirect kinematics. The solution to an indirect kinematic problem is not unique. In the case of the UR5 there are three parameters that cause the multiple solutions. These parameters reflect the fact that although the center of end effector is well defined, the robot can reach the point in different ways. This means that arm, elbow and wrist of the robot can be in either of two positions each. This leads to a solution set of 8 different possible joint solutions.
Therefore a method of choosing an adequate solution out of the provided solution set has to be defined. \\
The chosen solution for the tracking problem is the one that will keep the movement of the different links to a minimum. Define a vector diff as the difference of the current joint positions to the joint positions in the solution set. The minimum of total motion is then defined as the vector which respective diff vector has the minimal euclidean norm.

%########################################################

\subsection{Error Handling}
To assure continuous operation error cases need to be handled. The central error that we handle is the one that happens when
the object is outside of the parallely reachable workspace, which leads to invalid solutions of the inverse kinematics. In that instance, we catch the invalid solution and define the weight of the calculated solution as infinite, so that it loses out in the minimization procedure. If \textbf{all} solultion are invalid, the is debug output to the console stating that the point is at an unreachable position.

\section{RESULTS}

%\subsection{Object localisation in Image Plane}

Tracking of one or more red objects with the described method was implemented and tracking of a red object could be observed. As assumed, the tracking area was limited to a sub area of the reachable workspace: it is a circle which has a radius just below the maximum workspace of the robot. This is due to the fact that the decision to keep the end effector parallel to the X-Y-Plane requires the 5th joint stay fixed. Thus the last link stays vertical, which shortens the total arm length. The speed of tracking is mainly limited by the frame rate and the movment time, the processing time is negligible. For development, a speed factor of 0.05 was used. Interpreting the camera image multiple times while the robot is moving leads to a faster reaction to position changes of the red test object. On the other hand the motion of the robot becomes less smooth. Tracking will hold as long as either no red object is detected by the camera or the position of the red object is physically not reachable by the robot. 
\section{DISCUSSION}
Our tracking concept allows for tracking on a parallel plane - this is  the simpler solution from the implementational point of view, since one coordinate in world space is fixed - but the cost is the loss of 3D tracking ability. An extension to this could be achieved by using a different projection model, for instance by setting a midpoint of a virtual sphere inside the workspace and limiting the movement of the UR5's end effector to the sphere's diameter. Another well established solution is the POSIT procedure [2], which has the advantage of calculating the pose of a specific object of known dimension from a single-view camera setup. While this provided a very interesting solution to the location estimation problem, the main drawback is that only the tracking of one very specific object would be possible.\\Another drawback of the method outlined in this paper is the limit it shows when the pose the robot would have to take is not physically realizable. If the conditions for the end effector were less rigid, the area in which tracking was possible would be considerably larger: a tilting of the end effector in the direction of the object point of the end effector after reaching the boundary of the workspace could be implemented. Then the target position of the end effector could be defined by a sphere around the red dot with defined radius. A possible problem with this approach are the cases in which the red object is just at the boundary of the usual workspace.
% \begin{table}[h]
% \caption{An Example of a Table}
% \label{table_example}
% \begin{center}
% \begin{tabular}{|c||c|}
% \hline
% One & Two\\
% \hline
% Three & Four\\
% \hline
% \end{tabular}
% \end{center}
% \end{table}

%    \begin{figure}[thpb]
%       \centering
%       \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
% }}
      %\includegraphics[scale=1.0]{figurefile}
%       \caption{Inductance of oscillation winding on amorphous
%        magnetic core versus DC bias magnetic field}
%       \label{figurelabel}
%    \end{figure}

% Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

% \section{CONCLUSIONS}

% A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

% \addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{APPENDIX}

% Appendixes should appear before the acknowledgment.

% \section*{ACKNOWLEDGMENT}

% The preferred spelling of the word ÒacknowledgmentÓ in America is without an ÒeÓ after the ÒgÓ. Avoid the stilted expression, ÒOne of us (R. B. G.) thanks . . .Ó  Instead, try ÒR. B. G. thanksÓ. Put sponsor acknowledgments in the unnumbered footnote on the first page.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

% \bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
 \bibitem{c1}  Daniel F. DeMenthon , Larry S. Davis, Model-Based Object Pose in 25 Lines of Code (1995) ,International Journal of Computer Vision, 1995
 \end{thebibliography}

\end{document}

### 2015-01-25 src/laurin/CplexSolver.java
Added stepwise optimization via linear programming

@@ -1,0 +1,393 @@
+package laurin;
+
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.Set;
+
+import ilog.concert.IloException;
+import ilog.concert.IloLinearNumExpr;
+import ilog.concert.IloNumExpr;
+import ilog.concert.IloNumVar;
+import ilog.concert.IloObjective;
+import ilog.concert.IloRange;
+import ilog.cplex.IloCplex;
+import utils.Config;
+import utils.Seed;
+import utils.Voxel;
+
+public class CplexSolver {
+	
+	private Voxel[][][] body;
+	private Seed[] seeds;
+	private int xSize;
+	private int ySize;
+	private int zSize;
+	double[] dwellTimes;
+	
+	IloCplex cplex;
+	
+	IloLinearNumExpr c_all_objective; // objective for dwell times
+	IloLinearNumExpr[] c_s_lower_voi_objectives; // obj. single lower slacks
+	IloLinearNumExpr[] c_s_upper_voi_objectives; // obj. single upper slacks
+	IloLinearNumExpr[] c_t_lower_voi_objectives; // obj. PTV lower slacks
+	IloLinearNumExpr[] c_t_upper_voi_objectives; // obj. PTV upper slacks
+	
+	IloNumVar[] t; // dwell times
+	IloNumVar[] s_lower; // single lower slacks
+	IloNumVar[] s_upper; // single upper slack
+	IloNumVar[] t_lower; // lower slacks for body parts
+	IloNumVar[] t_upper; // upper slacks for body parts
+	
+	IloRange[][][] s_lower_ranges; // single lower slack range 
+	IloRange[][][] s_upper_ranges; // single upper slack range
+	IloRange[] t_lower_ranges; // lower slack range per body part
+	IloRange[] t_upper_ranges; // upper slack range per body part
+	IloRange[][][] voi_lower_ranges; // lower dose ranges
+	IloRange[][][] voi_upper_ranges; // upper dose ranges
+
+	public CplexSolver(Voxel[][][] body, Seed[] seeds) throws IloException, IOException
+	{
+		this.body = body;
+		this.seeds = seeds;
+		this.xSize = body.length;
+		this.ySize = body[0].length;
+		this.zSize = body[0][0].length;
+		initialize();
+	}
+	
+	private void initialize() throws IloException, IOException
+	{
[…]

### 2015-01-25 src/dawid/Solver.java
LP läuft wieder

@@ -50,7 +50,7 @@ public class Solver {
 	public void solveLP() throws IloException
 	{
 		int nonZeroCounter = 0;
-		new LPTreatment(body, seeds);
+		//new LPTreatment(body, seeds);
 		//LPTreatment.solveLP();
 		//LPTreatment.solveLPMin();
 		LPTreatment.solveLPIT();

### 2015-01-25 src/erik/Solver.java

@@ -78,8 +78,9 @@ public class Solver {
 	
 	public void solveLP() throws IloException
 	{
-		int nonZeroCounter = 0;
-		new LPTreatment(body, seeds);
+		int nonZeroCounter = 0;		
+		
+		new LPTreatment(body, seeds, xBoundsTumor, yBoundsTumor, zBoundsTumor, dimensions);
 		//LPTreatment.solveLP();
 		//LPTreatment.solveLPMin();
 		LPTreatment.solveLPIT();

### 2015-01-25 src/laurin/TreatmentAnalyzer.java

@@ -315,7 +315,7 @@ public class TreatmentAnalyzer {
 	public void printResults() {
 		for (int i = 0; i < anatomies.size(); i++)
 		{
-			LogTool.print("Body type: " + i + ": minDose: " + minDoses[i] + ", maxDose: " + maxDoses[i] + ", avgDose: " + avgDoses[i],"notification");
+			LogTool.print("Body type " + i + ": minDose: " + minDoses[i] + ", maxDose: " + maxDoses[i] + ", avgDose: " + avgDoses[i],"notification");
 		}
 		LogTool.print("Tumor conformality index: " + conformalityIndex, "notification");
 		LogTool.print("Tumor homogenity index: " + homogenityIndex, "notification");

### 2015-01-25 src/thobi/LPTreatment.java

@@ -13,13 +13,21 @@ public class LPTreatment {
 	
 	static Voxel[][][] body;
 	static Seed[] seed;
+	static int[] xBoundsTumor;
+	static int[] yBoundsTumor;
+	static int[] zBoundsTumor;
+	static int[] dim;
 	
 	static boolean[] seedUsed = new boolean[Config.numberOfSeeds];
 	
-	public LPTreatment(Voxel[][][] body, Seed[] seed)
+	public LPTreatment(Voxel[][][] body, Seed[] seed, int[] x, int[] y, int[] z, int[] dim)
 	{
 		LPTreatment.body = body;
 		LPTreatment.seed = seed;
+		LPTreatment.xBoundsTumor = x;
+		LPTreatment.yBoundsTumor = y;
+		LPTreatment.zBoundsTumor = z;
+		LPTreatment.dim = dim;
 	}
 	
 	public static Seed[] getSeed()
@@ -174,6 +182,7 @@ public class LPTreatment {
 		IloCplex cplex = new IloCplex();
 		
 		IloNumExpr dosepart[] = new IloNumExpr[Config.numberOfSeeds];
+		
 		for(int i = 0; i < Config.numberOfSeeds; i++)
 		{
 			dosepart[i] = cplex.constant(0);
@@ -195,14 +204,14 @@ public class LPTreatment {
 		
 		//iterate over world
 		System.out.println("Adding constraints...");
-		for(int x = 0; x < Config.xDIM; x++)
+		for(int x = xBoundsTumor[0]; x < xBoundsTumor[1]; x++)
 		{
-			for(int y = 0; y < Config.yDIM; y++)
+			for(int y = yBoundsTumor[0]; y < yBoundsTumor[1]; y++)
 			{
-				for(int z = 0; z < Config.zDIM; z++)
+				for(int z = zBoundsTumor[0]; z < zBoundsTumor[1]; z++)
 				{
 					//check state of world
-					/*
+					
 					if(body[x][y][z].getBodyType() == Config.tumorType)
 					{
 						for(int i = 0; i < Config.numberOfSeeds; i++)
@@ -227,94 +236,6 @@ public class LPTreatment {
 							}
 						}
 					}
-					
-					*/
-					
-					switch(body[x][y][z].getBodyType())
-					{
-						case Config.normalType:
-						{
-							for(int i = 0; i < Config.numberOfSeeds; i++)
-							{
-								if(seedUsed[i])
-								{
[…]

### 2015-01-25 src/utils/Config.java

@@ -29,7 +29,7 @@ public class Config {
 	public static final double liverGoalDose = 30; 		//max
 	public static final double pancreasGoalDose = 0; 	//max
 	public static final double tumorGoalDose = 50; 		//min
-	
+	public static final double relaxDose = 5;
 	// Parameters for classification
 	public static final double VOXEL_DIST = 1; // in cm
 	public static final int TUMOR = 5; // BodyType for Tumor

### 2015-01-25 src/thobi/LPTreatment.java
test

@@ -336,7 +336,6 @@ public class LPTreatment {
 		//expressions
 		IloLinearNumExpr objective = cplex.linearNumExpr();
 		int tumor = 0;
-		//iterate over world
 		System.out.println("Adding constraints...");
 		for(int x = xBoundsTumor[0]; x < xBoundsTumor[1]; x++)
 		{

### 2015-01-25 src/utils/Config.java
Added Descriptions for body types and min/max doses

@@ -30,6 +30,21 @@ public class Config {
 	public static final double pancreasGoalDose = 0; 	//max
 	public static final double tumorGoalDose = 50; 		//min
 	
+	public static final double normalMinDose = 0;
+	public static final double spineMinDose = 0;
+	public static final double liverMinDose = 0;
+	public static final double pancreasMinDose = 0;
+	public static final double tumorMinDose = 30;
+	
+	public static final double normalMaxDose = 40;
+	public static final double spineMaxDose = 30;
+	public static final double liverMaxDose = 50;
+	public static final double pancreasMaxDose = 30;
+	public static final double tumorMaxDose = 70;
+	
+	public static final String[] bodyTypeDescriptions = {"Normal", "Spine", "Liver", "Pancreas", "Tumor"};
+	
+	
 	// Parameters for classification
 	public static final double VOXEL_DIST = 1; // in cm
 	public static final int TUMOR = 5; // BodyType for Tumor

### 2015-01-25 src/erik/BodyAnalyzer.java
Added further methods to BodyAnalyzer

@@ -1,5 +1,9 @@
 package erik;
 
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.Set;
+
 import utils.Config;
 import utils.Voxel;
 
@@ -266,5 +270,89 @@ public class BodyAnalyzer {
 	public void setVoxelCountTumor(int voxelCountTumor) {
 		this.voxelCountTumor = voxelCountTumor;
 	}
-
+	
+	/**
+	 * Creates sets of voxels that share the same body type.
+	 * 
+	 * @param body array of body voxels
+	 * 
+	 * @return ArrayList of sets of voxels that share a body type
+	 */
+	public static ArrayList<Set<Voxel>> splitBodyTypes (Voxel[][][] body) {
+		ArrayList<Set<Voxel>> anatomies = new ArrayList<Set<Voxel>>();
+		for (int i = Config.normalType; i <= Config.tumorType; i++)
+		{
+			anatomies.add(new HashSet<Voxel>());
+		}
+		
+		for(int x = 0; x < body.length; x++) {				
+			for(int y = 0; y < body[0].length; y++) {		
+				for(int z = 0; z < body[0][0].length; z++) {		
+					anatomies.get(body[x][y][z].getBodyType()-1).add(body[x][y][z]);									
+				}
+			}	
+		}	
+		return anatomies;
+	}
+	
+	/**
+	 * Checks if voxel is outter voxel.
+	 * An outter voxel is a voxel that has at least one neighboring pixel (in 6-neighborhood) that either
+	 * 		- exceeds global body dimensions
+	 * 		- is of another body type
+	 * 
+	 * @return true if voxel is outter voxel,
+	 * 		   false otherwise
+	 **/
+	private static boolean isOutterVoxel(Voxel[][][] body, Voxel voxel)
+	{
+		int x = (int) voxel.getX();
+		int y = (int) voxel.getY();
+		int z = (int) voxel.getZ();
+		int bodyType = voxel.getBodyType();
+		
+		// voxel is outter body pixel
+		if (x-1 < 0 
+				|| x+1 > body.length
+				|| y-1 < 0
+				|| y+1 > body[0].length
+				|| z-1 < 0
+				|| z+1 > body[0][0].length)
+		{
+			return true;
+		}
+		
+		// voxel is on body part surface
+		if (body[x-1][y][z].getBodyType() != bodyType
[…]

### 2015-01-25 src/laurin/TreatmentAnalyzer.java
Method splitBodyTypes() transferred to BodyAnalyzer

@@ -6,6 +6,8 @@ import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.Set;
 
+import erik.BodyAnalyzer;
+
 import utils.Config;
 import utils.LogTool;
 import utils.Seed;
@@ -52,7 +54,7 @@ public class TreatmentAnalyzer {
 		this.seeds = seeds;
 		this.title = title;
 		
-		splitBodyTypes();
+		this.anatomies = BodyAnalyzer.splitBodyTypes(body);
 		irradiate();
 		analyzeAll();
 	}
@@ -120,27 +122,6 @@ public class TreatmentAnalyzer {
 	public void setCoverage(double coverage) 
 	{
 		this.coverage = coverage;
-	}
-
-	/**
-	 * Creates sets of voxels for each body type
-	 */
-	private void splitBodyTypes () {
-		ArrayList<Set<Voxel>> anatomies = new ArrayList<Set<Voxel>>();
-		for (int i = Config.normalType; i <= Config.tumorType; i++)
-		{
-			anatomies.add(new HashSet<Voxel>());
-		}
-		
-		for(int x = 0; x < this.dimensions[0]; x++) {				
-			for(int y = 0; y < this.dimensions[1]; y++) {		
-				for(int z = 0; z < this.dimensions[2]; z++) {		
-					anatomies.get(this.body[x][y][z].getBodyType()-1).add(body[x][y][z]);									
-				}
-			}	
-		}
-		
-		this.anatomies = anatomies;
 	}
 	
 	/**

### 2015-01-25 src/utils/Coordinate.java
Transferred distance calculation to Coordinate and added auxiliary methods to Coordinate

@@ -2,11 +2,14 @@ package utils;
 
 import java.io.Serializable;
 
+import org.apache.commons.math3.geometry.euclidean.threed.Vector3D;
+
 public class Coordinate implements Serializable {
 	private static final long serialVersionUID = 12L;
 	private double x;
 	private double y;
 	private double z;
+	public static final double GRID_RESOLUTION = 0.2;  // Distance between two voxel in cm
 	
 	public Coordinate(double x, double y , double z) {
 		this.x = x;
@@ -34,5 +37,36 @@ public class Coordinate implements Serializable {
 	public void setZ(double z) {
 		this.z = z;
 	}
-
+	
+	/**
+	 * Calculates coordinate on line through base and other.
+	 * 
+	 * @param base		start point
+	 * @param other		other point that defines line
+	 * @param t			distance to start point in direction of other
+	 * 
+	 * @return	point on line
+	 */
+	public static Coordinate getPointOnLine(Coordinate base, Coordinate other, double t)
+	{
+		Vector3D vBase = base.ToVector();
+		Vector3D vOther = other.ToVector();
+		Vector3D vPoint = vBase.add(vOther.subtract(vBase).scalarMultiply(t));
+		return vectorToCoordinate(vPoint);
+	}
+	
+	public static Coordinate vectorToCoordinate(Vector3D vector)
+	{
+		return new Coordinate(vector.getX(), vector.getY(), vector.getZ());
+	}
+	
+	public Vector3D ToVector()
+	{
+		return new Vector3D(this.getX(), this.getY(), this.getZ());
+	}
+	
+	public double distanceToCoordiante(Coordinate position) {
+		
+	    return(( Math.sqrt( Math.pow(this.getX()-position.getX(), 2) + Math.pow(this.getY()-position.getY(), 2) + Math.pow(this.getZ()-position.getZ(), 2) ) ) * GRID_RESOLUTION);
+	}
 }

### 2015-01-25 src/utils/Voxel.java

@@ -5,7 +5,6 @@ import java.io.Serializable;
 public class Voxel implements Serializable {
 	private static final long serialVersionUID = 21L;
 	
-	public static final double GRID_RESOLUTION = 0.2;  // Distance between two voxel in cm
 	private double maxDosis;
 	private double minDosis;
 	private double goalDosis;
@@ -28,9 +27,7 @@ public class Voxel implements Serializable {
 	 * The distance to the voxel
 	 */
 	public double distanceToVoxel (Coordinate position) {
-		
-	    return(( Math.sqrt( Math.pow(this.getCoordinate().getX()-position.getX(), 2) + Math.pow(this.getCoordinate().getY()-position.getY(), 2) + Math.pow(this.getCoordinate().getZ()-position.getZ(), 2) ) ) * GRID_RESOLUTION);
-				
+	    return this.getCoordinate().distanceToCoordiante(position);
 	}
 	
 	

### 2015-01-25 src/utils/Needle.java
Added methods for placing needles prior to treatment and choose best needles after treatment

@@ -1,10 +1,18 @@
 package utils;
 
+import java.io.Serializable;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.Collections;
+import java.util.Comparator;
 import java.util.HashSet;
 import java.util.Set;
 
-public class Needle {
-	
+import erik.BodyAnalyzer;
+
+public class Needle implements Serializable {
+
+	private static final long serialVersionUID = 1L;
 	private Set<Seed> seeds;
 	
 	public Needle()
@@ -20,5 +28,134 @@ public class Needle {
 	public void addSeed(Seed seed)
 	{
 		seeds.add(seed);
+	}
+	
+	public int getSize()
+	{
+		return this.seeds.size();
+	}
+	
+	/**
+	 * Create array of seeds from needles
+	 * 
+	 * All seeds in list of needles will be placed in a single array.
+	 * 
+	 * @param needles 		needles whose seeds shall be collected
+	 * 
+	 * @return				seeds
+	 */
+	public static Seed[] needlesToSeeds(ArrayList<Needle> needles)
+	{
+		Set<Seed> seeds = new HashSet<Seed>();
+		
+		for (Needle needle : needles)
+		{
+			for (Seed seed : needle.getSeeds())
+			{
+				seeds.add(seed);
+			}
+		}
+		return seeds.toArray(new Seed[seeds.size()]);
+	}
+	
+	/**
+	 * Gets best needles from set of needles.
+	 * 
+	 * The best needles are specified as the numNeedles needles 
+	 * with the largest sum of its assigned seeds' dwell times.
+	 * 
+	 * @param numNeedles      number of best needles to be returned
+	 * 
+	 * @return best needles
+	 */
+	public static ArrayList<Needle> getBestNeedles(ArrayList<Needle> needles, int numNeedles )
+	{
+		ArrayList<Needle> bestNeedles = new ArrayList<Needle>();
+
+		Collections.sort(needles, new Comparator<Needle>() {
+		    public int compare(Needle n1, Needle n2) {    	
+		    	Double needleSum1 = new Double(0.0);
+		    	Double needleSum2 = new Double(0.0);
+		    	for (Seed seed : n1.getSeeds())
+		    		needleSum1 += seed.getDurationMilliSec();
+		    	for (Seed seed : n2.getSeeds())
+		    		needleSum2 += seed.getDurationMilliSec();
+	
+		        return needleSum2.compareTo(needleSum1);
+		    }
+		});
+		
[…]

### 2015-01-25 src/utils/Needle.java
Added angle depency for dose calculation

@@ -2,25 +2,26 @@ package utils;
 
 import java.io.Serializable;
 import java.util.ArrayList;
-import java.util.Arrays;
 import java.util.Collections;
 import java.util.Comparator;
 import java.util.HashSet;
 import java.util.Set;
+
+import org.apache.commons.math3.geometry.euclidean.threed.Vector3D;
 
 import erik.BodyAnalyzer;
 
 public class Needle implements Serializable {
 
 	private static final long serialVersionUID = 1L;
-	private Set<Seed> seeds;
+	private ArrayList<Seed> seeds;
 	
 	public Needle()
 	{
-		seeds = new HashSet<Seed>();
+		seeds = new ArrayList<Seed>();
 	}
 	
-	public Set<Seed> getSeeds()
+	public ArrayList<Seed> getSeeds()
 	{
 		return seeds;
 	}
@@ -33,6 +34,22 @@ public class Needle implements Serializable {
 	public int getSize()
 	{
 		return this.seeds.size();
+	}
+	
+	/**
+	 * Calculates direction of needle as difference between second and first seeds' coordinates.
+	 * 
+	 * @return direction of needle
+	 */
+	public Vector3D getDirection()
+	{
+		if (this.getSize() < 2)
+			return null;
+		else
+		{
+			// subtract second coordinate from first
+			return (seeds.get(1).getCoordinate().ToVector()).subtract(seeds.get(0).getCoordinate().ToVector());
+		}
 	}
 	
 	/**

### 2015-01-25 src/utils/Seed.java

@@ -142,5 +142,26 @@ public class Seed extends Voxel {
 		
 		return rearrangedSeeds.toArray(new Seed[rearrangedSeeds.size()]);
 	}
+	
+	
+	/**
+	 * Calculates the angle (phi) for dose calculation
+	 * 
+	 * The angle can only be calculated if the seed is assigned to a needle.
+	 * 
+	 * @param voxel		voxel for which angle shall be determined
+	 * @return			angle in degrees
+	 */
+	public double getPhi(Voxel voxel)
+	{
+		if (this.needle == null)
+			return 90;
+		else
+		{
+			Vector3D vNeedle = needle.getDirection();
+			Vector3D vLineSeedToVoxel = (voxel.getCoordinate().ToVector()).subtract(this.getCoordinate().ToVector());
+			return (Vector3D.angle(vNeedle, vLineSeedToVoxel) * 180 * Math.PI);
+		}
+	}
 
 }

### 2015-01-25 src/utils/Voxel.java

@@ -43,12 +43,13 @@ public class Voxel implements Serializable {
 	 * The postition where the seed is 
 	 * @param durationMilliSec
 	 * The time the seed radiates
+	 * @param phi
+	 * angle phi 
 	 * @return
 	 */
-	public double radiationIntensity(Coordinate position, double durationMilliSec){
+	public double radiationIntensity(Coordinate position, double durationMilliSec, double phi){
 		double distance = distanceToVoxel(position);
 		
-		double phi = 90;  //TODO Magic number
 		double dose= 0;
 		
 		if(distance > 10) {
@@ -64,6 +65,10 @@ public class Voxel implements Serializable {
 		}
 		
 		return dose;
+	}
+	
+	public double radiationIntensity(Coordinate position, double durationMilliSec){
+		return radiationIntensity(position, durationMilliSec, 90);
 	}
 	
 	public double radiationIntensity(double r, double durationMilliSec){

### 2015-01-26 src/erik/BodyAnalyzer.java
changed Voxel.GridResolution to Coordinate.GridResolution

@@ -5,6 +5,7 @@ import java.util.HashSet;
 import java.util.Set;
 
 import utils.Config;
+import utils.Coordinate;
 import utils.Voxel;
 
 /**
@@ -147,8 +148,8 @@ public class BodyAnalyzer {
 			return xBoundsTumor;
 		} else if ( type == 2) {
 		
-			lowerBound = xBoundsTumor[0] - (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
-			upperBound = xBoundsTumor[1] + (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
+			lowerBound = xBoundsTumor[0] - (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
+			upperBound = xBoundsTumor[1] + (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
 			
 			if(lowerBound > 0) {
 				result[0] = lowerBound;
@@ -192,8 +193,8 @@ public class BodyAnalyzer {
 			return yBoundsTumor;
 		} else if ( type == 2) {
 		
-			lowerBound = yBoundsTumor[0] - (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
-			upperBound = yBoundsTumor[1] + (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
+			lowerBound = yBoundsTumor[0] - (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
+			upperBound = yBoundsTumor[1] + (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
 			
 			if(lowerBound > 0) {
 				result[0] = lowerBound;
@@ -236,8 +237,8 @@ public class BodyAnalyzer {
 			return zBoundsTumor;
 		} else if ( type == 2) {
 		
-			lowerBound = zBoundsTumor[0] - (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
-			upperBound = zBoundsTumor[1] + (int) (Config.treatmentRange/Voxel.GRID_RESOLUTION);
+			lowerBound = zBoundsTumor[0] - (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
+			upperBound = zBoundsTumor[1] + (int) (Config.treatmentRange/Coordinate.GRID_RESOLUTION);
 			
 			if(lowerBound > 0) {
 				result[0] = lowerBound;

### 2015-01-26 src/sebastian/PlannerGUI.java
- Created GUI for step by step execution  - modified db and treatment entry accordingly

@@ -1,0 +1,360 @@
+package sebastian;
+
+import ilog.concert.IloException;
+
+import java.awt.BorderLayout;
+import java.awt.FlowLayout;
+import java.awt.GridLayout;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.awt.event.MouseEvent;
+import java.awt.event.MouseListener;
+import java.io.File;
+import java.util.Vector;
+
+import javax.swing.BorderFactory;
+import javax.swing.BoxLayout;
+import javax.swing.JButton;
+import javax.swing.JFileChooser;
+import javax.swing.JFrame;
+import javax.swing.JList;
+import javax.swing.JOptionPane;
+import javax.swing.JPanel;
+import javax.swing.JScrollPane;
+import javax.swing.border.EtchedBorder;
+
+import utils.Config;
+import utils.Coordinate;
+import utils.LogTool;
+import utils.RandGenerator;
+import utils.Seed;
+import erik.BodyAnalyzer;
+import erik.Solver;
+
+public class PlannerGUI extends JFrame implements ActionListener, MouseListener {
+	private static final long serialVersionUID = 1L;
+	private JButton loadBody = null;
+	private JButton measureBody = null;
+	private JButton measureAll = null;
+	private JButton closestSeeds = null;
+	private JButton randomSeeds = null;
+	private JButton solveLP = null;
+	private JButton solveSA = null;
+	private JButton solveGA = null;
+	private JList<String> bodies = null;
+	private JList<String> treatments = null;
+	private JPanel panel = null;
+	private SimpleDB db = null;
+	private Seed[] seeds = null;
+	
+	private void updateBodies () {
+		bodies.removeAll();
+		if (db.getBodySize() > 0) {
+		  bodies.setListData(db.getBodies ());
+		}
+		else {
+			bodies.setListData(new Vector<String> ());
+		}
+		this.revalidate ();
+	}
+	
+	private void updateTreatments () {
+		treatments.removeAll();
+		if (db.getTreatmentSize() > 0) {
+		  treatments.setListData(db.getTreatments ());
+		}
+		else {
+			treatments.setListData(new Vector<String> ());
+		}
+		this.revalidate();
+	}
+	
[…]

### 2015-01-27 src/utils/Seed.java
Changed return type of method rearrangeSeeds

@@ -57,10 +57,11 @@ public class Seed extends Voxel {
 	 * 	
      * @return seeds in rearranged positions
 	 */
-	public static Seed[] rearrangeSeeds(Seed[] seeds, int numberOfNeedles, double maxDistanceToLine, Set<Needle> needles)
+	public static ArrayList<Needle> rearrangeSeeds(Seed[] seeds, int numberOfNeedles, double maxDistanceToLine)
 	{
 		ArrayList<Seed> rearrangedSeeds = new ArrayList<Seed>();
 		ArrayList<Seed> originalSeeds = new ArrayList<Seed>(Arrays.asList(seeds));
+		ArrayList<Needle> needles = new ArrayList<Needle>();
 		
 		// remove seeds with zero dwell time
 		for (Iterator<Seed> iterator = originalSeeds.iterator(); iterator.hasNext(); ) {
@@ -85,7 +86,6 @@ public class Seed extends Voxel {
 		// create needles until no are seeds left or maximum needle number reached
 		while (originalSeeds.size() > 0 && needleCounter < numberOfNeedles)
 		{
-			System.out.println(originalSeeds.size());
 			Needle needle = new Needle();
 			if (originalSeeds.size() > 1)
 			{
@@ -140,7 +140,7 @@ public class Seed extends Voxel {
 			needleCounter++;
 		}
 		
-		return rearrangedSeeds.toArray(new Seed[rearrangedSeeds.size()]);
+		return needles;
 	}
 	
 	

### 2015-01-27 src/erik/Population.java
Fixed problem with the body

@@ -3,6 +3,7 @@ package erik;
 import java.util.Random;
 
 import utils.Config;
+import utils.LogTool;
 
 public class Population
 {
@@ -161,15 +162,17 @@ public class Population
             
             // reevaluate current population
             pop.evaluate();
-            System.out.print("Iteration : "+iter+" "+"Total Fitness = " + pop.totalFitness);
-            System.out.println(" ; Best Fitness = " + pop.findBestIndividual().getFitnessValue()); 
+           // System.out.print("Iteration : "+iter+" "+"Total Fitness = " + pop.totalFitness);
+           // System.out.println(" ; Best Fitness = " + pop.findBestIndividual().getFitnessValue()); 
+           
+            LogTool.print("Iteration : "+iter, "notification");
+            LogTool.print("Improvement: "+Math.abs(bestResultOld-bestResult)+" Cancel counter= " + counter, "debug");
             
             if(pop.findBestIndividual().getFitnessValue() < bestResult) {
             	bestResultOld = bestResult;
             	bestResult =  pop.findBestIndividual().getFitnessValue();
             }
             
-            System.out.println("bestOld-best: "+Math.abs(bestResultOld-bestResult)+"counter= " + counter);
             if(Math.abs(bestResultOld-bestResult) <= Config.cancelValue) {
             	counter++;	
             } else {

### 2015-01-27 src/sebastian/PlannerGUI.java

@@ -28,6 +28,7 @@ import utils.Coordinate;
 import utils.LogTool;
 import utils.RandGenerator;
 import utils.Seed;
+import utils.Voxel;
 import erik.BodyAnalyzer;
 import erik.Solver;
 
@@ -237,6 +238,7 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 
 	private void solveGA () {
 		BodyEntry entry = null;
+		Voxel[][][] body = null;
 		if (bodies.isSelectionEmpty()) {
 			JOptionPane.showMessageDialog(this, "No body selected!");
 			return;

### 2015-01-27 src/sebastian/PlannerGUI.java
Minor bugfixes for the gui; Print redirection of the logtool into the gui

@@ -21,7 +21,9 @@ import javax.swing.JList;
 import javax.swing.JOptionPane;
 import javax.swing.JPanel;
 import javax.swing.JScrollPane;
+import javax.swing.JTextArea;
 import javax.swing.border.EtchedBorder;
+import javax.swing.filechooser.FileFilter;
 
 import utils.Config;
 import utils.Coordinate;
@@ -47,6 +49,11 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 	private JPanel panel = null;
 	private SimpleDB db = null;
 	private Seed[] seeds = null;
+	private JTextArea out = null;
+	
+	public void appendText (String text) {
+		out.append (text);
+	}
 	
 	private void updateBodies () {
 		bodies.removeAll();
@@ -71,6 +78,7 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 	}
 	
 	private void init () {
+		out = new JTextArea ();
 		panel = new JPanel ();
 		loadBody = new JButton ("Load Body");
 		measureBody = new JButton ("Measure Body");
@@ -81,16 +89,13 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 		solveSA = new JButton ("Simulated Annealing");
 		solveGA = new JButton ("Generic Algorithm");
 		bodies = new JList<String> ();
-		treatments = new JList<String> ();
+		treatments = new JList<String> ();		
 		
 		bodies.addMouseListener(this);
 		bodies.setToolTipText("Bodies");
 		treatments.addMouseListener(this);
 		treatments.setToolTipText("Treatments");
-		
-		updateBodies ();
-		updateTreatments ();
-		
+				
 		loadBody.addActionListener(this);
 		loadBody.setActionCommand("loadBody");
 		measureBody.addActionListener(this);
@@ -140,18 +145,23 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 		tmp4.add (new JScrollPane (bodies));
 		tmp4.setBorder(BorderFactory.createEtchedBorder(EtchedBorder.LOWERED));		
 		panel.add (tmp4, BorderLayout.WEST);
+		
+		panel.add(new JScrollPane (out), BorderLayout.CENTER);
 	}
 	
 	public PlannerGUI () {
 		super ("Treatment Planning");
-		db = new SimpleDB ();
[…]

### 2015-01-27 src/sebastian/SimpleDB.java

@@ -179,7 +179,7 @@ public class SimpleDB {
 		  tclDists[i] = 50000;
 	  }
 	  for (int i = 0; i < bEntry.getDimensions()[0]; i++) {
-		  System.out.println ("" + i);
+		  System.out.println(String.format("Measuring done: %f", (float) i / (float) bEntry.getDimensions ()[0]));
 		  for (int j = 0; j < bEntry.getDimensions()[1]; j++) {
 			  for (int k = 0; k < bEntry.getDimensions()[2]; k++) {
 				  for (int l = 0; l < tumorVoxel.size(); l++) {

### 2015-01-27 src/utils/LogTool.java

@@ -1,24 +1,37 @@
 package utils;
 
-import java.util.List;
+import sebastian.PlannerGUI;
 
 public class LogTool {
-	
+	private static PlannerGUI gui = null;
 	
 	public static void print(String str, String level) {
-		
-		if(level.equals("error") && Config.errors) {
-			System.out.println(str);
-		} else if(level.equals("warning") && Config.warnings) {
-			System.out.println(str);
-		} else if (level.equals("debug") && Config.debug) {
-			System.out.println(str);
-		} else if (level.equals("notification") && Config.notification) {
-			System.out.println(str);
+		if (gui == null) {
+			if(level.equals("error") && Config.errors) {
+				System.out.println(str);
+			} else if(level.equals("warning") && Config.warnings) {
+				System.out.println(str);
+			} else if (level.equals("debug") && Config.debug) {
+				System.out.println(str);
+			} else if (level.equals("notification") && Config.notification) {
+				System.out.println(str);
+			}
+		}
+		else {
+			if(level.equals("error") && Config.errors) {
+				gui.appendText(str + "\n");
+			} else if(level.equals("warning") && Config.warnings) {
+				gui.appendText(str + "\n");
+			} else if (level.equals("debug") && Config.debug) {
+				gui.appendText(str + "\n");
+			} else if (level.equals("notification") && Config.notification) {
+				gui.appendText(str + "\n");
+			}
 		}
 		
 	}
 
-
-
+	public static void addGUI(PlannerGUI plannerGui) {
+		gui = plannerGui;
+	}
 }

### 2015-01-27 src/sebastian/PlannerGUI.java
added display methods

@@ -15,6 +15,7 @@ import java.util.Vector;
 import javax.swing.BorderFactory;
 import javax.swing.BoxLayout;
 import javax.swing.JButton;
+import javax.swing.JComboBox;
 import javax.swing.JFileChooser;
 import javax.swing.JFrame;
 import javax.swing.JList;
@@ -44,12 +45,17 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 	private JButton solveLP = null;
 	private JButton solveSA = null;
 	private JButton solveGA = null;
+	private JButton trainWeights = null;
+	private JButton display = null;
+	private JComboBox<ScatterDisplay.ChartType> dispType = null;
 	private JList<String> bodies = null;
 	private JList<String> treatments = null;
 	private JPanel panel = null;
 	private SimpleDB db = null;
 	private Seed[] seeds = null;
 	private JTextArea out = null;
+	private Voxel[][][] optBody = null;
+	private int[] optDims = null;
 	
 	public void appendText (String text) {
 		out.append (text);
@@ -89,13 +95,20 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 		solveSA = new JButton ("Simulated Annealing");
 		solveGA = new JButton ("Generic Algorithm");
 		bodies = new JList<String> ();
-		treatments = new JList<String> ();		
-		
+		treatments = new JList<String> ();
+		trainWeights = new JButton ("Train Weights");
+		display = new JButton ("Show Display");
+		dispType = new JComboBox<ScatterDisplay.ChartType>(ScatterDisplay.ChartType.values());
+				
 		bodies.addMouseListener(this);
 		bodies.setToolTipText("Bodies");
 		treatments.addMouseListener(this);
 		treatments.setToolTipText("Treatments");
 				
+		trainWeights.addActionListener (this);
+		trainWeights.setActionCommand("trainWeights");
+		display.addActionListener(this);
+		display.setActionCommand("display");
 		loadBody.addActionListener(this);
 		loadBody.setActionCommand("loadBody");
 		measureBody.addActionListener(this);
@@ -132,11 +145,17 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 		tmp2.add(solveLP);
 		tmp2.add(solveSA);
[…]

### 2015-01-27 src/erik/Population.java
slide update

@@ -1,5 +1,8 @@
 package erik;
 
+import java.text.DateFormat;
+import java.text.SimpleDateFormat;
+import java.util.Date;
 import java.util.Random;
 
 import utils.Config;
@@ -125,6 +128,7 @@ public class Population
         // main loop
         int count;
         while (done) {
+        	long start = System.currentTimeMillis();
             count = 0;
             iter++;
 
@@ -162,11 +166,11 @@ public class Population
             
             // reevaluate current population
             pop.evaluate();
-           // System.out.print("Iteration : "+iter+" "+"Total Fitness = " + pop.totalFitness);
-           // System.out.println(" ; Best Fitness = " + pop.findBestIndividual().getFitnessValue()); 
+            System.out.print("Iteration : "+iter+" "+"Total Fitness = " + pop.totalFitness);
+            System.out.println(" ; Best Fitness = " + pop.findBestIndividual().getFitnessValue()); 
            
-            LogTool.print("Iteration : "+iter, "notification");
-            LogTool.print("Improvement: "+Math.abs(bestResultOld-bestResult)+" Cancel counter= " + counter, "debug");
+            //LogTool.print("Iteration : "+iter, "notification");
+            //LogTool.print("Improvement: "+Math.abs(bestResultOld-bestResult)+" Cancel counter= " + counter, "debug");
             
             if(pop.findBestIndividual().getFitnessValue() < bestResult) {
             	bestResultOld = bestResult;
@@ -182,6 +186,11 @@ public class Population
             if(counter > 5) {
             	done = false;
             }
+            long end = System.currentTimeMillis();
+            Date date = new Date(end - start);
+    		DateFormat formatter = new SimpleDateFormat("mm:ss:SSS");
+    		String dateFormatted = formatter.format(date);
+    		System.out.println("Iteration Runtime: " + dateFormatted);
             
         }
        

### 2015-01-27 src/erik/TreatmentPlanner.java

@@ -126,11 +126,13 @@ public class TreatmentPlanner {
 		ta.analyzeAll();
 		ta.printResults();
 		// runtime measurement
+		
 		Date date = new Date(end - start);
 		DateFormat formatter = new SimpleDateFormat("mm:ss:SSS");
 		String dateFormatted = formatter.format(date);
 		System.out.println("Runtime: " + dateFormatted);
 
+		
 		ScatterDisplay display5 = new ScatterDisplay(ChartType.BodyType);
 		display5.fill(Solver.body, entry.getDimensions()[0], entry.getDimensions()[1],
 				entry.getDimensions()[2]);

### 2015-01-28 src/sebastian/CrossDistanceCalculator.java
- added multithread-support for measurement - added display - fixed display errors

@@ -7,15 +7,16 @@ import utils.Coordinate;
 
 public class CrossDistanceCalculator implements Callable<double[]> {
     private BodyEntry body = null;
-    private int z;
+    private int z, max;
     private ArrayList<double[]> tumor = null;
     private double[] partRes = null;
 	
 	
-	public CrossDistanceCalculator (BodyEntry body, int z, ArrayList<double[]> tumor) {
+	public CrossDistanceCalculator (BodyEntry body, int z, int max, ArrayList<double[]> tumor) {
 		this.body = body;
 		this.tumor = tumor;
 		this.z = z;
+		this.max = max;
 		partRes = new double[body.getMaxType() + 1];
 		for (int i = 0; i < partRes.length; i++) {
 			partRes[i] = 50000;
@@ -63,7 +64,7 @@ public class CrossDistanceCalculator implements Callable<double[]> {
 	
 	@Override
 	public double[] call() throws Exception {
-		System.out.println(z);
+		System.out.println(z + "/" + max);
 		crossCalculate ();		
 		return partRes;
 	}

### 2015-01-28 src/sebastian/PlannerGUI.java

@@ -170,7 +170,6 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 	
 	public PlannerGUI () {
 		super ("Treatment Planning");
-		LogTool.addGUI (this);
 		Runtime.getRuntime().addShutdownHook(new Thread () {
 			public void run () {
 				db.close();
@@ -214,6 +213,7 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 			JOptionPane.showMessageDialog(this, "No body selected!");
 			return;
 		}
+		
 		
 		db.classify (bodies.getSelectedValue());
 		updateTreatments ();
@@ -360,6 +360,22 @@ public class PlannerGUI extends JFrame implements ActionListener, MouseListener 
 			ScatterDisplay disp = new ScatterDisplay ((ScatterDisplay.ChartType) dispType.getSelectedItem());
 			disp.fill(optBody, optDims[0], optDims[1], optDims[2]);
 			disp.display();
+		}
+		else {
+			if (bodies.isSelectionEmpty()) {
+				JOptionPane.showMessageDialog (this, "No optimized data stored");
+			}
+			else {
+				BodyEntry entry = db.getBodyByName(bodies.getSelectedValue());
+				if (entry != null) {
+					ScatterDisplay disp = new ScatterDisplay (ScatterDisplay.ChartType.BodyType);
+					disp.fill(entry.getBodyArray(), entry.getDimensions()[0], entry.getDimensions()[1], entry.getDimensions()[2]);
+					disp.display ();
+				}
+				else {
+					JOptionPane.showMessageDialog (this, "No body selected!");
+				}
+			}
 		}
 	}
 	

### 2015-01-28 src/sebastian/ScatterDisplay.java

@@ -112,16 +112,16 @@ public class ScatterDisplay{
         		color = new Color (0.8f, 0.8f, 0.8f, 0f);
         		break;
         	case 2: // spine
-        		color = new Color (1f, 1f, 0, 1f);
+        		color = new Color (1f, 1f, 0, 0.2f);
         		break;
         	case 3: // liver
-        		color = new Color (0f, 1f, 0f, 1f);
+        		color = new Color (0f, 1f, 0f, 0.2f);
         		break;
         	case 4: // pancreas
-        		color = new Color (0f, 0f, 0f, 1f);
+        		color = new Color (0f, 0f, 0f, 0.2f);
         		break;
         	case 5: // tumor
-        		color = new Color (1f, 0f, 0f, 1f);
+        		color = new Color (1f, 0f, 0f, 0.2f);
         		break;
     		default:
     			color = new Color (0f, 0f, 0f, 0f);

### 2015-01-28 src/sebastian/SimpleDB.java

@@ -9,6 +9,11 @@ import java.io.ObjectInputStream;
 import java.io.ObjectOutputStream;
 import java.util.ArrayList;
 import java.util.Vector;
+import java.util.concurrent.CompletionService;
+import java.util.concurrent.ExecutionException;
+import java.util.concurrent.ExecutorCompletionService;
+import java.util.concurrent.ExecutorService;
+import java.util.concurrent.Executors;
 
 import utils.Config;
 import utils.Coordinate;
@@ -93,6 +98,7 @@ public class SimpleDB {
 			    classify (getBodyByName (folder.getName ()));
 			    
 			  }
+			  storeDB();
 		  }
 		  else {
 			  LogTool.print ("Entry already exists: " + folder.getName (),"debug");
@@ -173,13 +179,42 @@ public class SimpleDB {
 	  
 	  // Calculate closest distance between tumor and other volumes
 	  LogTool.print ("Calculating closest distances", "debug");
+	  ExecutorService threadPool = Executors.newFixedThreadPool(Config.numberOfThreads);
+  	  CompletionService<double[]> pool = new ExecutorCompletionService<double[]>(threadPool);
+  	
+  	  for (int i = 0; i < tclDists.length; i++) {
+		  tclDists[i] = 50000;
+	  }
+  	  
+  	  // Create tasks
+  	  for (int i = 0; i < bEntry.getDimensions()[2]; i++) {
+  		  pool.submit (new CrossDistanceCalculator(bEntry, i, bEntry.getDimensions()[2], tumorVoxel));
+  	  }
+  	  
+  	  // Gather results
+  	  double[] partRes = null;
+  	  for (int i = 0; i < bEntry.getDimensions()[2]; i++) {
+  		  try {
+			partRes = pool.take().get();
+			for (int j = 0; j < partRes.length; j++) {
+				if (partRes[j] < tclDists[j]) {
+					tclDists[j] = partRes[j];
+				}
+			}
+		} catch (InterruptedException e) {
+			LogTool.print("Error retreiving result: " + e, "error");
+		} catch (ExecutionException e) {
+			LogTool.print("Error accessing result: " + e, "error");
+		}
+  	  }
+  	  /*
 	  int type = -1;
 	  double dist = -1;
 	  for (int i = 0; i < tclDists.length; i++) {
 		  tclDists[i] = 50000;
 	  }
[…]

### 2015-01-28 src/sebastian/TreatmentEntry.java

@@ -27,6 +27,7 @@ public class TreatmentEntry implements Serializable {
   private double[][] volumeCenters = null;
   private double[] tumorCenterDistances = null;
   private double[] tumorClosestDistances = null;
+  private double[] coverage = null;
   
   public TreatmentEntry (String name) {
 	this.name = name;

### 2015-01-28 src/utils/LogTool.java

@@ -1,37 +1,17 @@
 package utils;
 
-import sebastian.PlannerGUI;
 
 public class LogTool {
-	private static PlannerGUI gui = null;
 	
-	public static void print(String str, String level) {
-		if (gui == null) {
-			if(level.equals("error") && Config.errors) {
-				System.out.println(str);
-			} else if(level.equals("warning") && Config.warnings) {
-				System.out.println(str);
-			} else if (level.equals("debug") && Config.debug) {
-				System.out.println(str);
-			} else if (level.equals("notification") && Config.notification) {
-				System.out.println(str);
-			}
+	public static void print(String str, String level) {	
+		if(level.equals("error") && Config.errors) {
+			System.out.println(str);
+		} else if(level.equals("warning") && Config.warnings) {
+			System.out.println(str);
+		} else if (level.equals("debug") && Config.debug) {
+			System.out.println(str);
+		} else if (level.equals("notification") && Config.notification) {
+			System.out.println(str);
 		}
-		else {
-			if(level.equals("error") && Config.errors) {
-				gui.appendText(str + "\n");
-			} else if(level.equals("warning") && Config.warnings) {
-				gui.appendText(str + "\n");
-			} else if (level.equals("debug") && Config.debug) {
-				gui.appendText(str + "\n");
-			} else if (level.equals("notification") && Config.notification) {
-				gui.appendText(str + "\n");
-			}
-		}
-		
-	}
-
-	public static void addGUI(PlannerGUI plannerGui) {
-		gui = plannerGui;
 	}
 }

### 2015-01-28 src/thobi/LPTreatment.java
added LP Präse

@@ -443,9 +443,9 @@ public class LPTreatment {
 		{
 			seedUsed[i] = true;
 		}
-		solveLPMax();
-		solveLPMin();
 		
+		solveLPMin();
+		//solveLPMax();
 	}
 
 }

### 2015-01-28 src/utils/Config.java

@@ -30,7 +30,7 @@ public class Config {
 	public static final double pancreasGoalDose = 0; 	//max
 	public static final double tumorGoalDose = 50; 		//min
 	
-	public static final double relaxDose = 5;
+	public static final double relaxDose = 10;
 	
 	public static final double normalMinDose = 0;
 	public static final double spineMinDose = 0;

### 2015-01-28 src/erik/Population.java
parameter adjustment

@@ -11,7 +11,7 @@ import utils.LogTool;
 public class Population
 {
     final static int ELITISM_K =2;
-    final static int POP_SIZE = 10 + ELITISM_K;  // population size
+    final static int POP_SIZE = 8 + ELITISM_K;  // population size
     final static int MAX_ITER = 800;             // max number of iterations
     final static double MUTATION_RATE = 0.55;     // probability of mutation
     final static double CROSSOVER_RATE = 0.8;     // probability of crossover

### 2015-01-28 src/erik/TreatmentPlanner.java

@@ -133,10 +133,10 @@ public class TreatmentPlanner {
 		System.out.println("Runtime: " + dateFormatted);
 
 		
-		ScatterDisplay display5 = new ScatterDisplay(ChartType.BodyType);
+		/*ScatterDisplay display5 = new ScatterDisplay(ChartType.BodyType);
 		display5.fill(Solver.body, entry.getDimensions()[0], entry.getDimensions()[1],
 				entry.getDimensions()[2]);
-		display5.display();
+		display5.display();/*
 
 		
 		// System.exit(0);

### 2015-01-29 src/sebastian/TreatmentEntry.java
fixed things

@@ -146,12 +146,12 @@ public class TreatmentEntry implements Serializable {
 	}
 	res = res / sizeErrors.length / centerErrors.length / centerDistsErrors.length / closestDistsErrors.length;
 	
-	LogTool.print("Entry " + tEntry.getName(), "debug");
+	/*LogTool.print("Entry " + tEntry.getName(), "debug");
 	LogTool.print(String.format("Size: (%f|%f|%f|%f|%f|%f)", sizeErrors[0],sizeErrors[1],sizeErrors[2],sizeErrors[3],sizeErrors[4],sizeErrors[5]), "debug");
 	LogTool.print(String.format("Centers: (%f|%f|%f|%f|%f|%f)", centerErrors[0], centerErrors[1],centerErrors[2],centerErrors[3],centerErrors[4],centerErrors[5]), "debug");
 	LogTool.print(String.format("CenterDists: (%f|%f|%f|%f|%f|%f)", centerDistsErrors[0], centerDistsErrors[1],centerDistsErrors[2],centerDistsErrors[3],centerDistsErrors[4],centerDistsErrors[5]), "debug");
 	LogTool.print(String.format("ClosestDists: (%f|%f|%f|%f|%f|%f)", closestDistsErrors[0], closestDistsErrors[1],closestDistsErrors[2],closestDistsErrors[3],closestDistsErrors[4],closestDistsErrors[5]), "debug");
-	LogTool.print (String.format ("Error: %f\n", res), "debug");
+	LogTool.print (String.format ("Error: %f\n", res), "debug");*/
 	return res;
   }
 }

### 2015-01-29 src/utils/Voxel.java
Adjustments in LP stuff by Thobi

@@ -258,7 +258,11 @@ public class Voxel implements Serializable {
 	}
 	
 	public double getRelaxedGoalDosis() {
-		if(goalDosis == 0)
+		if(goalDosis == 50)
+		{
+			return goalDosis - Config.relaxDose;
+		}		
+		else if(goalDosis == 0)
 		{
 			return Config.relaxDose;
 		}

### 2015-01-29 src/dawid/Looper.java
Final SA local implementation without Multithreading

@@ -13,6 +13,7 @@ import utils.RandGenerator;
 import utils.Seed;
 import utils.Voxel;
 import java.lang.Math;
+import java.util.Arrays;
 import java.util.concurrent.CompletionService;
 import java.util.concurrent.ExecutionException;
 import java.util.concurrent.ExecutorCompletionService;
@@ -37,6 +38,8 @@ public class Looper {
     
     private  Voxel [][][] body; //Thobi hat das als ganz einfache Variable in seiner Loesermethode...nicht so OOP
     private  Seed[] seeds = new Seed[Config.SAnumberOfSeeds];
+    private  Voxel [][][] body2; //Thobi hat das als ganz einfache Variable in seiner Loesermethode...nicht so OOP
+    private  Seed[] seeds2 = new Seed[Config.SAnumberOfSeeds];
     public double[] Cur_state = new double[Config.SAnumberOfSeeds];
     public double[] New_state = new double[Config.SAnumberOfSeeds]; // Do I even need this ?
     public double[] Global_Lowest_state = new double[Config.SAnumberOfSeeds]; // Do I even need this ?
@@ -56,13 +59,15 @@ public class Looper {
         this.temperature = Config.StartTemp;
 		this.body = body;
 		this.seeds = seeds;
-                this.Cur_cost = Cur_cost;
-                this.New_cost = New_cost;
-                this.temperature = temperature;
-                this.Cur_state = Cur_state;
-                this.New_state = New_state;
+                this.body2 = body;
+		this.seeds2 = seeds;
+//                this.Cur_cost = Cur_cost;
+//                this.New_cost = New_cost;
+//                this.temperature = temperature;
+//                this.Cur_state = Cur_state;
+//                this.New_state = New_state;
                 this.Global_lowest_cost = Double.MAX_VALUE;
-                this.Global_Lowest_state = Global_Lowest_state;
+//                this.Global_Lowest_state = Global_Lowest_state;
 //                LogTool.print("Cur_state : " + Cur_state + " Zeit 2 : " + Zeit2 + " body : " + New_state, "notification");
 //                this.dwelltimes = Looper.seeds().;
                         }
@@ -91,6 +96,23 @@ public class Looper {
     * @param New_state Contains randomized dwelltimes
     * @param Config.SAnumberOfSeeds Contains randomized dwelltimes
     */
+    
+    /**
+     * Works.Randomly determines a new state vector
[…]

### 2015-01-29 src/dawid/TreatmentSA.java

@@ -157,7 +157,11 @@ public class TreatmentSA {
             }
         }
         
-        LogTool.print("     Seedtimes : A)" + seeds[0].getDurationMilliSec() + " B) " + seeds[1].getDurationMilliSec() + " C) " + seeds[2].getDurationMilliSec(),"notification");
+        String Global_Lowest_state_string = new String();
+        for (int cc = 0; cc < Config.SAnumberOfSeeds; cc++) {
+            Global_Lowest_state_string = Global_Lowest_state_string.concat(" " + cc + ") " + seeds[cc].getDurationMilliSec());
+            }
+        LogTool.print("     Seedtimes : " + Global_Lowest_state_string,"notification");
         
     /* Attack the problem using SA */
         
@@ -165,8 +169,9 @@ public class TreatmentSA {
         // Dawid inslucde the verbose switch pls
         LogTool.print("Initialized Looper Object!","notification");
         LogTool.print("Beginning Annealing...","notification");
+        looper.solveSA();
 //        LogTool.print("seeds, curstate,newstate" + looper.getCur_state() + " " + looper.getNew_state() + " ","notification");    
-       // GlobalState GLS = looper.solveSA();
+//        GlobalState GLS = looper.solveSA();
 //        GlobalState GLS = looper;
         LogTool.print("GLC: " + looper.getGlobal_lowest_cost()+ " CURC: " + looper.getCur_cost(),"notification");
 //        LogTool.print("FitnessGlobalLow: " + looper.+ " FitnessCURC: " + looper.getcurfitnessValue(),"notification");

### 2015-01-29 src/utils/Config.java

@@ -37,14 +37,14 @@ public class Config {
         // Parameters for the Simulated Annealing algorithm
         
         public static final boolean SAResets = false;
-        public static final int NumberOfMetropolisResets = 100;
-        public static final int NumberOfMetropolisRounds = 1000;
+        public static final int NumberOfMetropolisResets = 10;
+        public static final int NumberOfMetropolisRounds = 100;
         public static final double StartTemp = NumberOfMetropolisRounds;
         public static final int SAnumberOfSeeds = 50;
-        public static final boolean SAdebug = true;
-        public static final boolean SAdebugkoords = true;
-        public static final boolean SAdebugbody = true;
-        public static final int SAverboselvl = 2;
+        public static final boolean SAdebug = false;
+        public static final boolean SAdebugkoords = false;
+        public static final boolean SAdebugbody = false;
+        public static final int SAverboselvl = 0;
         
 	// Parameters for the genetic algorithm
 	

### 2015-01-29 src/dawid/Looper.java
fina final

@@ -1,8 +1,3 @@
-/*
- * To change this license header, choose License Headers in Project Properties.
- * To change this template file, choose Tools | Templates
- * and open the template in the editor.
- */
 package dawid;
 
 import erik.DoseEvaluator;
@@ -13,6 +8,7 @@ import utils.RandGenerator;
 import utils.Seed;
 import utils.Voxel;
 import java.lang.Math;
+import java.util.Arrays;
 import java.util.concurrent.CompletionService;
 import java.util.concurrent.ExecutionException;
 import java.util.concurrent.ExecutorCompletionService;
@@ -28,7 +24,7 @@ public class Looper {
     /**
  * Creates 
  *  <strong> Using the default constructor will not work.</strong>
- * Use <i>new Looper(body,seeds);	</i>
+ * Use <i>new Looper(body,seeds);   </i>
  * @param  seeds Seed[][][] Necessary for constructor
  * @param  body Voxel[][][]Necessary for constructor
  * @return      the image at the specified URL
@@ -37,6 +33,8 @@ public class Looper {
     
     private  Voxel [][][] body; //Thobi hat das als ganz einfache Variable in seiner Loesermethode...nicht so OOP
     private  Seed[] seeds = new Seed[Config.SAnumberOfSeeds];
+    private  Voxel [][][] body2; //Thobi hat das als ganz einfache Variable in seiner Loesermethode...nicht so OOP
+    private  Seed[] seeds2 = new Seed[Config.SAnumberOfSeeds];
     public double[] Cur_state = new double[Config.SAnumberOfSeeds];
     public double[] New_state = new double[Config.SAnumberOfSeeds]; // Do I even need this ?
     public double[] Global_Lowest_state = new double[Config.SAnumberOfSeeds]; // Do I even need this ?
@@ -45,24 +43,26 @@ public class Looper {
     private double Cur_cost, New_cost, Global_lowest_cost, temperature, newfitnessValue, curfitnessValue;
     
     /**
-	 * WHAT THIS DOES
-	 * @param body
-	 * The body of the patient
-	 * @param seeds
-	 * The seeds which should be optimized
-	 */
-	
+     * WHAT THIS DOES
+     * @param body
+     * The body of the patient
+     * @param seeds
+     * The seeds which should be optimized
+     */
+    
     public Looper(Voxel [][][] body, Seed[] seeds) {
         this.temperature = Config.StartTemp;
-		this.body = body;
-		this.seeds = seeds;
-                this.Cur_cost = Cur_cost;
[…]

### 2015-01-29 src/erik/Population.java
Finished presentation

@@ -170,9 +170,9 @@ public class Population
             System.out.println(" ; Best Fitness = " + pop.findBestIndividual().getFitnessValue()); 
            
             //LogTool.print("Iteration : "+iter, "notification");
-            //LogTool.print("Improvement: "+Math.abs(bestResultOld-bestResult)+" Cancel counter= " + counter, "debug");
+            LogTool.print("Improvement: "+Math.abs(bestResultOld-bestResult)+" Cancel counter= " + counter, "debug");
             
-            if(pop.findBestIndividual().getFitnessValue() < bestResult) {
+            if(pop.findBestIndividual().getFitnessValue() <= bestResult) {
             	bestResultOld = bestResult;
             	bestResult =  pop.findBestIndividual().getFitnessValue();
             }

### 2015-01-29 src/laurin/TreatmentAnalyzer.java
Error fix

@@ -133,7 +133,7 @@ public class TreatmentAnalyzer {
 				for(int z = 0; z < this.dimensions[2]; z++) {
 					double dose = 0.0;
 					for (Seed seed : seeds)
-					{;
+					{
 						dose += seed.radiationIntensity(body[x][y][z].distanceToVoxel(seed.getCoordinate()), seed.getDurationMilliSec());
 					}
 					body[x][y][z].setCurrentDosis(dose);									
@@ -141,6 +141,8 @@ public class TreatmentAnalyzer {
 			}	
 		}
 	}
+	
+	
 	
 	/**
 	 * Calculates average dose for specified body part
@@ -192,7 +194,9 @@ public class TreatmentAnalyzer {
 		for (Voxel voxel : anatomies.get(bodyType-1))
 		{
 			if (voxel.getCurrentDosis() > maxDose)
+			{
 				maxDose = voxel.getCurrentDosis();
+			}
 		}
 		return maxDose;
 	}
@@ -327,7 +331,7 @@ public class TreatmentAnalyzer {
 	 * 
 	 * @param treatmentAnalyzers
 	 */
-	public static void printTreatmentComparison(ArrayList<TreatmentAnalyzer> treatmentAnalyzers)
+	public static void printTreatmentComparison(ArrayList<TreatmentAnalyzer> treatmentAnalyzers, boolean showHistograms)
 	{
 		String spacing = String.format("%5s","");
 		String headLine = String.format("%-15s%s", "Treatment", spacing);
@@ -359,7 +363,7 @@ public class TreatmentAnalyzer {
 			for(int i = 0; i < Config.tumorType; i++)
 			{
 				minLines[i] += String.format("%15s%s", decimalFormat.format(ta.getMinDoses()[i]), spacing);
-				maxLines[i] += String.format("%15s%s", decimalFormat.format(ta.getMinDoses()[i]), spacing); 
+				maxLines[i] += String.format("%15s%s", decimalFormat.format(ta.getMaxDoses()[i]), spacing); 
 				avgLines[i] += String.format("%15s%s", decimalFormat.format(ta.getAvgDoses()[i]), spacing);			
 			}
 			conformalityLine += String.format("%15s%s", decimalFormat.format(ta.getConformalityIndex()), spacing);
@@ -390,6 +394,17 @@ public class TreatmentAnalyzer {
 		System.out.println(homogeinityLine);
[…]

### 2015-01-29 src/utils/Coordinate.java
Fixed bug in point on line calculation

@@ -47,11 +47,14 @@ public class Coordinate implements Serializable {
 	 * 
 	 * @return	point on line
 	 */
-	public static Coordinate getPointOnLine(Coordinate base, Coordinate other, double t)
+	public static Coordinate getPointOnLine(Coordinate base, Vector3D vDirection, double t)
 	{
+		if ((vDirection.getX() == 0) && (vDirection.getY() == 0) && (vDirection.getZ() == 0))
+			return base;
+		
 		Vector3D vBase = base.ToVector();
-		Vector3D vOther = other.ToVector();
-		Vector3D vPoint = vBase.add(vOther.subtract(vBase).scalarMultiply(t));
+		Vector3D vDirectionNormalized = vDirection.normalize();
+		Vector3D vPoint = vBase.add(vDirectionNormalized.scalarMultiply(t));
 		return vectorToCoordinate(vPoint);
 	}
 	

### 2015-01-29 src/utils/Config.java
Added LUT for dose approximation

@@ -99,6 +99,9 @@ public class Config {
 	public static final int scaleFactor = 5;
 	
 	public static final double cancelValue = 1;
+
+	//other
+	public static final int LUTSize = 10000; // size of lookup table for fast dose approximation
 	
 	
 	

### 2015-01-29 src/utils/Voxel.java

@@ -11,12 +11,30 @@ public class Voxel implements Serializable {
 	private double currentDosis;
 	private Coordinate coordinate;
 	private int bodyType = -1;
+	private static double[] lut;
 	
 	
 	public Voxel(double x, double y, double z) {
 		
 		coordinate = new Coordinate(x, y, z);
 		
+	}
+	
+	/**
+	 * LUT creation for fast dose approximation
+	 * @param maxDist
+	 * @param num
+	 */
+	public static void setLUT(double maxDist, int num)
+	{
+		lut = new double[num+1];
+		double phi = 90.0;
+		
+		for (int i = 0; i <= num; i++)
+		{
+			double distance = i*(maxDist/(double)(num+1));
+			lut[i] = ((1.12 * GL(1,90)/100) * Config.SK * (GL(distance, phi)/GL(1,90)) * gl(distance) * F(phi,distance) );  
+		}
 	}
 	
 	/**
@@ -67,6 +85,26 @@ public class Voxel implements Serializable {
 		return dose;
 	}
 	
+	public double radiationIntensityLUT(Coordinate position, double durationMilliSec, double phi){
+		double distance = distanceToVoxel(position);
+		
+		double dose= 0;
+		
+		if(distance > 10) {
+			return 0.0;
+		}
+
+		//  Interpolation for dose function 
+		
+		if(distance == 0) {
+			dose = this.getGoalDosis();
+		} else {
+			dose = lut[(int)(distance/10.0*Config.LUTSize)] * durationMilliSec;
+		}
+		
+		return dose;
+	}
+	
 	public double radiationIntensity(Coordinate position, double durationMilliSec){
 		return radiationIntensity(position, durationMilliSec, 90);
 	}
@@ -107,7 +145,7 @@ public class Voxel implements Serializable {
 	/*  Mathematical help functions to calculate all partial resultes for the dose */ 
 	
 	
-	private double F(double phi, double r) {
+	private static double F(double phi, double r) {
 		
 		
 		if(phi == 90) {
@@ -146,7 +184,7 @@ public class Voxel implements Serializable {
 		return F;
 	}
 	
-	private double gl(double r) {
+	private static double gl(double r) {
 		double gl = 0;
 		
 		double[] coeff = { 		//coefficents for the interpolationspolynom
@@ -171,7 +209,7 @@ public class Voxel implements Serializable {
 		
 	}
 	
-	private double GL(double r, double phi) {
[…]

### 2015-01-29 src/utils/Needle.java
Changed seed and needle placing methods

@@ -15,10 +15,18 @@ public class Needle implements Serializable {
 
 	private static final long serialVersionUID = 1L;
 	private ArrayList<Seed> seeds;
+	private Voxel tumorSurfacePointA;
+	private Voxel tumorSurfacePointB;
 	
-	public Needle()
+	private Needle()
 	{
-		seeds = new ArrayList<Seed>();
+		this.seeds = new ArrayList<Seed>();
+	}
+	
+	public Needle(Voxel insideTumor, Voxel outsideTumor, Voxel[][][] body)
+	{
+		this();
+		setTumorSurfacePointsFromVoxels (insideTumor, outsideTumor, body);
 	}
 	
 	public ArrayList<Seed> getSeeds()
@@ -31,25 +39,120 @@ public class Needle implements Serializable {
 		seeds.add(seed);
 	}
 	
+	
 	public int getSize()
 	{
 		return this.seeds.size();
 	}
 	
-	/**
-	 * Calculates direction of needle as difference between second and first seeds' coordinates.
-	 * 
-	 * @return direction of needle
-	 */
-	public Vector3D getDirection()
+	public Voxel getTumorSurfacePointA() 
 	{
-		if (this.getSize() < 2)
-			return null;
-		else
+		return tumorSurfacePointA;
+	}
+	
+	public Voxel getTumorSurfacePointB() 
+	{
+		return tumorSurfacePointB;
+	}
+	
+	/**
+	 * Sorts seeds by their distance from the first intersection with the tumor surface
+	 */
+	public void sortSeedsByDistanceToTumorSurfacePointA() 
+	{
+		Collections.sort(this.seeds, new Comparator<Seed>() {
+		    public int compare(Seed s1, Seed s2) {    
+		    	Needle needle = s1.getNeedle();
+		    	Double s1Dist = new Double(s1.distanceToVoxel(needle.tumorSurfacePointA.getCoordinate()));
+		    	Double s2Dist = new Double(s1.distanceToVoxel(needle.tumorSurfacePointA.getCoordinate()));
+		        return s1Dist.compareTo(s2Dist);
+		    }
+		});
+	}
+	
+	/**
+	 * Calculates the needles' intersecting points with the tumor from one point inside the tumor and one point outside
+	 * @param insideVoxel
+	 * @param otherVoxel
+	 * @param body
+	 */
[…]

### 2015-01-29 src/laurin/CplexSolver.java
Added GUI for stepwise LP

@@ -5,6 +5,8 @@ import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.Set;
 
+import laurin.LPStepwiseGUI.DoseGoalType;
+
 import ilog.concert.IloException;
 import ilog.concert.IloLinearNumExpr;
 import ilog.concert.IloNumExpr;
@@ -18,42 +20,288 @@ import utils.Voxel;
 
 public class CplexSolver {
 	
-	private Voxel[][][] body;
+	public static enum SlackType {PER_VOI, PER_VOXEL};
+	public static enum GoalType {MAXIMIZE, MINIMIZE};
+	public static enum BoundType {LOWER_BOUND, UPPER_BOUND};
+	
+	private Voxel[][] voxels;
 	private Seed[] seeds;
-	private int xSize;
-	private int ySize;
-	private int zSize;
 	double[] dwellTimes;
 	
-	IloCplex cplex;
+	private IloCplex cplex;
 	
-	IloLinearNumExpr c_all_objective; // objective for dwell times
-	IloLinearNumExpr[] c_s_lower_voi_objectives; // obj. single lower slacks
-	IloLinearNumExpr[] c_s_upper_voi_objectives; // obj. single upper slacks
-	IloLinearNumExpr[] c_t_lower_voi_objectives; // obj. PTV lower slacks
-	IloLinearNumExpr[] c_t_upper_voi_objectives; // obj. PTV upper slacks
+	private IloLinearNumExpr c_all_objective; // objective for dwell times
+	private IloLinearNumExpr[] c_s_lower_voi_objectives; // obj. single lower slacks
+	private IloLinearNumExpr[] c_s_upper_voi_objectives; // obj. single upper slacks
+	private IloLinearNumExpr[] c_t_lower_voi_objectives; // obj. PTV lower slacks
+	private IloLinearNumExpr[] c_t_upper_voi_objectives; // obj. PTV upper slacks
 	
-	IloNumVar[] t; // dwell times
-	IloNumVar[] s_lower; // single lower slacks
-	IloNumVar[] s_upper; // single upper slack
-	IloNumVar[] t_lower; // lower slacks for body parts
-	IloNumVar[] t_upper; // upper slacks for body parts
+	private IloNumVar[] t; // dwell times
+	private IloNumVar[][] s_lower; // single lower slacks
+	private IloNumVar[][] s_upper; // single upper slack
[…]

### 2015-01-29 src/laurin/LPStepwise.java

@@ -1,0 +1,213 @@
+package laurin;
+
+import ilog.concert.IloException;
+
+import java.io.IOException;
+import java.util.AbstractMap;
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.Comparator;
+import java.util.HashSet;
+import java.util.Map.Entry;
+import java.util.Set;
+
+import sebastian.BodyEntry;
+import sebastian.SimpleDB;
+import utils.Config;
+import utils.Coordinate;
+import utils.Needle;
+import utils.RandGenerator;
+import utils.Seed;
+import utils.Voxel;
+import erik.BodyAnalyzer;
+
+public class LPStepwise {
+
+	public static Voxel[][][] getVoxelArray(BodyEntry entry) 
+	{
+		Voxel[][][] body = new Voxel[entry.getDimensions()[0]][entry.getDimensions()[1]][entry.getDimensions()[2]];
+		
+		for (int x = 0; x < entry.getDimensions()[0]; x++) 
+		{
+			for (int y = 0; y < entry.getDimensions()[1]; y++)
+			{
+				for (int z = 0; z < entry.getDimensions()[2]; z++)
+				{
+					body[x][y][z] = new Voxel(x, y, z);
+					
+					switch (entry.getBodyArray()[x][y][z].getBodyType()) {
+						case Config.normalType: {
+							body[x][y][z].setGoalDosis(Config.normalGoalDose);
+							body[x][y][z].setBodyType(Config.normalType);
+							break;
+						}
+						case Config.spineType: {
+							body[x][y][z].setGoalDosis(Config.spineGoalDose);
+							body[x][y][z].setBodyType(Config.spineType);
+							break;
+						}
+						case Config.liverType: {
+							body[x][y][z].setGoalDosis(Config.liverGoalDose);
+							body[x][y][z].setBodyType(Config.liverType);
+							break;
+						}
+						case Config.pancreasType: {
+							body[x][y][z].setGoalDosis(Config.pancreasGoalDose);
+							body[x][y][z].setBodyType(Config.pancreasType);
+							break;
+						}
+						case Config.tumorType: {
+							body[x][y][z].setGoalDosis(Config.tumorGoalDose);
+							body[x][y][z].setBodyType(Config.tumorType);
[…]

### 2015-01-29 src/laurin/LPStepwiseGUI.java

@@ -1,0 +1,376 @@
+package laurin;
+
+import ilog.concert.IloException;
+
+import java.awt.BorderLayout;
+import java.awt.Checkbox;
+import java.awt.CheckboxGroup;
+import java.awt.EventQueue;
+import java.awt.FlowLayout;
+import java.awt.GridLayout;
+import java.awt.Panel;
+import java.awt.event.ActionEvent;
+import java.awt.event.ActionListener;
+import java.util.ArrayList;
+import java.util.Set;
+
+import javax.swing.BorderFactory;
+import javax.swing.BoxLayout;
+import javax.swing.JButton;
+import javax.swing.JCheckBox;
+import javax.swing.JComboBox;
+import javax.swing.JFrame;
+import javax.swing.JLabel;
+import javax.swing.JOptionPane;
+import javax.swing.JPanel;
+import javax.swing.JTextField;
+
+import javax.swing.border.EtchedBorder;
+import javax.swing.event.DocumentEvent;
+import javax.swing.event.DocumentListener;
+
+import laurin.CplexSolver.BoundType;
+
+import erik.Solver;
+
+import sebastian.BodyEntry;
+import utils.Config;
+import utils.Seed;
+import utils.Voxel;
+
+
+public class LPStepwiseGUI extends JFrame implements ActionListener{
+
+	public static enum DoseGoalType {GOAL_DOSE, MIN_DOSE, MAX_DOSE, MANUALLY};
+	public static enum BodyType {NORMAL, SPINE, LIVER, PANCREAS, TUMOR};
+	public static enum ActionType {OPTIMIZE};
+	
+	private CplexSolver cplexSolverSampled;
+	private Voxel[][][] body;
+	private int[] dimensions;
+	
+	private static final long serialVersionUID = 1L;
+	private JPanel panel = null;
+	private JComboBox<LPStepwiseGUI.DoseGoalType> doseGoalBox;
+	private JComboBox<CplexSolver.BoundType> boundBox;
+	private JComboBox<CplexSolver.SlackType> slackBox;
+	private Checkbox[] bodyTypeBoxes;
+	private CheckboxGroup bodyTypeBoxesGroup;
+	private JTextField[] lowerBoundsFields;
+	private JTextField[] upperBoundsFields;
+	private JTextField manualDoseField;
+
[…]

### 2015-01-29 src/dawid/Looper.java
Optimaleres SA fuer Sebastian

@@ -54,7 +54,7 @@ public class Looper {
         this.temperature = Config.StartTemp;
         this.body = body;
         this.seeds = seeds;
-                this.body2 = body;
+        this.body2 = body;
         this.seeds2 = seeds;
 //                this.Cur_cost = Cur_cost;
 //                this.New_cost = New_cost;
@@ -74,6 +74,8 @@ public class Looper {
  * @return      Writes to Cur_state, doesnt return anything
  */
     private void initState() {
+        double metaintensity=0;
+        double diffr=0;
          for(int ii=0; ii < Config.SAnumberOfSeeds; ii++)
             {
                 Cur_state[ii] = this.seeds[ii].getDurationMilliSec();
@@ -85,6 +87,32 @@ public class Looper {
 //                Cur_state[1] = Zeit2;
 //                Cur_state[2] = Zeit3;
             }
+        for(int x=0; x < Config.xDIM; x++) {
+//        for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x++) {
+            for(int y=0; y < Config.yDIM; y++) {
+//            for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y++) {
+                for(int z=0; z < Config.zDIM; z++) {
+//                for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z++) {
+
+//                    this.body2[x][y][z].setCurrentDosis(0.0);  //Set currentPtvVoxel Dose to 0 
+                    this.body2[x][y][z].metavalue = 0.0;
+                        for(int i=0; i<Config.SAnumberOfSeeds;++i) { 
+                            // Calculate intensity based based on current dwelltime
+                            metaintensity = this.body2[x][y][z].radiationIntensityNoTime((this.seeds2[i].getCoordinate()));
+    //                                radiationIntensityNoTime(this.seeds2[i].getCoordinate(), New_state[i]);
+                                                    if (metaintensity>0) {
+    //                                                LogTool.print("Cost: Intensity :" + intensity + "@ " + x + " " + y + " " + z,"notification");
+                                                    }
+    //                        this.body2[x][y][z].addCurrentDosis(metaintensity);
+                              this.body2[x][y][z].metavalue += metaintensity;                      
+                        }   
+                }    
+            }
+        }
+        this.body = this.body2;
[…]

### 2015-01-29 src/dawid/TreatmentSA.java

@@ -18,7 +18,7 @@ import utils.Voxel;
  */
 public class TreatmentSA {
 
-        public static int[] dimensions;
+    public static int[] dimensions;
     static public int[] xBoundsTumor;
     static public int[] yBoundsTumor;
     static public int[] zBoundsTumor;
@@ -116,10 +116,12 @@ public class TreatmentSA {
                     if( (x >= Config.ptvXLow && x <= Config.ptvXHigh) && (y >= Config.ptvYLow && y <= Config.ptvYHigh) && (z >= Config.ptvZLow && z <= Config.ptvZHigh)  ) {        //consider whether the current voxel belongs to the tumor or not
                         body[x][y][z].setGoalDosis(Config.ptvGoalDose);
                         body[x][y][z].setBodyType(Config.tumorType);
+                        body[x][y][z].metavalue = 0;
                         ptvSize++;
                     } else {
                         body[x][y][z].setGoalDosis(Config.oarGoalDose);
                         body[x][y][z].setBodyType(Config.normalType);
+                        body[x][y][z].metavalue = 0;
                         oarSize++;
                     }
                 }
@@ -129,9 +131,9 @@ public class TreatmentSA {
         
     /* Verify body */     
         if (Config.SAdebugbody) {
-        LogTool.print("     MaxDose 3,0,1 : " + body[53][50][50].getMaxDosis() + "","notification");
+        LogTool.print("     MaxDose 46,46,46 : " + body[46][46][46].getMaxDosis() + "","notification");
         LogTool.print("     MaxDose 1,1,1 : " + body[5][5][5].getMaxDosis() + "","notification");
-        LogTool.print("     Type 3,0,1 : " + body[3][0][1].getBodyType()+ "","notification");
+        LogTool.print("     Type 46,46,46 : " + body[46][46][46].getBodyType()+ "","notification");
         LogTool.print("     Type 1,1,1 : " + body[5][5][5].getBodyType() + "","notification");
         }
                 

### 2015-01-29 src/utils/Config.java

@@ -54,8 +54,8 @@ public class Config {
         // Parameters for the Simulated Annealing algorithm
         
         public static final boolean SAResets = false;
-        public static final int NumberOfMetropolisResets = 100;
-        public static final int NumberOfMetropolisRounds = 1000;
+        public static final int NumberOfMetropolisResets = 5;
+        public static final int NumberOfMetropolisRounds = 60;
         public static final double StartTemp = NumberOfMetropolisRounds;
         public static final int SAnumberOfSeeds = 50;
         public static final boolean SAdebug = true;

### 2015-01-29 src/utils/Voxel.java

@@ -11,12 +11,13 @@ public class Voxel implements Serializable {
 	private double currentDosis;
 	private Coordinate coordinate;
 	private int bodyType = -1;
+        public double metavalue;
 	
 	
 	public Voxel(double x, double y, double z) {
 		
 		coordinate = new Coordinate(x, y, z);
-		
+                
 	}
 	
 	/**
@@ -70,6 +71,10 @@ public class Voxel implements Serializable {
 	public double radiationIntensity(Coordinate position, double durationMilliSec){
 		return radiationIntensity(position, durationMilliSec, 90);
 	}
+        
+        public double radiationIntensityNoTime(Coordinate position){
+		return radiationIntensityNoTimeCALC(position, 90);
+	}
 	
 	public double radiationIntensity(double r, double durationMilliSec){
 		double distance = r;
@@ -92,6 +97,26 @@ public class Voxel implements Serializable {
 		return dose;
 	}
 	
+        public double radiationIntensityNoTimeCALC(Coordinate position, double phi){
+		double distance = distanceToVoxel(position);
+		
+		double dose= 0;
+		
+		if(distance > 10) {
+			return 0.0;
+		}
+
+		//  Interpolation for dose function 
+		
+		if(distance == 0) {
+			dose = this.getGoalDosis();
+		} else {
+			dose = ((1.12 * GL(1,90)/100) * Config.SK * (GL(distance, phi)/GL(1,90)) * gl(distance) * F(phi,distance) ); 
+		}
+		
+		return dose;
+	}
+        
 	public static void main(String args[]) {
 		Voxel v = new Voxel(0, 0, 0);
 		

### 2015-01-29 src/utils/Voxel.java
Namen angepasst, javadoc

@@ -136,6 +136,13 @@ public class Voxel implements Serializable {
 		return dose;
 	}
 	
+        /**
+         * @author Dawid
+         * @param position
+         * @param phi
+         * @return metavalue
+         */
+        
         public double radiationIntensityNoTimeCALC(Coordinate position, double phi){
 		double distance = distanceToVoxel(position);
 		

### 2015-01-29 src/dawid/Looper.java
setFinalSolution fuer GUI

@@ -87,11 +87,11 @@ public class Looper {
 //                Cur_state[1] = Zeit2;
 //                Cur_state[2] = Zeit3;
             }
-        for(int x=0; x < Config.xDIM; x++) {
+        for(int x=Config.ptvXLow-0; x < Config.ptvXHigh+0; x++) {
 //        for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x++) {
-            for(int y=0; y < Config.yDIM; y++) {
+            for(int y=Config.ptvYLow-0; y < Config.ptvYHigh+0; y++) {
 //            for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y++) {
-                for(int z=0; z < Config.zDIM; z++) {
+                for(int z=Config.ptvZLow-0; z < Config.ptvZHigh+0; z++) {
 //                for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z++) {
 
 //                    this.body2[x][y][z].setCurrentDosis(0.0);  //Set currentPtvVoxel Dose to 0 
@@ -556,6 +556,16 @@ public class Looper {
 
     public double getcurfitnessValue() {
         return curfitnessValue;
+    }
+    
+    public void setFinalSolution () {
+        for(int i=0; i<Config.SAnumberOfSeeds;i++) {
+                this.seeds[i].setDurationMilliSec(GLowestState.getDwelltimes()[i]);
+                Cur_state[i] = GLowestState.getDwelltimes()[i];
+            }
+        setCur_cost(costCURsuper());
+//        setCur_cost(costCUR());
+        
     }
     
     

### 2015-01-29 src/dawid/TreatmentSA.java

@@ -179,6 +179,13 @@ public class TreatmentSA {
 //        LogTool.print("FitnessGlobalLow: " + looper.+ " FitnessCURC: " + looper.getcurfitnessValue(),"notification");
         //LogTool.print("GLS external: " + GLS.getGlobal_Lowest_state_string(),"notification");
         LogTool.print("SolveSA: Global Current Best Solution : " + looper.getGlobal_Lowest_state_string(),"notification");
+        
+//        for(int i=0; i<Config.SAnumberOfSeeds;i++) {
+//            double theval = looper.getGlobal_Lowest_state()[i];
+//			Solver.seeds[i].setDurationMilliSec(theval);
+//			System.out.println("Verweildauer: "+ Solver.seeds[i].getDurationMilliSec() );
+//		}
+        
 //        looper.rausfindenWarumCUrCostsichnichtaendert();
 
 //        The looper needs to establish an initial solution and temperature

### 2015-01-29 src/erik/TreatmentPlanner.java

@@ -118,7 +118,7 @@ public class TreatmentPlanner {
 
 		long start = System.currentTimeMillis();
 		//solver.solveGeneticAlg();
-        //solver.solveSA();
+                //solver.solveSA();
 		solver.solveLP();
 		long end = System.currentTimeMillis();
 		TreatmentAnalyzer ta = new TreatmentAnalyzer(Solver.body,

### 2015-01-29 src/utils/Voxel.java

@@ -108,11 +108,20 @@ public class Voxel implements Serializable {
 	}
 	
 	public double radiationIntensity(Coordinate position, double durationMilliSec){
-		return radiationIntensity(position, durationMilliSec, 90);
+            double temp=radiationIntensityNoTimeCALC(position, 90);
+             if (temp>0.0) {
+//                LogTool.print("Wert in Time " + temp, "notification");
+//                LogTool.print("Wert in Time " + temp-, "notification");
+            }
+		return temp;
 	}
         
         public double radiationIntensityNoTime(Coordinate position){
-		return radiationIntensityNoTimeCALC(position, 90);
+            double temp = radiationIntensityNoTimeCALC(position, 90);
+            if (temp>0.0) {
+                LogTool.print("Wert in nOtIME " + temp, "notification");
+            }
+		return temp;
 	}
 	
 	public double radiationIntensity(double r, double durationMilliSec){

### 2015-01-30 src/dawid/Looper.java
Aufrufen von SA fuer die GUI 1) in der Config unter SA den CostFunctionType setzen (falls Werte schlecht oder etwas nicht geht, 3 ist failsafe) 2) die methode erik.Solver.solveSA() setzt, wenn sie ausgefuehrt wird, Solver.body auf den optimalen Body und Solver.seeds auf die optimalen seeds.

@@ -87,15 +87,16 @@ public class Looper {
 //                Cur_state[1] = Zeit2;
 //                Cur_state[2] = Zeit3;
             }
-        for(int x=Config.ptvXLow-0; x < Config.ptvXHigh+0; x++) {
-//        for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x++) {
-            for(int y=Config.ptvYLow-0; y < Config.ptvYHigh+0; y++) {
-//            for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y++) {
-                for(int z=Config.ptvZLow-0; z < Config.ptvZHigh+0; z++) {
-//                for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z++) {
-
-//                    this.body2[x][y][z].setCurrentDosis(0.0);  //Set currentPtvVoxel Dose to 0 
-                    this.body2[x][y][z].metavalue = 0.0;
+        
+         if ((Config.SACostFunctionType==3)||(Config.SACostFunctionType==1)) {
+            for(int x=Config.ptvXLow-0; x < Config.ptvXHigh+0; x++) {
+//        for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x+= Config.scaleFactor) {
+                for(int y=Config.ptvYLow-0; y < Config.ptvYHigh+0; y++) {
+//                  for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y+= Config.scaleFactor) {
+                    for(int z=Config.ptvZLow-0; z < Config.ptvZHigh+0; z++) {
+//                      for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z+= Config.scaleFactor) {
+//                          this.body2[x][y][z].setCurrentDosis(0.0);  //Set currentPtvVoxel Dose to 0 
+                        this.body2[x][y][z].metavalue = 0.0;
                         for(int i=0; i<Config.SAnumberOfSeeds;++i) { 
                             // Calculate intensity based based on current dwelltime
                             metaintensity = this.body2[x][y][z].radiationIntensityNoTime((this.seeds2[i].getCoordinate()));
@@ -104,15 +105,41 @@ public class Looper {
     //                                                LogTool.print("Cost: Intensity :" + intensity + "@ " + x + " " + y + " " + z,"notification");
                                                     }
[…]

### 2015-01-30 src/erik/Solver.java

@@ -104,21 +104,19 @@ public class Solver {
         
         public void solveSA() {
             Looper looper = new Looper(Solver.body,Solver.seeds);
-            looper.solveSA();
             LogTool.print("Initialized Looper Object!","notification");
-        LogTool.print("Beginning Annealing...","notification");
-//        LogTool.print("seeds, curstate,newstate" + looper.getCur_state() + " " + looper.getNew_state() + " ","notification");    
-//        GlobalState GLS = looper.solveSA();
-//        GlobalState GLS = looper;
-        LogTool.print("GLC: " + looper.getGlobal_lowest_cost()+ " CURC: " + looper.getCur_cost(),"notification");
-//        LogTool.print("FitnessGlobalLow: " + looper.+ " FitnessCURC: " + looper.getcurfitnessValue(),"notification");
-//        LogTool.print("GLS external: " + GLS.getGlobal_Lowest_state_string(),"notification");
-        LogTool.print("SolveSA: Global Current Best Solution : " + looper.getGlobal_Lowest_state_string(),"notification");
+            LogTool.print("Beginning Annealing...","notification");
+            looper.solveSA();
+            LogTool.print("GLC: " + looper.getGlobal_lowest_cost()+ " CURC: " + looper.getCur_cost(),"notification");
+            LogTool.print("SolveSA: Global Current Best Solution : " + looper.getGlobal_Lowest_state_string(),"notification");
         
-        for(int i=0; i<Config.SAnumberOfSeeds;i++) {
-			Solver.seeds[i].setDurationMilliSec(looper.getGlobal_Lowest_state()[i]);
-			System.out.println("Verweildauer: "+ Solver.seeds[i].getDurationMilliSec() );
-		}
+            looper.setFinalSolution();
+            Solver.body = looper.getBody();
+            Solver.seeds = looper.getSeeds();
+            for(int i=0; i<Config.SAnumberOfSeeds;i++) {
+//                            Solver.seeds[i].setDurationMilliSec(looper.getGlobal_Lowest_state()[i]);
+                            System.out.println("Verweildauer: "+ Solver.seeds[i].getDurationMilliSec() );
+                    }
         }
 	
 	public void solveLP() throws IloException

### 2015-01-30 src/utils/Config.java

@@ -53,6 +53,14 @@ public class Config {
 	
         // Parameters for the Simulated Annealing algorithm
         
+        // SAostFunctionType legend
+        // 0: fast using the Solver.TumorBounds
+        // 1: fast using ptv static bounds above
+        // 2: slow using the Solver.TumorBounds
+        // 3: slow using the static bounds above
+        // in case of errors keep changing, option 3 is SAFE
+        
+        public static final int SACostFunctionType = 0; //see above comments 
         public static final boolean SAResets = false;
         public static final int NumberOfMetropolisResets = 5;
         public static final int NumberOfMetropolisRounds = 60;

### 2015-01-30 src/laurin/LPMatlabStepwise.java
Hotfix dose calculation

@@ -1,167 +1,0 @@
-package laurin;
-
-import java.io.IOException;
-import java.util.ArrayList;
-import java.util.HashSet;
-import java.util.Set;
-
-import ilog.concert.IloException;
-import ilog.concert.IloLinearNumExpr;
-import ilog.concert.IloNumExpr;
-import ilog.concert.IloNumVar;
-import ilog.cplex.IloCplex;
-import sebastian.BodyEntry;
-import sebastian.SimpleDB;
-import sebastian.TreatmentEntry;
-import utils.Config;
-import utils.Coordinate;
-import utils.LogTool;
-import utils.RandGenerator;
-import utils.Seed;
-import utils.Voxel;
-import erik.BodyAnalyzer;
-
-public class LPMatlabStepwise {
-	
-	public static Voxel[][][] getVoxelArray(BodyEntry entry) 
-	{
-		Voxel[][][] body = new Voxel[entry.getDimensions()[0]][entry.getDimensions()[1]][entry.getDimensions()[2]];
-		
-		for (int x = 0; x < entry.getDimensions()[0]; x++) 
-		{
-			for (int y = 0; y < entry.getDimensions()[1]; y++)
-			{
-				for (int z = 0; z < entry.getDimensions()[2]; z++)
-				{
-					body[x][y][z] = new Voxel(x, y, z);
-					
-					switch (entry.getBodyArray()[x][y][z].getBodyType()) {
-						case Config.normalType: {
-							body[x][y][z].setGoalDosis(Config.normalGoalDose);
-							body[x][y][z].setBodyType(Config.normalType);
-							break;
-						}
-						case Config.spineType: {
-							body[x][y][z].setGoalDosis(Config.spineGoalDose);
-							body[x][y][z].setBodyType(Config.spineType);
-							break;
-						}
-						case Config.liverType: {
-							body[x][y][z].setGoalDosis(Config.liverGoalDose);
-							body[x][y][z].setBodyType(Config.liverType);
-							break;
-						}
-						case Config.pancreasType: {
-							body[x][y][z].setGoalDosis(Config.pancreasGoalDose);
-							body[x][y][z].setBodyType(Config.pancreasType);
-							break;
-						}
-						case Config.tumorType: {
-							body[x][y][z].setGoalDosis(Config.tumorGoalDose);
[…]

### 2015-01-30 src/utils/Voxel.java

@@ -113,8 +113,9 @@ public class Voxel implements Serializable {
 //                LogTool.print("Wert in Time " + temp, "notification");
 //                LogTool.print("Wert in Time " + temp-, "notification");
             }
-		return temp;
+		return temp*durationMilliSec;
 	}
+	
         
         public double radiationIntensityNoTime(Coordinate position){
             double temp = radiationIntensityNoTimeCALC(position, 90);

### 2015-03-02 src/laurin/ResultReader.java
Latest Version

@@ -1,15 +1,30 @@
 package laurin;
 
+import java.io.File;
+import java.util.ArrayList;
 public class ResultReader {
 
 	/**
 	 * @param args
 	 */
 	public static void main(String[] args) {
-		TreatmentAnalyzer treatmentAnalyzer = TreatmentAnalyzer.readFromFile("treatments/LP_50_5.0_.ser");
+		/*TreatmentAnalyzer treatmentAnalyzer = TreatmentAnalyzer.readFromFile("treatments/LP_50_5.0_.ser");
 		//TreatmentAnalyzer treatmentAnalyzer = TreatmentAnalyzer.readFromFile("treatments/GA_50_2.0_10.0_0.55_0.8_1.0_1.0_1.0_1.0_1.0_2.0_5.0_.ser");
 		treatmentAnalyzer.printResults();
-		System.out.println(treatmentAnalyzer.getSAlgorithm());
+		System.out.println(treatmentAnalyzer.getSAlgorithm());*/
+		
+		ArrayList<TreatmentAnalyzer> treatmentAnalyzers = new ArrayList<TreatmentAnalyzer>();
+		
+		final File folder = new File("treatments/");
+		
+	    for (final File fileEntry : folder.listFiles()) {
+	        if (fileEntry.getName().startsWith("LP")) {
+	        	String filename = "treatments/" + fileEntry.getName();
+	            treatmentAnalyzers.add(TreatmentAnalyzer.readFromFile(filename));
+	        }
+	    }
+	    
+	    TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, true);
 	}
 
 }

### 2015-03-02 src/laurin/ResultReader.java
Added further methods for treatment comparison

@@ -1,30 +1,82 @@
 package laurin;
 
 import java.io.File;
+import java.io.FileFilter;
 import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.Map;
+import java.util.regex.Pattern;
 public class ResultReader {
 
 	/**
 	 * @param args
 	 */
-	public static void main(String[] args) {
-		/*TreatmentAnalyzer treatmentAnalyzer = TreatmentAnalyzer.readFromFile("treatments/LP_50_5.0_.ser");
-		//TreatmentAnalyzer treatmentAnalyzer = TreatmentAnalyzer.readFromFile("treatments/GA_50_2.0_10.0_0.55_0.8_1.0_1.0_1.0_1.0_1.0_2.0_5.0_.ser");
-		treatmentAnalyzer.printResults();
-		System.out.println(treatmentAnalyzer.getSAlgorithm());*/
-		
+	
+	public static HashMap<TreatmentAnalyzer, String> getFiles(String folderName, String regex)
+	{
 		ArrayList<TreatmentAnalyzer> treatmentAnalyzers = new ArrayList<TreatmentAnalyzer>();
 		
-		final File folder = new File("treatments/");
+		HashMap<TreatmentAnalyzer, String> treatments = new HashMap<TreatmentAnalyzer, String>();
 		
-	    for (final File fileEntry : folder.listFiles()) {
-	        if (fileEntry.getName().startsWith("LP")) {
-	        	String filename = "treatments/" + fileEntry.getName();
-	            treatmentAnalyzers.add(TreatmentAnalyzer.readFromFile(filename));
-	        }
-	    }
+		final File folder = new File(folderName);		
+		File[] files = listFilesMatching(folder, regex);
+
+		
+		for (File file : files)
+		{
+			String filename = folderName + file.getName();
+            treatments.put(TreatmentAnalyzer.readFromFile(filename), file.getName());
+		}
 	    
-	    TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, true);
+	    return treatments;
+	}
+	
+	public static void compareHistogramsByBodyType(HashMap<TreatmentAnalyzer, String> treatments, String bodyType)
+	{
+		ArrayList<Histogram> histograms = new ArrayList<Histogram>();
+		ArrayList<HashMap<String, double[]>> dataSets = new ArrayList<HashMap<String, double[]>>();
+		
+		Histogram histogramComparison = new Histogram("Comparison of " + bodyType);
+		
[…]

### 2015-03-02 src/laurin/TreatmentAnalyzer.java

@@ -123,6 +123,11 @@ public class TreatmentAnalyzer implements Serializable {
 		return gridResolution;
 	}
 	
+	public Histogram getHistogram()
+	{
+		return histogram;
+	}
+	
 	public void setMinDoses(double[] minDoses) {
 		this.minDoses = minDoses;
 	}
@@ -178,6 +183,10 @@ public class TreatmentAnalyzer implements Serializable {
 		this.gridResolution = gridResolution;
 	}
 	
+	public void setTitle(String title)
+	{
+		this.title = title;
+	}
 	/**
 	 * Calculates dose for each voxel.
 	 */

### 2015-03-02 src/laurin/TreatmentAnalyzer.java
Fix of body type descriptions in histograms

@@ -367,9 +367,9 @@ public class TreatmentAnalyzer implements Serializable {
 		LogTool.print("Adding histogram data", "Notification");
 		Histogram histogram = new Histogram(this.title);
 		histogram.addDataSet("Normal", this.getAnatomy(Config.normalType));
-		histogram.addDataSet("Spine", this.getAnatomy(Config.bladderType));
-		histogram.addDataSet("Liver", this.getAnatomy(Config.rectumType));
-		histogram.addDataSet("Pancreas", this.getAnatomy(Config.urethraType));
+		histogram.addDataSet("Bladder", this.getAnatomy(Config.bladderType));
+		histogram.addDataSet("Rectum", this.getAnatomy(Config.rectumType));
+		histogram.addDataSet("Urethra", this.getAnatomy(Config.urethraType));
 		histogram.addDataSet("Tumor", this.getAnatomy(Config.tumorType));
 		setHistogram(histogram);
 	}

### 2015-03-02 src/erik/BodyAnalyzer.java
Fixed normalType issue

@@ -291,8 +291,11 @@ public class BodyAnalyzer {
 		
 		for(int x = 0; x < body.length; x++) {				
 			for(int y = 0; y < body[0].length; y++) {		
-				for(int z = 0; z < body[0][0].length; z++) {		
-					anatomies.get(body[x][y][z].getBodyType()-1).add(body[x][y][z]);									
+				for(int z = 0; z < body[0][0].length; z++) {
+					if ((body[x][y][z].getBodyType()) != Config.dummyType)
+					{
+						anatomies.get(body[x][y][z].getBodyType()-1).add(body[x][y][z]);	
+					}
 				}
 			}	
 		}	

### 2015-03-02 src/erik/Solver.java

@@ -65,8 +65,9 @@ public class Solver {
 							break;
 						}
 						default: {
-							Solver.body[x][y][z].setGoalDosis(Config.normalGoalDose);
-							Solver.body[x][y][z].setBodyType(Config.normalType);
+							//Solver.body[x][y][z].setGoalDosis(Config.normalGoalDose);
+							//Solver.body[x][y][z].setBodyType(Config.normalType);
+							Solver.body[x][y][z].setBodyType(Config.dummyType);
 							break;
 						}
 					

### 2015-03-02 src/erik/TreatmentPlanner.java

@@ -20,7 +20,7 @@ import utils.Seed;
 import utils.Voxel;
 
 public class TreatmentPlanner {
-	private static final boolean outputToFile = true;
+	private static final boolean outputToFile = false;
 
 	private static void planTreatment(String algo, double doubleArgs[]) throws IloException {
 
@@ -88,7 +88,7 @@ public class TreatmentPlanner {
 						case Config.normalType: {
 							//body[x][y][z].setGoalDosis(Config.normalGoalDose);
 							//body[x][y][z].setBodyType(Config.normalType);
-							body[x][y][z].setBodyType(7);
+							body[x][y][z].setBodyType(Config.dummyType);
 							break;
 						}
 						case Config.bladderType: {
@@ -114,7 +114,7 @@ public class TreatmentPlanner {
 						default: {
 							//body[x][y][z].setGoalDosis(Config.normalGoalDose);
 							//body[x][y][z].setBodyType(Config.normalType);
-							body[x][y][z].setBodyType(7);
+							body[x][y][z].setBodyType(Config.dummyType);
 							break;
 						}
 						}

### 2015-03-02 src/laurin/ResultReader.java

@@ -33,15 +33,13 @@ public class ResultReader {
 	
 	public static void compareHistogramsByBodyType(HashMap<TreatmentAnalyzer, String> treatments, String bodyType)
 	{
-		ArrayList<Histogram> histograms = new ArrayList<Histogram>();
-		ArrayList<HashMap<String, double[]>> dataSets = new ArrayList<HashMap<String, double[]>>();
-		
 		Histogram histogramComparison = new Histogram("Comparison of " + bodyType);
 		
 		for (Map.Entry<TreatmentAnalyzer, String> treatment : treatments.entrySet())
 		{
 			Histogram histogram = treatment.getKey().getHistogram();
 			histogramComparison.addDataSet(treatment.getValue(), histogram.getDataSets().get(bodyType));
+			HashMap<String, double[]> datasets = histogram.getDataSets();
 		}
 		
 		histogramComparison.display(100, 5000);
@@ -70,13 +68,15 @@ public class ResultReader {
 			treatment.getKey().setTitle(treatment.getValue());
 		}
 		
-		TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, false);
+		TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, true);
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_50_1.*");   
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_50_5.*");   
 		compareHistogramsByBodyType(treatments, "Tumor");
-		//printTreatmentComparison(treatments);
+		printTreatmentComparison(treatments);
+		
+		
 	}
 
 }

### 2015-03-02 src/utils/Config.java

@@ -96,6 +96,7 @@ public class Config {
 	public static final int rectumType 		= 3; //low-medium dose (liver)
 	public static final int urethraType 	= 4; // no dose (pancreas)
 	public static final int tumorType 		= 5; // high dose
+	public static final int dummyType 		= 6; // high dose
 	
 	//Multithreading
 	

### 2015-03-02 src/erik/TreatmentPlanner.java
Weight parameter

@@ -164,8 +164,8 @@ public class TreatmentPlanner {
 		else if(algo.equals("GA"))
 		{
 			double[] weighting_factors = new double[6];
-			weighting_factors[0] = doubleArgs[4];
-			weighting_factors[1] = 1;
+			weighting_factors[0] = 1;
+			weighting_factors[1] = doubleArgs[4];
 			weighting_factors[2] = doubleArgs[5];
 			weighting_factors[3] = doubleArgs[6];
 			weighting_factors[4] = doubleArgs[7];

### 2015-03-02 src/laurin/ResultReader.java
Changed TA output

@@ -58,7 +58,7 @@ public class ResultReader {
 	    });
 	}
 	
-	public static void printTreatmentComparison(HashMap<TreatmentAnalyzer, String> treatments)
+	public static void printTreatmentComparison(HashMap<TreatmentAnalyzer, String> treatments, boolean showHistograms)
 	{
 		ArrayList<TreatmentAnalyzer> treatmentAnalyzers = new ArrayList<TreatmentAnalyzer>();
 		
@@ -68,13 +68,13 @@ public class ResultReader {
 			treatment.getKey().setTitle(treatment.getValue());
 		}
 		
-		TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, true);
+		TreatmentAnalyzer.printTreatmentComparison(treatmentAnalyzers, showHistograms);
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_50_5.*");   
-		compareHistogramsByBodyType(treatments, "Tumor");
-		printTreatmentComparison(treatments);
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_50_.*");   
+		compareHistogramsByBodyType(treatments, "Rectum");
+		printTreatmentComparison(treatments, false);
 		
 		
 	}

### 2015-03-02 src/laurin/TreatmentAnalyzer.java

@@ -405,7 +405,7 @@ public class TreatmentAnalyzer implements Serializable {
 		String conformalityLine = String.format("%-15s%s", "Conformality", spacing);
 		String homogeinityLine = String.format("%-15s%s", "Homogeinity", spacing);
 		String coverageLine = String.format("%-15s%s", "Coverage", spacing);
-		DecimalFormat decimalFormat = new DecimalFormat("#####0.0000");
+		DecimalFormat decimalFormat = new DecimalFormat("#####0.000E00");
 		
 		for(int i = 0; i < Config.tumorType; i++)
 		{

### 2015-03-02 src/dawid/Looper.java
SNAFU

@@ -9,6 +9,7 @@ import utils.Seed;
 import utils.Voxel;
 import java.lang.Math;
 import java.util.Arrays;
+import java.util.Random;
 import java.util.concurrent.CompletionService;
 import java.util.concurrent.ExecutionException;
 import java.util.concurrent.ExecutorCompletionService;
@@ -129,15 +130,18 @@ public class Looper {
                 }    
             }
         }
-        this.body = this.body2;
+//        this.body = this.body2;
          } else {
          
     //        for(int x=Config.ptvXLow-0; x < Config.ptvXHigh+0; x++) {
-            for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x+= Config.scaleFactor) {
+//              for(int x=Solver.xBoundsTumor[0]; x < Solver.xBoundsTumor[1]; x+= Config.scaleFactor) {                     
+                    for(int x=0; x < Solver.dimensions[0]; x+= Config.scaleFactor) {
     //            for(int y=Config.ptvYLow-0; y < Config.ptvYHigh+0; y++) {
-                for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y+= Config.scaleFactor) {
+//                for(int y=Solver.yBoundsTumor[0]; y < Solver.yBoundsTumor[1]; y+= Config.scaleFactor) {
+                        for(int y=0; y < Solver.dimensions[1]; y+= Config.scaleFactor) {
     //                for(int z=Config.ptvZLow-0; z < Config.ptvZHigh+0; z++) {
-                    for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z+= Config.scaleFactor) {
+//                    for(int z=Solver.zBoundsTumor[0]; z < Solver.zBoundsTumor[1]; z+= Config.scaleFactor) {
+                            for(int z=0; z < Solver.dimensions[2]; z+= Config.scaleFactor) {
     //                    this.body2[x][y][z].setCurrentDosis(0.0);  //Set currentPtvVoxel Dose to 0 
                         this.body2[x][y][z].metavalue = 0.0;
                             for(int i=0; i<Config.numberOfSeeds;++i) { 
@@ -148,13 +152,15 @@ public class Looper {
         //                                                LogTool.print("Cost: Intensity :" + intensity + "@ " + x + " " + y + " " + z,"notification");
                                                         }
[…]

### 2015-03-02 src/erik/Solver.java

@@ -113,7 +113,7 @@ public class Solver {
             looper.setFinalSolution();
             Solver.body = looper.getBody();
             Solver.seeds = looper.getSeeds();
-            for(int i=0; i<Config.SAnumberOfSeeds;i++) {
+            for(int i=0; i<Config.numberOfSeeds;i++) {
 //                            Solver.seeds[i].setDurationMilliSec(looper.getGlobal_Lowest_state()[i]);
                             System.out.println("Verweildauer: "+ Solver.seeds[i].getDurationMilliSec() );
                     }

### 2015-03-02 src/erik/TreatmentPlanner.java

@@ -20,7 +20,7 @@ import utils.Seed;
 import utils.Voxel;
 
 public class TreatmentPlanner {
-	private static final boolean outputToFile = true;
+	private static final boolean outputToFile = false;
 
 	private static void planTreatment(String algo, double doubleArgs[]) throws IloException {
 
@@ -175,7 +175,7 @@ public class TreatmentPlanner {
 		}
 		else if(algo.equals("SA"))
 		{
-			//solver.solveSA();
+			solver.solveSA(doubleArgs);
                         //ToDo: solveSA so modden dass sie Parameter akzeptiert
                         // In meinem Fall MINDESTENS die Seedzahl
                         // Darueber hinaus gibt es nur varianten der

### 2015-03-02 src/utils/Voxel.java

@@ -126,7 +126,7 @@ public class Voxel implements Serializable {
         public double radiationIntensityNoTime(Coordinate position){
             double temp = radiationIntensityNoTimeCALC(position, 90);
             if (temp>0.0) {
-                LogTool.print("Wert in nOtIME " + temp, "notification");
+                //LogTool.print("Wert in nOtIME " + temp, "notification");
             }
 		return temp;
 	}

### 2015-03-02 src/dawid/Looper.java
SNAFU deluxe

@@ -305,15 +305,17 @@ public class Looper {
             LogTool.print("==================== INACTIVE: START CALC FOR OUTER ROUND " + ab + "=========================","notification");
 
                 if (Config.SAverboselvl>1) {
+                    //<editor-fold defaultstate="collapsed" desc="My custom code folding">
                     LogTool.print("SolveSA: Cur_State Read before Metropolis : A)" + Cur_state[0] + " B) " + Cur_state[1] + " C) " + Cur_state[2],"notification");
                     LogTool.print("Debug: GLS get before loop only once each reset: " + this.getGlobal_Lowest_state_string(),"notification");
                     LogTool.print("Debug: GLC_field get before loop only once each reset: " + Cur_cost,"notification");
                     LogTool.print("Debug: GLC_getter get before loop only once each reset: " + this.getCur_cost(),"notification");
                     LogTool.print("Debug: NXC_field get before loop only once each reset: " + New_cost,"notification");
                     LogTool.print("Debug: NXC_getter get before loop only once each reset: " + this.getNew_cost(),"notification");
+                    //</editor-fold>
                 }
-            
             setCur_cost(costCURsuperfinal()); //One of four possible costNEX variants
+            //<editor-fold defaultstate="collapsed" desc="My custom code folding">
                     LogTool.print("CUR COST SUPER FINAL ","notification");
                     LogTool.print("SolveSA: Cur_State Read before Metropolis : A)" + Cur_state[0] + " B) " + Cur_state[1] + " C) " + Cur_state[2],"notification");
                     LogTool.print("Debug: GLS get before loop only once each reset: " + this.getGlobal_Lowest_state_string(),"notification");
@@ -321,8 +323,10 @@ public class Looper {
                     LogTool.print("Debug: GLC_getter get before loop only once each reset: " + this.getCur_cost(),"notification");
                     LogTool.print("Debug: NXC_field get before loop only once each reset: " + New_cost,"notification");
                     LogTool.print("Debug: NXC_getter get before loop only once each reset: " + this.getNew_cost(),"notification");
-            /* [Newstate] with random dwelltimes */
+                    //</editor-fold>
+                /* [Newstate] with random dwelltimes */
             New_state = newstater(); 
+            //<editor-fold defaultstate="collapsed" desc="My custom code folding">
                     LogTool.print("NEW STATE ","notification");
[…]

### 2015-03-03 src/laurin/ResultReader.java
Fix

@@ -51,7 +51,6 @@ public class ResultReader {
 	    }
 	    final Pattern p = Pattern.compile(regex); // careful: could also throw an exception!
 	    return root.listFiles(new FileFilter(){
-	        @Override
 	        public boolean accept(File file) {
 	            return p.matcher(file.getName()).matches();
 	        }
@@ -72,8 +71,8 @@ public class ResultReader {
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_50_.*");   
-		compareHistogramsByBodyType(treatments, "Rectum");
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_100_1.*");   
+		compareHistogramsByBodyType(treatments, "Tumor");
 		printTreatmentComparison(treatments, false);
 		
 		

### 2015-03-03 src/laurin/ResultReader.java
Added comments

@@ -6,16 +6,25 @@ import java.util.ArrayList;
 import java.util.HashMap;
 import java.util.Map;
 import java.util.regex.Pattern;
+
+/**
+ * ResultReader
+ * 
+ * Provides methods for comparison of results
+ * @author Laurin Mordhorst
+ */
 public class ResultReader {
 
 	/**
-	 * @param args
+	 * Get files in folder that match regular expression.
+	 * 
+	 * @param folderName 	name of folder relative to workspace
+	 * @param regex			regular expression
+	 * 
+	 * @return filenames and corresponding TreatmentAnalyzers
 	 */
-	
 	public static HashMap<TreatmentAnalyzer, String> getFiles(String folderName, String regex)
 	{
-		ArrayList<TreatmentAnalyzer> treatmentAnalyzers = new ArrayList<TreatmentAnalyzer>();
-		
 		HashMap<TreatmentAnalyzer, String> treatments = new HashMap<TreatmentAnalyzer, String>();
 		
 		final File folder = new File(folderName);		
@@ -31,6 +40,12 @@ public class ResultReader {
 	    return treatments;
 	}
 	
+	/**
+	 * Create Histogram for given body part from multiple treatment analyzers
+	 * 
+	 * @param treatments		set of treatments
+	 * @param bodyType			body type
+	 */
 	public static void compareHistogramsByBodyType(HashMap<TreatmentAnalyzer, String> treatments, String bodyType)
 	{
 		Histogram histogramComparison = new Histogram("Comparison of " + bodyType);
@@ -42,10 +57,10 @@ public class ResultReader {
 			HashMap<String, double[]> datasets = histogram.getDataSets();
 		}
 		
-		histogramComparison.display(100, 5000);
+		histogramComparison.display(1, 5000);
 	}
 	
-	public static File[] listFilesMatching(File root, String regex) {
+	private static File[] listFilesMatching(File root, String regex) {
 	    if(!root.isDirectory()) {
 	        throw new IllegalArgumentException(root+" is no directory.");
 	    }
@@ -57,6 +72,12 @@ public class ResultReader {
 	    });
 	}
 	
+	/**
+	 * Print comparison of treatments
+	 * 
+	 * @param treatments		set of treatments
+	 * @param showHistograms	if true, histograms will be displayed
+	 */
[…]

### 2015-03-03 src/laurin/ResultReader.java
Fix

@@ -54,7 +54,6 @@ public class ResultReader {
 		{
 			Histogram histogram = treatment.getKey().getHistogram();
 			histogramComparison.addDataSet(treatment.getValue(), histogram.getDataSets().get(bodyType));
-			HashMap<String, double[]> datasets = histogram.getDataSets();
 		}
 		
 		histogramComparison.display(1, 5000);

### 2015-03-03 src/erik/TreatmentPlanner.java
Dont use LUT

@@ -20,7 +20,7 @@ import utils.Seed;
 import utils.Voxel;
 
 public class TreatmentPlanner {
-	private static final boolean outputToFile = false;
+	private static final boolean outputToFile = true;
 
 	private static void planTreatment(String algo, double doubleArgs[]) throws IloException {
 
@@ -35,8 +35,8 @@ public class TreatmentPlanner {
 		 */
 
 		// Database test
-		Config.useLUT = true;
-		Voxel.setLUT(10, 10000);
+		//Config.useLUT = true;
+		//Voxel.setLUT(10, 10000);
 		
 		SimpleDB db = new SimpleDB();
 		

### 2015-03-03 src/erik/TreatmentPlanner.java
Changed to older version

@@ -86,8 +86,9 @@ public class TreatmentPlanner {
 
 						switch (entry.getBodyArray()[x][y][z].getBodyType()) {
 						case Config.normalType: {
-							body[x][y][z].setGoalDosis(Config.normalGoalDose);
-							body[x][y][z].setBodyType(Config.normalType);
+							//body[x][y][z].setGoalDosis(Config.normalGoalDose);
+							//body[x][y][z].setBodyType(Config.normalType);
+							body[x][y][z].setBodyType(Config.dummyType);
 							break;
 						}
 						case Config.bladderType: {
@@ -111,8 +112,9 @@ public class TreatmentPlanner {
 							break;
 						}
 						default: {
-							body[x][y][z].setGoalDosis(Config.normalGoalDose);
-							body[x][y][z].setBodyType(Config.normalType);
+							//body[x][y][z].setGoalDosis(Config.normalGoalDose);
+							//body[x][y][z].setBodyType(Config.normalType);
+							body[x][y][z].setBodyType(Config.dummyType);
 							break;
 						}
 						}
@@ -162,8 +164,8 @@ public class TreatmentPlanner {
 		else if(algo.equals("GA"))
 		{
 			double[] weighting_factors = new double[6];
-			weighting_factors[0] = doubleArgs[4];
-			weighting_factors[1] = 1;
+			weighting_factors[0] = 0;
+			weighting_factors[1] = doubleArgs[4];
 			weighting_factors[2] = doubleArgs[5];
 			weighting_factors[3] = doubleArgs[6];
 			weighting_factors[4] = doubleArgs[7];
@@ -175,7 +177,7 @@ public class TreatmentPlanner {
 		}
 		else if(algo.equals("SA"))
 		{
-			solver.solveSA(doubleArgs);
+			//solver.solveSA();
                         //ToDo: solveSA so modden dass sie Parameter akzeptiert
                         // In meinem Fall MINDESTENS die Seedzahl
                         // Darueber hinaus gibt es nur varianten der
@@ -286,3 +288,4 @@ public class TreatmentPlanner {
 		System.exit(0);				// necessary to terminate all threads 
 	}
 }
+

### 2015-03-04 Paper BMT/article-sample-proceedings.tex
Added latex sources for paper

\documentclass[USenglish,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[big,online]{dgruyter}
\usepackage{hypernat}
\setcitestyle{numbers,square,comma,sort&compress}

\begin{document}

%%%--------------------------------------------%%%
%%% Please do not alter the following 7 lines: %%%
%%%--------------------------------------------%%%
	\articletype{Proceedings}
  \journalname{Current~Directions~in~Biomedical~Engineering}
  \journalyear{2015}
  \journalvolume{1}
  \journalissue{???}
  \startpage{1}
  %\aop
  \DOI{10.1515/bmt-XXXX}
%%%--------------------------------------------%%%

\title{Insert your title here}
\runningtitle{Short title}
%\subtitle{Insert subtitle if needed}

\author*[1]{First Author}
\author[2]{Second Author}
\author[2]{Third Author} 
\runningauthor{F.~Author et al.}

\affil[1]{\protect\raggedright 
  Institution, address of first, e-mail: author\_one@rmi.ge}
\affil[2]{\protect\raggedright 
  Institution, address of second author and third author, e-mail: author\_two@rmi.ge, author\_three@rmi.ge}

\abstract{Hallo na Please insert your abstract here. Remember that online
systems rely heavily on the content of titles and abstracts to
identify articles in electronic bibliographic databases and search
engines. We ask you to take great care in preparing the abstract.}

\keywords{Please insert your keywords here, separated by commas.}

\maketitle

\section{Headline 1st level}

Please insert your manuscript here. \textit{Please insert your manuscript here}. Please insert your manuscript here. Please insert your manuscript here. \textbf{Please insert your manuscript here}. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

New paragraph: Please insert your manuscript here in \citep{2,3,4,5}. Please insert your manuscript here in \citep{1}. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

\subsection{Headline 2nd level}
Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here, see Table \ref{tab:Table1}. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

\begin{table}
\caption{Please insert your table caption here.}
\begin{tabular}{llr}
Table head 1 			& Table head 2 			& Table head 3 		\\ \midrule
Table content 1a 	& Table content 2a 	& Table content 3 \\
Table content 1b	& Table content 2b	& 0.1111 					\\
Table content 1c	& Table content 2c 	& 0.3333 					\\
\end{tabular}
\label{tab:Table1}
\end{table}

\subsubsection{Headline 3rd level}
Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.
Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

\begin{itemize}
\item Item 1
\item Item 2
\end{itemize}

New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here (see Figure \ref{img:Figure1}).

New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

%If possible, please provide figures in the grayscale or CMYK color mode rather than the RGB mode%
\begin{figure}
	\includegraphics[width=\columnwidth]{graphics/figure1}
	\caption{Please insert your figure caption here.}
	\label{img:Figure1}
\end{figure}

\paragraph{Headline 4th level}
Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.

In the subsequent discussion we can conveniently set
$$g=g_{0}, \quad a=a_{0}, \quad x=x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n}$$
$0 \leqslant x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n} \leqslant q-2,$
where $x_{i} \in \{0,1 \}$, for $0 \leqslant i \leqslant n$, and then
rewrite the relation (\ref{eq:1}) in the form

\begin{equation}\label{eq:1}
g_{0}^{x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n}}=a_{0} \mbox{ ,}
\end{equation}
Rising both parts of the last equation to the power $(q-1)/2$ and
taking into account that the following relation ...

\begin{enumerate}
\item Item 1
\item Item 2
\end{enumerate}

\begin{acknowledgement}
Please insert acknowledgments of the assistance of colleagues or similar notes of appreciation here.
\end{acknowledgement}

\def\acknowledgementname{Funding}
\begin{acknowledgement}
Please insert information concerning research grant support here
\end{acknowledgement}

%\bibliographystyle{...}
%\bibliography{...}

\begin{thebibliography}{9}
%---------------------------------------------------------------------------------------------------------------------%
% The Reference list at the end of the manuscript should be in alphanumerical order (see samples below).							%
%---------------------------------------------------------------------------------------------------------------------%

% Books
\bibitem{1}
Eisen HN. Immunology: an introduction to molecular and cellular
principles of the immune response. 5th ed. New York: Harper and
Row 1974.

% Institutional publications:
\bibitem{2}
Institute of Medicine (US). Looking at the future of the Medicaid
program. Washington: The Institute 1999.

% Chapters in books:
\bibitem{3}
Weinsten L, Swartz MN. Pathogenic properties of invading
microorganisms. In: Sodeman WA Jr, Sodeman WA, editors. Pathologic
physiology: mechanisms of disease. Philadelphia: WB Saunders 1974:
457--472.

% Articles in journals with up to 6 authors:
\bibitem{4}
Ge Q, Filip L, Bai A, Nguyen T, Eisen HN, Chen J. Inhibition
of influenza virus production in virus-infected mice by RNA interference.
Proc Natl Acad Sci USA 2004; 101: 8676--8681.

% Articles in journals with more than 6 authors:
\bibitem{5}
You CH, Lee KY, Chey WY, et al. Electrogastrographic study
of patients with unexplained nausea, bloating and vomiting. Gastroenterology 1980; 79: 311--314.

\end{thebibliography}
\end{document}

### 2015-03-04 src/laurin/Histogram.java
Added histogram output to file

@@ -1,6 +1,9 @@
 package laurin;
 
 
+import java.io.File;
+import java.io.FileWriter;
+import java.io.IOException;
 import java.io.Serializable;
 import java.util.HashMap;
 import java.util.Map;
@@ -252,5 +255,34 @@ public class Histogram extends JFrame implements Serializable {
 					max = number;
 			}
 		return max;
+	}
+	
+	public void writeSeriesToFile(String fileName, double normalizationFactor, int numBins)
+	{
+		writeSeriesToFile(fileName, normalizationFactor, numBins, getMin(), getMax());
+	}
+	
+	public void writeSeriesToFile(String fileName, double normalizationFactor, int numBins, double min, double max)
+	{
+		for (Map.Entry<String, double[]> entry : dataSets.entrySet())
+		{
+			System.out.println(numBins);
+			XYSeries xySeries = calcHistogram(entry.getKey(), entry.getValue(), normalizationFactor, numBins, min, max);
+			double[][] xyArray = xySeries.toArray();
+			File file = new File(fileName + "_" + entry.getKey() + ".dat");
+		    try {
+		    	 FileWriter writer = new FileWriter(file, false);
+		    	 for (int i = 0; i < xyArray[0].length; i++)
+		    	 {
+		    		 writer.write(Double.toString(xyArray[0][i]) + " " + Double.toString(xyArray[1][i]) + " \\\\\n");
+		    	 }
+		       writer.flush();
+		       
+		       // Schließt den Stream
+		       writer.close();
+		    } catch (IOException e) {
+		      e.printStackTrace();
+		    }
+		}
 	}
 }

### 2015-03-04 src/laurin/TreatmentAnalyzer.java
Fixed dose calculation error

@@ -197,8 +197,7 @@ public class TreatmentAnalyzer implements Serializable {
 					double dose = 0.0;
 					for (Seed seed : seeds)
 					{
-						dose += seed.radiationIntensity(body[x][y][z].distanceToVoxel(seed.getCoordinate()), seed.getDurationMilliSec());
-						//dose += body[x][y][z].radiationIntensityLUT(seed.getCoordinate(), seed.getDurationMilliSec(), 90);
+						dose += body[x][y][z].radiationIntensity(seed.getCoordinate(), seed.getDurationMilliSec());
 					}
 					body[x][y][z].setCurrentDosis(dose);									
 				}

### 2015-03-04 src/laurin/Histogram.java
Fixed termination problem for Histogram

@@ -129,6 +129,7 @@ public class Histogram extends JFrame implements Serializable {
         setContentPane(chartPanel);
         pack();
         RefineryUtilities.centerFrameOnScreen(this);
+        setDefaultCloseOperation(JFrame.DISPOSE_ON_CLOSE);
         setVisible(true);
 	}
 	

### 2015-03-04 src/erik/Solver.java
Added Stepwise LP support for Solver and TreatmentPlanner

@@ -1,5 +1,8 @@
 package erik;
 
+import java.io.IOException;
+
+import laurin.LPStepwiseTreatment;
 import dawid.Looper;
 import ilog.concert.IloException;
 import thobi.LPTreatment;
@@ -158,6 +161,13 @@ public class Solver {
 		System.out.println("tumorpoint 50-50-50 dose: " + LPTreatment.getBody()[50][50][50].getCurrentDosis());
 		System.out.println("zentrum");
 		*/
+	}
+	
+	public void solveLPSW() throws IloException, IOException
+	{
+		BodyAnalyzer ba = new BodyAnalyzer(body, dimensions, Config.treatmentRange);
+		LPStepwiseTreatment lpStepwiseTreatment = new LPStepwiseTreatment(body, seeds, ba);
+		lpStepwiseTreatment.startGUI();
 	}
 
 

### 2015-03-04 Paper BMT/article-sample-proceedings.tex
Edited the BMT paper

@@
 
-\title{Insert your title here}
-\runningtitle{Short title}
+\title{On optimisation in HDR brachytherapy}
+\runningtitle{On optimisation in HDR brachytherapy}
 %\subtitle{Insert subtitle if needed}
-
-\author*[1]{First Author}
-\author[2]{Second Author}
-\author[2]{Third Author} 
-\runningauthor{F.~Author et al.}
+\author[1]{Laurin Mordhorst}
+\author[2]{Thobias Karthe}
+\author[2]{Sebastian Elm}
+\author[2]{Dawid Golebiewski} 
+\author[2]{Malte Erik Schröder}
+\runningauthor{Mordhorst, L.; Karthe, T.; Elm, S.; Golebiewski, D.; Schröder, M. E.}
 
 \affil[1]{\protect\raggedright 
-  Institution, address of first, e-mail: author\_one@rmi.ge}
+  TUHH,  e-mail: laurin.mordhorst@tuhh.de}
 \affil[2]{\protect\raggedright 
-  Institution, address of second author and third author, e-mail: author\_two@rmi.ge, author\_three@rmi.ge}
+  TUHH,  e-mail: thobias.karthe@tuhh.de, sebastian.elm@tuhh.de, dawid.golebiewski@tuhh.de, malte.schroeder@tuhh.de}
 
@@
 
-\keywords{Please insert your keywords here, separated by commas.}
+\keywords{Optimisation, Treatment planning, Brachytherapy, Genetic Algorithm, Simulated Annealing, Linear Programmin}
 
@@
 
-\section{Headline 1st level}
+\section{Introduction}
 
-Please insert your manuscript here.
-\textit{Please insert your manuscript here}.
-Please insert your manuscript here.
-Please insert your manuscript here.
-\textbf{Please insert your manuscript here}.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+The aim of this article is to compare three different approaches in HDR-Brachytherapy treatment planning.
+The three analysed algorithms are linear programming, simulated annealing and genetic algorithms.
+The results are evaluated by analysing the conformity and homogenity of the dose-distribution as well as runtime for different parameters.
 
-New paragraph: Please insert your manuscript here in \citep{2,3,4,5}.
-Please insert your manuscript here in \citep{1}.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+\section{Genetic Algorithm}
+The Genetic Algorithm as introduced in \citep{1} has a lot of paramters which are described in the following.
 
-\subsection{Headline 2nd level}
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+\subsection{Parameters}
 
-New paragraph: Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here, see Table \ref{tab:Table1}.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+\subsubsection{Weights}
+The Fitness-Function used in the Genetic Algorithm is implemented as a sum of squared distances.
+But every tissue type can be weighted with an individual coefficent.
+Different weightings for the tissue types should clearly visible in the DVH.
 
-\begin{table}
-\caption{Please insert your table caption here.}
-\begin{tabular}{llr}
-Table head 1 			& Table head 2 			& Table head 3 		\\ \midrule
-Table content 1a 	& Table content 2a 	& Table content 3 \\
-Table content 1b	& Table content 2b	& 0.1111 					\\
-Table content 1c	& Table content 2c 	& 0.3333 					\\
-\end{tabular}
-\label{tab:Table1}
-\end{table}
+\subsubsection{Probabilities} 
+There are two important probabilites within the algorithm.
+The first one is the crossover rate.
+It describes how likely it is for two individuals to reproduce.
+The second one is the mutation rate.
+It describes, analogous to the crossover rate, the probability for an individual to perform a mutation.
+\\
+Higher probabilties should cause higher fluctuations within the population.
+This may effect the runtime, because the individuals may change even when a steady state is reached, so the termination criterion is not fullfilled.
 
-\subsubsection{Headline 3rd level}
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+\subsubsection{Scaling and Accuracy}
+The fitness-function for the given optimisation problem has very high computational complexity.
+Hence two scaling parameters are introduced to achieve shorter runtimes.
+The challenge here is to find a compromise between good accuracy for the optimisation and practicable runtimes.\\ The first parameter is called the \textit{treatment range}.
+It chooses the range around the PTV which shall be evaluated by the algorithm.
+The second parameter is a simple scaling value which defines that just every  $n^{th}$ voxel is evaluated.
+Within the treatment range of course.
 
-New paragraph: Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
+\subsection{Results}
 
-\begin{itemize}
-\item Item 1
-\item Item 2
-\end{itemize}
+\subsubsection{Weights}
+The results have shown that the best CI and HI is achieved by a weighting composition of bla bla bla.
+But it should be mentioned, that the weightings should be different for every patient.
+It seems very unlikely to find a set of weighting coefficents, which achieves equally good results for a broader spectrum of patients.
 
-New paragraph: Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here (see Figure \ref{img:Figure1}).
+\subsubsection{Probabilities} 
+The results have shown, that some values are better than others.
 
-New paragraph: Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-
-%If possible, please provide figures in the grayscale or CMYK color mode rather than the RGB mode%
-\begin{figure}
-	\includegraphics[width=\columnwidth]{graphics/figure1}
-	\caption{Please insert your figure caption here.}
-	\label{img:Figure1}
-\end{figure}
-
-\paragraph{Headline 4th level}
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-Please insert your manuscript here.
-
-New paragraph: Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here. Please insert your manuscript here.
-
-In the subsequent discussion we can conveniently set
-$$g=g_{0}, \quad a=a_{0}, \quad x=x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n}$$
-$0 \leqslant x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n} \leqslant q-2,$
-where $x_{i} \in \{0,1 \}$, for $0 \leqslant i \leqslant n$, and then
-rewrite the relation (\ref{eq:1}) in the form
-
-\begin{equation}\label{eq:1}
-g_{0}^{x_{0}+x_{1}2+x_{2}2^{2}+ \cdots +x_{n}2^{n}}=a_{0} \mbox{ ,}
-\end{equation}
-Rising both parts of the last equation to the power $(q-1)/2$ and
-taking into account that the following relation ...
-
-\begin{enumerate}
-\item Item 1
-\item Item 2
-\end{enumerate}
-
+\subsubsection{Scaling and Accuracy}
+The dose distribution of a radioactive seed shows that the dose at a radius of 10 cm is negligible small.
+So any treatment range higher than 10 cm wouldn't lead to better results anyway.
+(TODO: which range is acceptable)
 \begin{acknowledgement}
@@
 % Books
+
 \bibitem{1}
-Eisen HN. Immunology: an introduction to molecular and cellular
-principles of the immune response. 5th ed. New York: Harper and
-Row 1974.
-
-% Institutional publications:
-\bibitem{2}
-Institute of Medicine (US). Looking at the future of the Medicaid
-program. Washington: The Institute 1999.
-
-% Chapters in books:
-\bibitem{3}
-Weinsten L, Swartz MN. Pathogenic properties of invading
-microorganisms. In: Sodeman WA Jr, Sodeman WA, editors. Pathologic
-physiology: mechanisms of disease. Philadelphia: WB Saunders 1974:
-457--472.
-
-% Articles in journals with up to 6 authors:
-\bibitem{4}
-Ge Q, Filip L, Bai A, Nguyen T, Eisen HN, Chen J. Inhibition
-of influenza virus production in virus-infected mice by RNA interference.
-Proc Natl Acad Sci USA 2004; 101: 8676--8681.
-
-% Articles in journals with more than 6 authors:
-\bibitem{5}
-You CH, Lee KY, Chey WY, et al. Electrogastrographic study
-of patients with unexplained nausea, bloating and vomiting. Gastroenterology 1980; 79: 311--314.
+Koza, John R. Genetic Programming. 3rd ed. Cambridge: MIT Press 1992.
 

### 2015-03-05 src/laurin/ResultReader.java
Added Runtime to treatmentcomparison

@@ -91,9 +91,9 @@ public class ResultReader {
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_100_.*");   
-		compareHistogramsByBodyType(treatments, "Tumor");
-		printTreatmentComparison(treatments, false);
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "GA_50_2.0_10.0_0.55_0.8_1.0_1.0_1.0_1.0_1.0_9.0_.*");   
+		compareHistogramsByBodyType(treatments, "Rectum");
+		printTreatmentComparison(treatments, true);
 	}
 
 }

### 2015-03-05 src/laurin/TreatmentAnalyzer.java

@@ -6,15 +6,15 @@ import java.io.IOException;
 import java.io.ObjectInputStream;
 import java.io.ObjectOutputStream;
 import java.io.Serializable;
+import java.text.DateFormat;
 import java.text.DecimalFormat;
+import java.text.SimpleDateFormat;
 import java.util.ArrayList;
-
 import java.util.Date;
 import java.util.HashSet;
 import java.util.Set;
 
 import erik.BodyAnalyzer;
-
 import utils.Config;
 import utils.LogTool;
 import utils.Seed;
@@ -404,7 +404,12 @@ public class TreatmentAnalyzer implements Serializable {
 		String conformalityLine = String.format("%-15s%s", "Conformality", spacing);
 		String homogeinityLine = String.format("%-15s%s", "Homogeinity", spacing);
 		String coverageLine = String.format("%-15s%s", "Coverage", spacing);
+		String runtimeLine = String.format("%-15s%s", "Runtime", spacing);
 		DecimalFormat decimalFormat = new DecimalFormat("#####0.000E00");
+		DateFormat dateFormat = new SimpleDateFormat("mm:ss:SSS");
+		//Formatter to convert Date to String
+		
+
 		
 		for(int i = 0; i < Config.tumorType; i++)
 		{
@@ -432,6 +437,7 @@ public class TreatmentAnalyzer implements Serializable {
 			conformalityLine += String.format("%15s%s", decimalFormat.format(ta.getConformalityIndex()), spacing);
 			homogeinityLine += String.format("%15s%s", decimalFormat.format(ta.getHomogenityIndex()), spacing);
 			coverageLine += String.format("%15s%s", decimalFormat.format(ta.getCoverage()), spacing);
+			runtimeLine += String.format("%15s%s", dateFormat.format(ta.getRuntime()), spacing);
 		}
 		
 		
@@ -456,6 +462,7 @@ public class TreatmentAnalyzer implements Serializable {
 		System.out.println(conformalityLine);
 		System.out.println(homogeinityLine);
 		System.out.println(coverageLine);
+		System.out.println(runtimeLine);
 		System.out.println();
 		
 		if (showHistograms)

### 2015-03-05 Paper BMT/article-sample-proceedings.tex
Added paragraph on model description.

@@
 The three analysed algorithms are linear programming, simulated annealing and genetic algorithms.
-The results are evaluated by analysing the conformity and homogenity of the dose-distribution as well as runtime for different parameters.
+The results are evaluated by analysing the conformity and homogeneity of the dose-distribution as well as runtime for different parameters.
+
+\section{Model description}
+HDR brachytherapy consists of several radioactive sources being placed inside the patient's body by needles.
+These sources will eradiate into the surrounding tissues, treating tumor cells and other critical volumes.
+Therefore, it relies heavily on optimal algorithms to determine the dwell times for certain, predefined goals.
+\\
+The body is represented by an equidistant grid of voxels, defining the single body cell type.
+The seeds are randomly positioned within the tumor voxels.
+Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible.
 

### 2015-03-05 src/laurin/ResultReader.java
merge

@@ -91,8 +91,11 @@ public class ResultReader {
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_100_.*");   
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_100_[0-3][.]{1}.*");   
 		compareHistogramsByBodyType(treatments, "Tumor");
+		compareHistogramsByBodyType(treatments, "Urethra");
+		compareHistogramsByBodyType(treatments, "Rectum");
+		compareHistogramsByBodyType(treatments, "Bladder");
 		printTreatmentComparison(treatments, false);
 	}
 

### 2015-03-05 Paper BMT/article-sample-proceedings.tex
LP start

@@
 \section{Genetic Algorithm}
-The Genetic Algorithm as introduced in \citep{1} has a lot of paramters which are described in the following.
+The Genetic Algorithm as introduced in \citep{1} has a lot of parameters which are described in the following.
 
@@
 (TODO: which range is acceptable)
+
+\section{Linear Programming}
+In Linear Programming a goal function and constraints are used to compute the optimal solution, as described in \citep{2}.
+
+\subsection{Goal function and Constraints}
+bla bla
+
+\subsection{Results}
+bla bla bla
+
 \begin{acknowledgement}
@@
 
+\bibitem{2}
+Williams, H. Paul Model Building in Mathmetical Programming, 4th ed., 1999.
+
 \end{thebibliography}

### 2015-03-05 Paper BMT/article-sample-proceedings.tex
all of LP for today

@@
 \section{Linear Programming}
-In Linear Programming a goal function and constraints are used to compute the optimal solution, as described in \citep{2}.
+In Linear Programming a objective function and constraints are used to compute the optimal solution, as described in \citep{2}.
 
-\subsection{Goal function and Constraints}
-bla bla
+\subsection{Objective function and Constraints}
+The objective function sums up the dose for every organ at risk voxel and is then minimized.
+The constraints restrict the optimization so that the level of radiation in the tumor does not fall under a fixed value.
+The contraints can be relaxed to lower the level of radiation in the surrounding organs at the cost of lowering the radiation at the border of the tumor.
 
@@
 bla bla bla
+unsure bla bla
 

### 2015-03-06 Paper BMT/article-sample-proceedings.tex
Added GenAlg Resultes to the paper

@@
 \subsubsection{Weights}
-The Fitness-Function used in the Genetic Algorithm is implemented as a sum of squared distances.
-But every tissue type can be weighted with an individual coefficent.
-Different weightings for the tissue types should clearly visible in the DVH.
+The Fitness-Function used in the Genetic Algorithm is implemented as a sum of squared distances. But every tissue type can be weighted with an individual coefficent.
 
@@
 It describes, analogous to the crossover rate, the probability for an individual to perform a mutation.
-\\
-Higher probabilties should cause higher fluctuations within the population.
-This may effect the runtime, because the individuals may change even when a steady state is reached, so the termination criterion is not fullfilled.
 
@@
 \subsubsection{Weights}
-The results have shown that the best CI and HI is achieved by a weighting composition of bla bla bla.
-But it should be mentioned, that the weightings should be different for every patient.
-It seems very unlikely to find a set of weighting coefficents, which achieves equally good results for a broader spectrum of patients.
+It seems hard to examine a specific sets of weights, that is applicable to any type of patients.
+The experiment has shown, that a higher weighted treatment has a much higher runtime and the convergency decreases.
 
 \subsubsection{Probabilities} 
-The results have shown, that some values are better than others.
+The results have shown, that a high mutation rate leads to a very low tumor coverage.
+Furthermore the algorithm often doesn't converges, because the changes of the individuals accure too often.
+A stable state becomes less likely.
+\\ For the crossover rate the results showed, that the value should be between 0.7 and 0.9.
 

### 2015-03-07 src/laurin/TreatmentAnalyzer.java
Fix: Runtime initialized

@@ -73,6 +73,7 @@ public class TreatmentAnalyzer implements Serializable {
 		this.anatomies = BodyAnalyzer.splitBodyTypes(body);
 		this.sAlgorithm = "";
 		this.doubleArgs = null;
+		this.runtime = new Date(0);
 		irradiate();
 		analyzeAll();
 	}

### 2015-03-07 Paper BMT/article-sample-proceedings.tex
Added LPSW draft

@@
 \documentclass[USenglish,twocolumn]{article}
+\usepackage{etex}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hypernat}
+\usepackage{tikz}
+\usetikzlibrary{shapes,arrows}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \setcitestyle{numbers,square,comma,sort&compress}
@@
 unsure bla bla
+\section{Stepwise Approach}
+The use of Linear Programming for HDR brachytherapy treatment planning is extended to a stepwise approach as described in \citep{3}.
+
+	\subsection{Principle}
+		Based on a initial solution the planner can take a variable number of steps until an appropiate solution is determined.
+By relaxing bounds for one VOI, an improvement in another VOI might be possible.
+Therefore, the method can be used to identify tradeoffs for conflicting goals by specyfying dose bounds rather than weighting factors.
+
+	\subsection{Sampling}
+		To reduce computational complexity only a subset of voxels is taken into account for the underlying linear program.
+A selection mechanism approximates the significance of voxels as a function of their distance to the closest seed.
+		While distant voxels are more likely to conflict with dose constraints for the PTV, voxels close to a seed are expected to be more informative for the OAR.
+
+	\subsection{Results}
+		The results in table \ref{table:LPSW_results} show different measurements after minimizing the upper dose bound for the OAR while relaxing the PTV lower bound.
+Clearly, OAR sparing and therefore an improvement in conformality comes at the cost of a decreasement in PTV coverage.
+		\begin{table}[ht]
+			\centering		
+		 	\caption{Fixed Pattern Rauschens bei Erhöhung der Umgebungstemperatur}
+		 	\label{table:LPSW_results}
+			\begin{tabular}{ccc}
+				PTV lower bound [Gy] 	& Coverage 	& Conformality Index\\	\hline
+				32.0 					& 1.000		& 1.70\\
+				31.5 					& 0.993 	& 1.61\\
+				31.0 					& 0.979 	& 1.53\\
+				30.5 					& 0.959 	& 1.47\\
+				30.0 					& 0.926		& 1.41\\		
+			\end{tabular}
+		\end{table}
+
+		The quality of the treatment plan depends on the number of seeds that is used for optimization.
+		The steepness of the curves for higher numbers of seeds in figure \ref{fig:LPSW_DVH_numberOfSeeds} indicates better conformality.
+		\begin{figure}[ht]
+			\centering
+			\input{graphics/LPSW_numberOfSeeds.tikz} 
+			\caption{Dose Volume Histograms of the PTV for different numbers of seeds.}
+			\label{fig:LPSW_DVH_numberOfSeeds}
+		\end{figure}
 
@@
 
+\bibitem{3}
+Schlaefer, A. and Schweikard, A. Stepwise multi-criteria optimization for robotic radiosurgery. Med. Phys. 2008.
+
 \end{thebibliography}

### 2015-03-07 Paper BMT/article-sample-proceedings.tex
added parts of PL results

@@
 \subsection{Results}
-bla bla bla
-unsure bla bla
+As shown in tabel \ref{table:LP_results50} and \ref{table:LP_results100}, the relaxation of the PTV increases the conformality, while letting the coverage of the PTV drop. 
+		\begin{table}[h]
+			\centering		
+		 	\caption{LP 50 seeds}
+		 	\label{table:LP_results50}
+			\begin{tabular}{ccc}
+			PTV lower bound [Gy] 	& Coverage 	& Conformality Index\\	\hline
+				32.0 	& 0.999		& 1.23\\
+				31.0 	& 0.988 	& 1.20\\
+				30.0 	& 0.953		& 1.17\\		
+				29.0 	& 0.899		& 1.14\\
+			\end{tabular}
+		\end{table}
+
+Using more initial seeds, the algorithm can achieve higher conformality without losing coverage.
+
+				\begin{table}[h]
+			\centering		
+		 	\caption{LP 100 seeds}
+		 	\label{table:LP_results100}
+			\begin{tabular}{ccc}
+			PTV lower bound [Gy] 	& Coverage 	& Conformality Index\\	\hline
+				32.0 	& 0.999		& 1.16\\
+				31.0 	& 0.981 	& 1.13\\
+				30.0 	& 0.931		& 1.10\\		
+				29.0 	& 0.851		& 1.08\\
+			\end{tabular}
+		\end{table}
+
+mehr tabellen bla. ahh ohh vergleich
+
 \section{Stepwise Approach}

### 2015-03-07 src/laurin/ResultReader.java

@@ -92,8 +92,13 @@ public class ResultReader {
 	}
 	
 	public static void main(String[] args) {
-		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "GA_50_2.0_10.0_0.55_0.8_1.0_1.0_1.0_1.0_1.0_9.0_.*");   
-		compareHistogramsByBodyType(treatments, "Rectum");
+		HashMap<TreatmentAnalyzer, String> treatments = getFiles("treatments/", "LP_100_[0-3][.]{1}.*");   
+		//compareHistogramsByBodyType(treatments, "Tumor");
+		
+		HashMap<TreatmentAnalyzer, String> treatments2 = getFiles("treatments/", "LP_50_[0-3][.]{1}.*");
+		HashMap<TreatmentAnalyzer, String> treatments3 = getFiles("treatments/", "LP_75_[0-3][.]{1}.*");
+		printTreatmentComparison(treatments2, true);
+		printTreatmentComparison(treatments3, true);
 		printTreatmentComparison(treatments, true);
 	}
 

### 2015-03-09 Paper BMT/article-sample-proceedings.tex
Edited LPSW

@@
 	\subsection{Principle}
-		Based on a initial solution the planner can take a variable number of steps until an appropiate solution is determined.
-By relaxing bounds for one VOI, an improvement in another VOI might be possible.
+		Based on an initial solution the planner can take a variable number of steps until an appropiate solution is determined.
+By relaxing bounds for one VOI, an improvement in another VOI might be achieved.
 Therefore, the method can be used to identify tradeoffs for conflicting goals by specyfying dose bounds rather than weighting factors.
@@
 	\subsection{Results}
-		The results in table \ref{table:LPSW_results} show different measurements after minimizing the upper dose bound for the OAR while relaxing the PTV lower bound.
-Clearly, OAR sparing and therefore an improvement in conformality comes at the cost of a decreasement in PTV coverage.
+		The results in Table \ref{table:LPSW_results} show measurements after various optimization steps.
+Based on a coverage optimized solution the urethra and other tissue's upper bound is minimized while the PTV lower bound is relaxed.
+The values for coverage and conformality index (CI) indicate the conflicting multicriteria nature of the problem since one comes at another's cost.
+
+		Improved treatment plans can be achieved by raising the number of seeds for the optimization process as demonstrated in table \ref{table:LPSW_results} and the corresponding dose volume histograms (Figure \ref{fig:LPSW_compare}).
+
+		\begin{figure}[ht]
+			\centering
+			\input{graphics/LPSW_compare.tikz} 
+			\caption{Final DVH for 50 and 100 seeds after multiple optimization steps.}
+			\label{fig:LPSW_compare}
+		\end{figure}
+
 		\begin{table}[ht]
 			\centering		
-		 	\caption{Fixed Pattern Rauschens bei Erhöhung der Umgebungstemperatur}
+		 	\caption{Treatment quality measurements after multiple optimization steps.}
 		 	\label{table:LPSW_results}
-			\begin{tabular}{ccc}
-				PTV lower bound [Gy] 	& Coverage 	& Conformality Index\\	\hline
-				32.0 					& 1.000		& 1.70\\
-				31.5 					& 0.993 	& 1.61\\
-				31.0 					& 0.979 	& 1.53\\
-				30.5 					& 0.959 	& 1.47\\
-				30.0 					& 0.926		& 1.41\\		
+			\begin{tabular}{p{0.24\columnwidth}|p{0.13\columnwidth}|p{0.18\columnwidth}|p{0.13\columnwidth}|p{0.13\columnwidth}}
+					Step 					  	& Seeds  & Tumor LB 	& Coverage 	& CI\\	\hline
+					Optimization for coverage 	& ~50 	 & 32.0 		& 0.995		& 2.14\\ \hline
+					Min. of urethra UB 		  	& ~50	 & 31.0 		& 0.993		& 1.74\\ \hline
+					Min. of other UB 			& ~50	 & 30.5			& 0.972		& 1.53\\ \hline \hline
+					Min. of other UB 			& 100	 & 30.5			& 0.973		& 1.28\\
 			\end{tabular}
 		\end{table}
-
-		The quality of the treatment plan depends on the number of seeds that is used for optimization.
-		The steepness of the curves for higher numbers of seeds in figure \ref{fig:LPSW_DVH_numberOfSeeds} indicates better conformality.
-		\begin{figure}[ht]
-			\centering
-			\input{graphics/LPSW_numberOfSeeds.tikz} 
-			\caption{Dose Volume Histograms of the PTV for different numbers of seeds.}
-			\label{fig:LPSW_DVH_numberOfSeeds}
-		\end{figure}
 

### 2015-03-09 Paper BMT/article-sample-proceedings.tex
New Graphics and tables included

@@
 \usetikzlibrary{shapes,arrows}
+\usepackage{multirow}
 \usepackage{pgfplots}
@@
 The experiment has shown, that a higher weighted treatment has a much higher runtime and the convergency decreases.
+\begin{figure}[ht]
+\caption{Difference between the goal dose and the average dose}
+\begin{tikzpicture}
+  \begin{axis}[
+  	name=one,
+  	width=0.86 \columnwidth,
+  	height=1.0in, 
+  	scale only axis,
+  	xmajorticks=false,
+  	xtick={0,10,20,30,40},
+  	ytick={-20,-10,...,10},
+  	minor ytick={-20,-15,...,10},
+  	legend style={anchor=north,at={(0.5,1.4)},font=\small},
+  	legend columns=2,
+    xmin=0,
+    xmax=40,
+  	ymax = 5,
+  	ymin = -20,
+    grid=both,
+    y label style={at={(axis description cs:0.1,.5)},anchor=south},
+    ylabel = {Diff. to GD [Gy]}
+    ]
+	\addplot [color=black,dashed,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -7.4061373\\
+						20 -7.3660937\\
+						30 -7.16107\\
+						40 -7.5223036\\	};
+	\addlegendentry{Bladder};
+
+	\addplot [color=black,dotted,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -6.6003648\\
+						20 -7.250544\\
+						30 -8.3557603\\
+						40 -5.8634733\\ };
+	\addlegendentry{Rectum};
+
+	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -17.7968793\\
+						20 -16.8252388\\
+						30 -17.5772547\\
+						40 -15.4201025\\ };
+	\addlegendentry{Urethra};
+
+	\addplot [color=black,solid,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+	 					10 -1.0171151\\
+	 					20 -1.1222384\\
+				 		30 -1.0559538\\
+	 					40  0.6725971\\};
+	\addlegendentry{Tumor};					
+  \node at (rel axis cs:1,1) [anchor=north east] {Bladder weighted};	
+  \end{axis}
+
+  \begin{axis}[
+  	name=two,
+  	width=0.86 \columnwidth,
+  	height=1.0in, 
+  	scale only axis,
+  	at=(one.below south),
+    anchor=above north,
+  	xmajorticks=false,
+  	xtick={0,10,20,30,40},
+  	ytick={-20,-0,...,100},
+  	minor ytick={-10,0,...,100},
+    xmin=0,
+    xmax=40,
+  	ymax = 95,
+  	ymin = -20,
+    grid=both,
+    y label style={at={(axis description cs:0.1,.5)},anchor=south},
+    ylabel = {Diff. to GD [Gy]}
+    ]
+	\addplot [color=black,dashed,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 46.8163804\\
+						20 -6.5099569\\
+						30 -6.4596163\\
+						40 45.5999754\\ };
+
+	%\addlegendentry{Bladder};
+
+	\addplot [color=black,dotted,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 46.4315353\\
+						20 -7.0846292\\
+						30 -8.312817\\
+						40 44.4589502\\ };
+
+	%\addlegendentry{Rectum};
+
+	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 73.834936\\
+						20 -16.600937\\
+						30 -15.1440168\\
+						40	80.79331\\  };
+	%\addlegendentry{Urethra};
+
+	\addplot [color=black,solid,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+	 					10 95.820661\\
+	 					20 -0.5231224\\
+	 					30 -1.3135843\\
+	 					40	94.652913\\ };
+	%\addlegendentry{Tumor};					
+  \node at (rel axis cs:1,1) [anchor=north east] {Rectum weighted};	
+  \end{axis}
+  \begin{axis}[
+  	name=three,
+  	width=0.86 \columnwidth,
+  	height=1.0in, 
+  	scale only axis,
+  	at=(two.below south),
+    anchor=above north,
+  	xmajorticks=false,
+  	xtick={0,10,20,30,40},
+  	ytick={-20,-0,...,100},
+  	minor ytick={-10,0,...,100},
+  	legend style={anchor=north,at={(0.5,1.3)},font=\small},
+  	legend columns=2,
+    xmin=0,
+    xmax=40,
+  	ymax = 95,
+  	ymin = -20,
+    grid=both,
+    y label style={at={(axis description cs:0.1,.5)},anchor=south},
+    ylabel = {Diff. to GD [Gy]}
+    ]
+	\addplot [color=black,dashed,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 46.8163804\\
+						20 -6.5099569\\
+						30 45.5999754\\
+						40 -6.4596163\\ };
+
+	%\addlegendentry{Bladder};
+
+	\addplot [color=black,dotted,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 46.4315353\\
+						20 -7.0846292\\
+						30 44.4589502\\
+						40 -8.312817\\ };
+
+	%\addlegendentry{Rectum};
+
+	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 73.834936\\
+						20 -16.600937\\
+						30 80.79331\\
+						40 -15.1440168\\ };
+	%\addlegendentry{Urethra};
+
+	\addplot [color=black,solid,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+	 					10 95.820661\\
+	 					20 -0.5231224\\
+	 					30 94.652913\\
+	 					40 -1.3135843\\ };
+	%\addlegendentry{Tumor};					
+  \node at (rel axis cs:1,1) [anchor=north east] {Urethra weighted};	
+  \end{axis}
+  \begin{axis}[
+  	name=four,
+  	width=0.86 \columnwidth,
+  	height=1.0in, 
+  	scale only axis,
+  	at=(three.below south),
+    anchor=above north,
+  	xmajorticks=true,
+  	xtick={0,10,20,30,40},
+  	ytick={-20,-10,...,10},
+  	minor ytick={-15,-10,...,5},
+    xmin=0,
+    xmax=40,
+  	ymax = 10,
+  	ymin = -20,
+    grid=both,
+    xlabel = {Weighting Factor},
+    y label style={at={(axis description cs:0.1,.5)},anchor=south},
+    ylabel = {Diff. to GD [Gy]}
+    ]
+	\addplot [color=black,dashed,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -7.9814571\\
+						20 -2.0920392\\
+						30 -6.5044591\\
+						40 -6.4056077\\ };
+
+	%\addlegendentry{Bladder};
+
+	\addplot [color=black,dotted,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -8.2390329\\
+						20 -3.2641176\\
+						30 -6.7960123\\
+						40 -6.5058489\\ };
+
+	%\addlegendentry{Rectum};
+
+	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+						10 -18.198227\\
+						20 -6.5576085\\
+						30 -18.4460054\\
+						40 -16.7663932\\ };
+	%\addlegendentry{Urethra};
+
+	\addplot [color=black,solid,line width=1.0pt]
+	table[row sep=crcr]{0 0\\
+	 					10 -1.8981359\\
+	 					20 7.2659238\\
+	 					30 -0.6839885\\
+	 					40 -0.955956\\ };
+	%\addlegendentry{Tumor};					
+  \node at (rel axis cs:1,1) [anchor=north east] {Tumor weighted};	
+  \end{axis}
+
+\end{tikzpicture}
+\end{figure}
 
@@
 \\ For the crossover rate the results showed, that the value should be between 0.7 and 0.9.
-
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{table}[h]
+\begin{tabular}{c|ccccc}
+\textbf{MR}          & \textbf{CR :}     & 0.3   & 0.5   & 0.7   & 0.9   \\ \hline
+\multirow{2}{*}{0.3} & \textbf{CI}       & 1,092 & 1,019 & 1,683 & 1,101 \\
+                     & \textbf{Coverage} & 0,778 & 0,387 & 1,000 & 0,770 \\ \hline
+\multirow{2}{*}{0.5} & \textbf{CI}       & 1,028 & 1,032 & 1,022 & 1,079 \\
+                     & \textbf{Coverage} & 0,360 & 0,410 & 0,395 & 0,403 \\ \hline
+\multirow{2}{*}{0.7} & \textbf{CI}       & 1,007 & 1,046 & 1,012 & 1,035 \\
+                     & \textbf{Coverage} & 0,281 & 0,316 & 0,011 & 0,341 \\ \hline
+\multirow{2}{*}{0.9} & \textbf{CI}       & 1,034 & 1,000 & 1,109 & 1,043 \\
+                     & \textbf{Coverage} & 0,256 & 0,001 & 0,736 & 0,273 \\
+\end{tabular}
+\end{table}
 \subsubsection{Scaling and Accuracy}

### 2015-03-24 Paper Seminar/isim_paper.tex
Seminar paper hinzugefügt

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{cite}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\begin{document}
\tableofcontents
\section{Dose-Function}
The dose-distribution within the patient's body is the accumulation of all the radiation-seeds. Therefore it is mandatory to know the radiation intensity of each seed. Equation \eqref{eq:dosefunction} shows the general approach to approximate the dose intensity function and \eqref{eq:dosefunction2} shows the used implementation of that equation.
\begin{equation}
\label{eq:dosefunction}
\dot{D}(r,\theta_{0})= S_{k}\Lambda \frac{G_{L}(r,\theta_{0})}{G_{L}(r_{0},\theta_{0})} g_{L}(r) F(r_{0},\theta_{0})
\end{equation}
\begin{equation}
\label{eq:dosefunction2}
\Rightarrow \ \dot{D}(r,\theta_{0})= S_{k}\Lambda \frac{G_{L}(r,\theta_{0})}{G_{L}(r_{0},\theta_{0})} g_{L}(r) \cdot 1
\end{equation}
\\
The following section will describe the used paramters:
\begin{description}
\item[Air-Kerma-Strength $S_{k}$]~\\
The $S_{k}$ describes the \textbf{k}inetic \textbf{e}nergy \textbf{r}eleased per unit \textbf{ma}ss called kerma. It is measured in U. Where $1\ U =1 \ \frac{cGy \cdot m^{2}}{h} $   
\item[Dose Rate Constant $\Lambda$]~\\
The dose rate constant is defined as the ratio of the dose rate at reference point $ P(r_{0},\theta_{0})  $ and the the air-kerma-strength $S_{k}$. Therefore $ \Lambda \ = \ \frac{\dot{D}(r_{0},\theta_{0})}{S_{k}} $.

\item[Dose Function $g_{L}(r)$]~\\
The dose function describes the dose intensity dependent on the radial distance.

\item[Two dimensional anisotropy function $ F(r_{0},\theta_{0})$]~\\
The anisotropy resolves issues due to the fact, that the dose distribution is'nt rotation invariant. Therefore the dose intensity is different for different angles.

\item[Geometry Function $G_{L}(r_{0},\theta_{0})$ ]~\\
Due to scatter and attenuation an estimation of the dose distribution with point sources is less precise then an estimation with a line source. Therefore a line source model is used. \ref{fig:geometry} gives an overview of the geometry of the model.

\begin{figure}[hbtp]

\centering
\includegraphics[width=0.7\textwidth]{pictures/geometry}
\caption{Geometry of a line source}
\label{fig:geometry}	
\end{figure}

\end{description}
\newpage

Figure \ref{fig:doseplot} shows the plotted dose function. It is easily seen that the curve has a realy steep slope. Hence it should be discussed for which radii the given function should be evaluated.

\begin{figure}[htbp]

\centering
\includegraphics[width=0.9\textwidth]{pictures/dosefunction}
\caption{Dose-Function Plot}
\label{fig:doseplot}
\end{figure}
\section{Genetic Algorithm}
\label{sec:genalg}

Genetic algorithm  or evolutionary algorithm \cite{genpro} describes a search procedure which belongs to the category local search\cite{aima}. Local search algorithms are used when the path to your solution is irrelavant (unlike e.g. breadth first search). The idea of the genetic algorithm is to find the best solution in a solution space by adapting the procedure of genetic reproduction to the problem. \\ 
The algorithm consists mainly of three components:

\begin{description}
\item[Individual] ~\\
\label{indiv}
An individual contains one possible solution for the problem. It could be seen as the interface between the problem and the algorithm.  It is forced to reproduce itself with other individuals so the solution might improve. The key here is to find an efficient encoding for your problem.
\item[Population]~\\
\label{pop}
The sum of all individuals is called population. The population provides the gene pool. Every individual shall reproduce it's gene with another one out of the population. 
\item[Fitness-Function] ~\\
\label{fifu}
The fitnees-function is a heuristic which is used to evaluate the fitness value of an individual. The fitness value indicates how good the solution is, which the individual contains. Further the fitness value is used to select the best individuals for reproduction.
\end{description}

The genetic algorithm is an iterative algorithm. Each iteration consists of four steps, which are described in the following:

\begin{description}
\item[1. Elitism]~\\
In the elitism step a specific number of the best individuals are taken into the next iteration without further reproduction or other actions. 
\item[2. Selection]~\\
During the selection phase each individual gets selected with a probabilty, that is proportional to it's fitnessvalue.
$P_{selection_{i}}= \tfrac{fitness_{i}}{\sum_{j}^I fitness_{j}}$ Where $I$ is the Population size and $i$ the specific index for a individual.
\item[3. Crossover]~\\
In the crossover step the previously selected individuals are taken pairwise to exchange their solution by using the one point crossover. One point crossover cuts the gene in two parts at a randomly chosen point and exchanges one half of each's gene.

\item[4. Mutation]~\\
In the end each individual changes randomly some parts of it's solution by a given probability.
\end{description}

\subsection{Implementation}
\label{subsec:impl}

The individual is ipmlemented as a $n^{th}$ dimensional vector of the dwell times, where $n$ denotes the number of seeds. Equation \eqref{eq:individual} shows one individual for $n  $ Seeds.
 \begin{equation}
 \label{eq:individual}
 Individual \ \ i := \begin{pmatrix}
 t_{0} \\ \vdots \\ t_{n-1} 	
\end{pmatrix}   
\end{equation}

 The population is, like in \ref{pop} said, the accumulation of all individuals. $k$ denotes the number of individuals.
\begin{equation}
 \label{eq:population}
 Population \ \ p := \begin{pmatrix}
 i_{0} \\ \vdots \\ i_{k-1} 	
\end{pmatrix}   
 \end{equation}

 The Fitness-Function calculates the squared differnce between the goaldose and the current dose within the whole body.
 \begin{equation}
 f(individual) = \sum_{body}(GD-CD)^2
 \end{equation}
 For further adjustment and prioritisation the fitness-function is extended by an weighting factor, which holds a specifig weight for any bodytype (tumor, liver, spine,...).
 \begin{equation}
 \label{eq:fifu}
 f(individual) = \sum_{body}w_{bodytype}(GD-CD)^2
 \end{equation}

\subsubsection{Optimisation}
\label{subsubsec:optimisation}
The fitness-function as described in equation \eqref{eq:fifu} is shown in algorithm \ref{alg:fifu}. It is easily seen that it iterates over every voxel in the body. This results in a very high computational complexity and high runtime. To improve the performance of the algorithm three different optimisation approaches are introduced. \\

\begin{algorithm}[H]
\label{alg:fifu}
 $temp=0$\;
 \For{x=0; x $\textless$ body.xDim; x++}{
 	\For{y=0; y $\textless$ body.yDim; y++}{
 		\For{z=0; z $\textless$ body.zDim; z++}{
 			\For{j=0; j $\textless$ numberOfSeeds; j++} {
				 	$temp\ \ = temp+  w_{bodytype}(GD-CD)^2$\;		
 			}
  		}
  	}
  }
 $fitnessvalue = \sqrt{temp}$\;
 \Return{fitnessvalue}\;
 \caption{Calculation of the fitnessvalue}

 \end{algorithm}

\paragraph{Scaling}
\label{para:scaling}
~\\
This strategy evaluates not every single voxel in the body. Only every $n^{th}$ voxel is evaluated. Algortithm \ref{alg:scalefif} shows how this is implemented. \\

\paragraph{Treatment Range}
\label{para:treatment range}
~\\
Like the dosefunction chapter has shown, the distribution of the radiation decreases exponential with the distance. This effect could be used to reduce the number of voxel, that has to be evaluated, because the influence of voxels that are relatively far away is negligibly small. Algorithm \ref{alg:scalefif} shows how that information is used. The Region of Interest (roi) denotes the treatment range. \\

\begin{algorithm}[H]
\label{alg:scalefif}

 \For{y=y-roi.min; y $\textless$ y-rio.max; y$\leftarrow$y+Scalefactor}{
 		\For{z=z-roi.min; z $\textless$ z-rio.max; z$\leftarrow$z+Scalefactor}{
 			\For{j=0; j $\textless$ numberOfSeeds; j++} {
				 	$temp\ \ = temp+  w_{bodytype}(GD-CD)^2$\;		
 			}
  		}
  	}
\caption{Optimised Fitness-Function}
\end{algorithm}

\paragraph{Multithreading}
\label{para:multithreading}
~\\ 
The multithreading approach seperates the 
three-dimensional
body into two-dimensional planes. Each plane is then
processed in an own thread. Algorithm \ref{alg:multithread} shows that approach.\\

 \begin{algorithm}[H]
 \label{alg:multithread}
 \SetKwFunction{cb}{CalculateBounds}
 \SetKwFunction{thread}{ThreadedFitnessFunction}

 \emph{x-roi,y-roi,z-roi $\leftarrow$} \cb{body}  \;
 \For{x-roi}{
 	\thread{x,y-roi,z-roi}
 }
 \caption{Multithreaded Fitness-Function}
 \end{algorithm}

\subsubsection{Stop Criterion}
\label{subsubsec: stopcrit}
One of the challenges for the genetic algorithm is to decide when and how the algorithm should terminate. It is difficult to detmine how close the current iteration is to the optimal solution. The approach used in this implementation calculates the difference between the current fitness-value and the fitness-value of the previous iteration. If the absolute value of this differnece is less or equal to a small boundary value a counter is incremented. If this counter is incremented five times in a row, the algorithm terminates. \\ 
This procedure shall ensure that small fluctuations, that can accure due to the mutations within the individuals, will have no effect on the termination. \\

\begin{algorithm}[H]
\SetKwFunction{iter}{GeneticAlgIteration}
$individualOld,currentIndividual \leftarrow 0$\;
$stop \leftarrow false$\;
$counter \leftarrow 0$\;
\While{! stop} {
 \tcc{Returns the best fitness value from the current iteration}

	$currentIndividual \leftarrow $\iter{}\;

	\If{$\left|individualOld-currentIndividual \right| \ \ \textless \ \ 1$} {
		\emph{counter++}
	} 
	\If{$counter \geq 5$} {
		$done \leftarrow true$
	}  
	$individualOld \leftarrow currentIndividual $

	}

\caption{Stop-Criterion}
\end{algorithm}

\newpage
\subsubsection{Parameters}
The implemented algorithm has a set of parameters, which is used to adjust it's outcome. The usage and the influence of those values are discussed in the following.

\paragraph{Weights}~\\
The fitness-function as described in \ref{fifu} uses individual weighting factors for each body type.  But every tissue type can be weighted with an individual coefficent.

\paragraph{Probabilities} ~\\ 
There are two important probabilites within the algorithm. The first one is the crossover rate. It describes how likely it is for two individuals to reproduce. The second one is the mutation rate. It describes, analogous to the crossover rate, the probability for an individual to perform a mutation.

\paragraph{Scaling and Accuracy}~\\
The fitness-function for the given optimisation problem has very high computational complexity. Hence two scaling parameters are introduced to achieve shorter runtimes. The challenge here is to find a compromise between good accuracy for the optimisation and practicable runtimes.\\ The first parameter is the treatment range. It chooses the range around the PTV which shall be evaluated by the algorithm. The second parameter is a simple scaling value which defines that just every  $n^{th}$ voxel is evaluated. Within the treatment range of course.

\subsubsection{Results}

\paragraph{Weights}
It seems hard to examine a specific sets of weights, that is applicable to any type of patients. The experiment has shown, that a higher weighted treatment has a much higher runtime and the convergency decreases.  Figure \ref{fig:weights} shows the difference between the goal dose and the actual dose for different weighted distributions.
\begin{figure}[hbtp]
\begin{tikzpicture}
  \begin{axis}[
  	name=one,
  	width=0.86 \columnwidth,
  	height=1.0in, 
  	scale only axis,
  	xmajorticks=false,
  	xtick={0,10,20,30,40},
  	ytick={-20,-10,...,10},
  	minor ytick={-20,-15,...,10},
  	legend style={anchor=north,at={(0.5,1.6)},font=\small},
  	legend columns=2,
    xmin=0,
    xmax=40,
  	ymax = 5,
  	ymin = -20,
    grid=both,
    y label style={at={(axis description cs:0,.5)},anchor=south},
    ylabel = {Diff. to GD [Gy]}
    ]
	\addplot [color=black,dashed,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -7.4061373\\
						20 -7.3660937\\
						30 -7.16107\\
						40 -7.5223036\\	};
	\addlegendentry{Bladder};

	\addplot [color=black,dotted,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -6.6003648\\
						20 -7.250544\\
						30 -8.3557603\\
						40 -5.8634733\\ };
	\addlegendentry{Rectum};

	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -17.7968793\\
						20 -16.8252388\\
						30 -17.5772547\\
						40 -15.4201025\\ };
	\addlegendentry{Urethra};

	\addplot [color=black,solid,line width=1.0pt]
	table[row sep=crcr]{0 0\\
	 					10 -1.0171151\\
	 					20 -1.1222384\\
				 		30 -1.0559538\\
	 					40  0.6725971\\};
	\addlegendentry{Tumor};					
  \node at (rel axis cs:1,1) [anchor=north east] {Bladder weighted};	
  \end{axis}

  \begin{axis}[
  	name=two,
  	width=0.86 \columnwidth,
  	height=1.0in, 
  	scale only axis,
  	at=(one.below south),
    anchor=above north,
  	xmajorticks=false,
  	xtick={0,10,20,30,40},
  	ytick={-20,-0,...,100},
  	minor ytick={-10,0,...,100},
    xmin=0,
    xmax=40,
  	ymax = 95,
  	ymin = -20,
    grid=both,
    y label style={at={(axis description cs:0,.5)},anchor=south},
    ylabel = {Diff. to GD [Gy]}
    ]
	\addplot [color=black,dashed,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 46.8163804\\
						20 -6.5099569\\
						30 -6.4596163\\
						40 45.5999754\\ };

	%\addlegendentry{Bladder};

	\addplot [color=black,dotted,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 46.4315353\\
						20 -7.0846292\\
						30 -8.312817\\
						40 44.4589502\\ };

	%\addlegendentry{Rectum};

	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 73.834936\\
						20 -16.600937\\
						30 -15.1440168\\
						40	80.79331\\  };
	%\addlegendentry{Urethra};

	\addplot [color=black,solid,line width=1.0pt]
	table[row sep=crcr]{0 0\\
	 					10 95.820661\\
	 					20 -0.5231224\\
	 					30 -1.3135843\\
	 					40	94.652913\\ };
	%\addlegendentry{Tumor};					
  \node at (rel axis cs:1,1) [anchor=north east] {Rectum weighted};	
  \end{axis}
  \begin{axis}[
  	name=three,
  	width=0.86 \columnwidth,
  	height=1.0in, 
  	scale only axis,
  	at=(two.below south),
    anchor=above north,
  	xmajorticks=false,
  	xtick={0,10,20,30,40},
  	ytick={-20,-0,...,100},
  	minor ytick={-10,0,...,100},
  	legend style={anchor=north,at={(0.5,1.3)},font=\small},
  	legend columns=2,
    xmin=0,
    xmax=40,
  	ymax = 95,
  	ymin = -20,
    grid=both,
    y label style={at={(axis description cs:0,.5)},anchor=south},
    ylabel = {Diff. to GD [Gy]}
    ]
	\addplot [color=black,dashed,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 46.8163804\\
						20 -6.5099569\\
						30 45.5999754\\
						40 -6.4596163\\ };

	%\addlegendentry{Bladder};

	\addplot [color=black,dotted,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 46.4315353\\
						20 -7.0846292\\
						30 44.4589502\\
						40 -8.312817\\ };

	%\addlegendentry{Rectum};

	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 73.834936\\
						20 -16.600937\\
						30 80.79331\\
						40 -15.1440168\\ };
	%\addlegendentry{Urethra};

	\addplot [color=black,solid,line width=1.0pt]
	table[row sep=crcr]{0 0\\
	 					10 95.820661\\
	 					20 -0.5231224\\
	 					30 94.652913\\
	 					40 -1.3135843\\ };
	%\addlegendentry{Tumor};					
  \node at (rel axis cs:1,1) [anchor=north east] {Urethra weighted};	
  \end{axis}
  \begin{axis}[
  	name=four,
  	width=0.86 \columnwidth,
  	height=1.0in, 
  	scale only axis,
  	at=(three.below south),
    anchor=above north,
  	xmajorticks=true,
  	xtick={0,10,20,30,40},
  	ytick={-20,-10,...,10},
  	minor ytick={-15,-10,...,5},
    xmin=0,
    xmax=40,
  	ymax = 10,
  	ymin = -20,
    grid=both,
    xlabel = {Weighting Factor},
    y label style={at={(axis description cs:0,.5)},anchor=south},
    ylabel = {Diff. to GD [Gy]}
    ]
	\addplot [color=black,dashed,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -7.9814571\\
						20 -2.0920392\\
						30 -6.5044591\\
						40 -6.4056077\\ };

	%\addlegendentry{Bladder};

	\addplot [color=black,dotted,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -8.2390329\\
						20 -3.2641176\\
						30 -6.7960123\\
						40 -6.5058489\\ };

	%\addlegendentry{Rectum};

	\addplot [color=black,dash pattern=on 7pt off 7pt,line width=1.0pt]
	table[row sep=crcr]{0 0\\
						10 -18.198227\\
						20 -6.5576085\\
						30 -18.4460054\\
						40 -16.7663932\\ };
	%\addlegendentry{Urethra};

	\addplot [color=black,solid,line width=1.0pt]
	table[row sep=crcr]{0 0\\
	 					10 -1.8981359\\
	 					20 7.2659238\\
	 					30 -0.6839885\\
	 					40 -0.955956\\ };
	%\addlegendentry{Tumor};					
  \node at (rel axis cs:1,1) [anchor=north east] {Tumor weighted};	
  \end{axis}

\end{tikzpicture}
\caption{Difference between the goal dose and the average dose}
\label{fig:weights}
\end{figure}

\newpage
\paragraph{Probabilities} 
The results have shown, that a high mutation rate leads to a very low tumor coverage. Furthermore the algorithm often doesn't converges, because the changes of the individuals accure too often. A stable state becomes less likely. \\ For the crossover rate the results showed, that the value should be between 0.7 and 0.9.
% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{table}[h]
\label{table:probabilities}
\begin{tabular}{c|ccccc}
\textbf{MR}          & \textbf{CR :}     & 0.3   & 0.5   & 0.7   & 0.9   \\ \hline
\multirow{2}{*}{0.3} & \textbf{CI}       & 1,092 & 1,019 & 1,683 & 1,101 \\
                     & \textbf{Coverage} & 0,778 & 0,387 & 1,000 & 0,770 \\ \hline
\multirow{2}{*}{0.5} & \textbf{CI}       & 1,028 & 1,032 & 1,022 & 1,079 \\
                     & \textbf{Coverage} & 0,360 & 0,410 & 0,395 & 0,403 \\ \hline
\multirow{2}{*}{0.7} & \textbf{CI}       & 1,007 & 1,046 & 1,012 & 1,035 \\
                     & \textbf{Coverage} & 0,281 & 0,316 & 0,011 & 0,341 \\ \hline
\multirow{2}{*}{0.9} & \textbf{CI}       & 1,034 & 1,000 & 1,109 & 1,043 \\
                     & \textbf{Coverage} & 0,256 & 0,001 & 0,736 & 0,273 \\
\end{tabular}
\caption{Mutation rate and coverage for different mutation-  \& crossover rates}

\end{table}
\subsubsection{Scaling and Accuracy}
The dose distribution of a radioactive seed shows that the dose at a radius of 10 cm is negligible small. So any treatment range higher than 10 cm wouldn't lead to better results.

\bibliographystyle{plain}
\bibliography{lit}
\end{document}

### 2015-03-26 Paper Seminar/isim_paper.tex
started with LP part..... a little more to come considering optimization and stuff

@@
 \usepackage{pgfplotstable}
+\usepackage{amssymb}
 
@@
 \tableofcontents
+
+\section{Model description}
+HDR brachytherapy consists of several radioactive sources being placed inside the patient's body by needles.
+These sources will eradiate into the surrounding tissues, treating tumor cells and other critical volumes.
+Therefore, it relies heavily on optimal algorithms to determine the dwell times for certain, predefined goals.
+\\
+The body is represented by an equidistant grid of voxels, defining the single body cell type.
+The seeds are randomly positioned within the tumor voxels.
+Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible.
+
 \section{Dose-Function}
@@
 
+\section{Linear Programming}
+In Linear Programming an objective function and constraints are used to compute the optimal solution to a minimization or maximization problem, as described in [?].
+The problem consists of a certain number of variables and a feasible region, which is defined by the contraints.
+Equation \eqref{eq:LPMatrix} shows the standard form of a linear problem.
+The objective function, modeled with \textit{c},  is to be minimized, while the constraints, modeled with \textit{A} and \textit{b}, are satisfied.
+The used algorithm yields a vector x for the optimal solution.
+
+ \begin{equation}
+ \label{eq:LPMatrix}
+ \begin{aligned}
+ A \in \mathbb{R}^{m,n}, \ \ \ b \in \mathbb{R}^m, \ \ \ c \in \mathbb{R}^n, \ \ \ x \in \mathbb{R}^n \\
+ objective function \ \ \ c^Tx = Min! \\
+ min\{c^Tx|Ax \leq b, x \geq 0\}
+\end{aligned}
+ \end{equation}
+
+ \subsection{Implementation}
+
+For our problem we used CPLEX, an implementation of the simplex algorithm by IBM[?].
+In simplex the feasible region is considered a polyhedron.
+The optimal solution must be in one of the vertices of the polyhedron.
+Simplex finds a start vertex and approaches the optimal solution by going on an edge from vertex to vertex, until there is no better solution.
+The constraints allow us a good modelling of the tumor, since we are able to define a constraint with certain limits for each body type.
+
+\paragraph{Basic Funtionality}
+The basic functionality is depicted in algorithm \ref{alg:cplex}.
+This implementation checks every voxel of the body, if it is a tumor voxel or not.
+In case of a tumor voxel the influence of the radiation of every seed will be summed up, building a greater-equal constraint.
+For body voxel the influence of the radiation builds a term for the objective function.
+\begin{algorithm}
+\label{alg:cplex}
+ \For{x=0; x $\textless$ body.xDim; x++}{
+ 	\For{y=0; y $\textless$ body.yDim; y++}{
+ 		\For{z=0; z $\textless$ body.zDim; z++}{
+			\If{$body[x][y][z].getBodyType() == tumorType$} {
+				\For{i=0; i $\textless$ numberOfSeeds; i++}{
+					$dosepart[i] \gets cplex.prod(dose(x,y,z,i),time[i])$	
+				}
+				$cplex.addGe(cplex.sum(dosepart),goalDosis(x,y,z));$
+				}
+			\Else{\For{i=0; i $\textless$ numberOfSeeds; i++}{
+					$objective.addTerm(dose(x,y,z,i),time[i]);$	
+				}
+
+				}
+
+  		}
+  	}
+  }
+  $cplex.addMinimize(objective);$\\
+  $cplex.solve();$
+ \caption{basic functionality}
+ \end{algorithm}
+
+\subsection{Stepwise Approach}
+
 \bibliographystyle{plain}

### 2015-03-27 Paper Seminar/isim_paper.tex
Added stepwise part for seminar paper

@@
 \usepackage{amssymb}
+\usepackage{subcaption}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible.
+
+\section{Evaluation criteria}
+	Reflecting the multicriteria nature of the problem, several measurements are considered for treatment evaluation.
+	\begin{description}
+		\item[Coverage (CO)]~\\
+			The fraction of PTV volume that receives at least the prescription dose. Therefore, the CO is always smaller than or equal to 1 and becomes 1 ideally.
+		\item[Homogeneity (HI)]~\\
+			The upper PTV bound divided by lower PTV bound. Therefore, the HI is always greater than or equal to 1 and becomes 1 ideally.
+		\item[Conformality Index (CI)]~\\
+			The volume receiving at least the prescription dose divided by the PTV volume receiving at least the prescription dose. Therefore, the CI is always greater than or equal to 1 and becomes 1 ideally.
+		\item[Minimum, maximum and average values]~\\
+			The extreme and average values per volume of interest.
+		\item[Dose Volume Histogram (DVH)]~\\
+			A two-dimensional representation of the dose distribution.
+For a given dose, the cumulative DVH shows the fraction of a volume that is exposed to at least the given dose.
+Figure \ref{fig:evalCritDVH} shows a typical DVH.
+			\begin{figure}[ht]
+				\centering
+				\input{pictures/EvalCrit_DVH.tikz} 
+				\caption{A cumulative Dose Volume Histogram}
+				\label{fig:evalCritDVH}
+			\end{figure}
+
+\end{description}
 
@@
 \subsection{Stepwise Approach}
+	The use of linear programming for HDR brachytherapy treatment planning is extended to a stepwise approach as described in \cite{schlaefer}.
+
+	\subsubsection{Idea}
+		\label{subsec:lpswIdea}
+		Based on an initial solution the planner can take a variable number of steps until an appropiate solution is determined.
+By relaxing bounds for one VOI, an improvement in another VOI might be achieved.
+Therefore, the method can be used to identiy tradeoffs for conflicting goals by specyfying dose bounds rather than weighting factors.
+Figure \ref{fig:lpswWorkflow} presents the basic workflow of this method.
+
+		\begin{figure}[ht]
+			\centering
+			\input{pictures/lpswWorkflow.tikz} 
+			\caption{Workflow for the Stepwise Approach}
+			\label{fig:lpswWorkflow}
+		\end{figure}
+
+	\subsubsection{Optimization problem}
+		The underlying linear program as formulated in equation \ref{eq:LPMatrix} is modified to meet the requirements arising from section \ref{subsec:lpswIdea}.
+Equation \ref{eq:lpswMatrix} shows the resulting linear problem.
+The slack variables $\hat{t}$, $\check{t}$, $\hat{s}$ and $\check{s}$ measure the deviation from the desired lower or upper bound.
+While $\hat{t}$ and $\check{t}$ are valid for a whole VOI, $\hat{s}$ and $\check{s}$ correspond to a single voxel of a VOI.
+In an optimization step only one VOI is taken into account by setting the coefficient values to 1, specyfying the desired bounds and minimizing the objective function.
+		\begin{equation}
+			\label{eq:lpswMatrix}
+			\begin{matrix}
+			\textit{min} &            & ~\hat{c}_s^T\hat{s} & +\check{c}_s^T\check{s}    & +\hat{c}_t^T\hat{t} & +\check{c}_t^T\check{t} & & \\\\
+			\textit{s.t.} & Ax         & -\hat{s}            &            & -\hat{t}  &            & \leq & b_u\\
+     & Ax         & 		            & +\check{s} &           & +\check{t} & \geq & b_l\\
+			   	 & \hat{s}    & 		            &            &           &            & \leq & b_{\hat{s}}\\
+          & \check{s}  & 	                &            &           &            & \leq & b_{\check{s}}\\
+          & \hat{t}    & 		            &            &           &            & \leq & b_{\hat{t}}\\
+          & \check{t}  & 	                &            &           &            & \leq & b_{\check{t}}\\
+				 & \multicolumn{2}{c}{x,\hat{s},\check{s},\hat{t},\check{t}} & & & & \geq & 0
+			\end{matrix}		
+		\end{equation}
+
+	\subsubsection{Sampling}
+		To reduce computational complexity only a subset of voxels is taken into account for the linear program.
+A selection mechanism approximates the significance of voxels as a function of their distance to the closest seed.
+Distant voxels are more likely to conflict with dose constraints for the PTV.
+Therefore, the voxels with the greatest distance are chosen.
+In contrast, voxels close to a seed are expected to be more informative for the OAR since they are more likely to conflict with the upper bound dose constraints of the corresponding VOI.
+
+	\subsubsection{Results}
+		The results in Table \ref{table:lpswTable} show measurements after various optimization steps.
+Based on a coverage optimized solution the urethra and other tissue's upper bound is minimized while the PTV lower bound is relaxed.
+Figure \ref{fig:lpswComparison} shows a comparison of the dose volume histograms after the last optimization step for 50 and 100 seeds.
+
+		\begin{table}[ht]	
+		 	\caption{Treatment quality measurements after multiple optimization steps}
+		 	\label{table:lpswTable}
+			\begin{tabular}{L{0.35\textwidth}|C{0.1\textwidth}|C{0.18\textwidth}|C{0.1\textwidth}|C{0.1\textwidth}}
+					\textbf{Step} 			& \textbf{Seeds}  & \textbf{Tumor LB} 	& \textbf{CO} 	& \textbf{CI}\\	\hline
+					Optimization for coverage 	& ~50 	 & 32.0 		& 0.995		& 2.14\\ \hline
+					Min. of urethra UB 		  	& ~50	 & 31.0 		& 0.993		& 1.74\\ \hline
+					Min. of other UB 			& ~50	 & 30.5			& 0.972		& 1.53\\ \hline
+					Min. of other UB 			& 100	 & 30.5			& 0.973		& 1.28\\
+			\end{tabular}
+		\end{table}
+
+		\begin{figure}[htbp]
+					\begin{subfigure}[t]{\textwidth}
+						%\centering
+						\input{pictures/lpsw50.tikz} 
+						\caption{50 seeds}
+						\label{fig:lpswComparison50Seeds}
+					\end{subfigure}
+					\begin{subfigure}[t]{\textwidth}
+						%\centering
+						\input{pictures/lpsw100.tikz} 
+						\caption{100 seeds}
+						\label{fig:lpswComparison10Seeds}
+					\end{subfigure}
+			\caption{Final dose volume histograms with different numbers of seeds used for optimization}
+			\label{fig:lpswComparison}
+		\end{figure}
+
+	\subsubsection{Discussion}
+		The values for coverage and conformality index (CI) listed in Table \ref{table:lpswTable} indicate the conflicting multicriteria nature of the problem since one comes at another's cost.
+Therefore, the balancing of goals by relaxing dose constraints seems to be a reasonable approach to this problem.
+
+		Improved treatment plans can be achieved by raising the number of seeds for the optimization process as demonstrated in the last two rows of Table \ref{table:lpswTable} and the corresponding dose volume histograms (Figure \ref{fig:lpswComparison}).
+The increased quality of a treatment plan comes at the cost of more computational complexity and therefore a longer computation time.
 
@@
 \bibliography{lit}
+
+\begin{thebibliography}{9}
+\bibitem{schlaefer}
+	Schlaefer, A. and Schweikard, A. Stepwise multi-criteria optimization for robotic radiosurgery. Med. Phys. 2008.
+
+\end{thebibliography}
 \end{document}

### 2015-03-28 Paper Seminar/classification.tex
Added file itself

A wonderful classification. To be edited. Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!

### 2015-03-28 Paper Seminar/isim_paper.tex
Removed duplicate bibliography
added basic title page and VOI OAR PTV definitions in model description
Included classification file

@@
 \begin{document}
+
+\title{Treatment planning for HDR Brachytherapy}
+\date{March 30, 2015}
+\author{S.~Elm \and
+T.~Karthe \and
+M.~E.~Schroeder \and
+D.~Golebiewski \and
+L.~Mordhorst}
+
+\maketitle
+\newpage
+
 \tableofcontents
@@
 Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible.
+\\
+In the model the tumor is called the planning target volume (PTV) and the surrounding organs are called organs at risk (OAR).
+The different body types might also be referred as volume of interest (VOI).
 
@@
 \end{description}
+
+\section{Classification}
+\input{classification}
 
@@
 The optimal solution must be in one of the vertices of the polyhedron.
-Simplex finds a start vertex and approaches the optimal solution by going on an edge from vertex to vertex, until there is no better solution.
+Simplex finds a start vertex and approaches the optimal solution by moving from vertex to vertex on an edge, until there is no better solution.
 The constraints allow us a good modelling of the tumor, since we are able to define a constraint with certain limits for each body type.
 
-\paragraph{Basic Funtionality}
-The basic functionality is depicted in algorithm \ref{alg:cplex}.
-This implementation checks every voxel of the body, if it is a tumor voxel or not.
-In case of a tumor voxel the influence of the radiation of every seed will be summed up, building a greater-equal constraint.
-For body voxel the influence of the radiation builds a term for the objective function.
-\begin{algorithm}
+\subsubsection{Basic Funtionality}
+
+\begin{algorithm}[H]
 \label{alg:cplex}
@@
 				}
-
 				}
@@
  \end{algorithm}
+
+The basic functionality is depicted in algorithm \ref{alg:cplex}.
+This implementation checks every voxel of the body, if it is a tumor voxel or not.
+In case of a tumor voxel the influence of the radiation of every seed will be summed up, building a greater-equal constraint.
+For body voxels the influence of the radiation builds a term for the objective function.
+
+ \subsubsection{Results}
+ The constraints restrict the optimization so that the level of radiation in the PTV does not fall under a fixed value.
+The contraints can be relaxed to lower the level of radiation in the surrounding OAR at the cost of lowering the radiation at the border of the PTV.
+
+ As shown in table \ref{table:LP_results50} and \ref{table:LP_results100}, the relaxation of the PTV increases the conformality, while letting the coverage of the PTV drop. 
+		\begin{table}[h]
+			\centering		
+		 	\caption{Treatment quality with 50 seeds and relaxation}
+		 	\label{table:LP_results50}
+			\begin{tabular}{C{0.35\textwidth}|C{0.11\textwidth}|C{0.1\textwidth}}
+			\textbf{PTV lower bound [Gy]} 	& \textbf{CO} 	& \textbf{CI}\\	\hline
+				32.0 	& 0.999		& 1.23\\	\hline
+				31.0 	& 0.988 	& 1.20\\	\hline
+				30.0 	& 0.953		& 1.17\\	\hline		
+				29.0 	& 0.899		& 1.14\\
+			\end{tabular}
+		\end{table}
+
+Using more initial seeds, the algorithm can achieve higher conformality without losing coverage.
+
+				\begin{table}[h]
+			\centering		
+		 	\caption{Treatment quality with 100 seeds and relaxation}
+		 	\label{table:LP_results100}
+			\begin{tabular}{C{0.35\textwidth}|C{0.11\textwidth}|C{0.1\textwidth}}
+			\textbf{PTV lower bound [Gy]} 	& \textbf{CO} 	& \textbf{CI}\\	\hline
+				32.0 	& 0.999		& 1.16\\ \hline
+				31.0 	& 0.981 	& 1.13\\ \hline
+				30.0 	& 0.931		& 1.10\\ \hline	
+				29.0 	& 0.851		& 1.08\\
+			\end{tabular}
+		\end{table}
 
@@
 
-\begin{thebibliography}{9}
-\bibitem{schlaefer}
-	Schlaefer, A. and Schweikard, A. Stepwise multi-criteria optimization for robotic radiosurgery. Med. Phys. 2008.
-
-\end{thebibliography}
 \end{document}

### 2015-03-28 Paper Seminar/classification.tex
Fixed file

@@
-A wonderful classification.
-To be edited.
-Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!Make it seen!
+While using computer aided systems to create treatment plans, this involves time-consuming calculations during optimizations (as seen in this project).
+Therefore, a classification attempt is proposed, using prior knowledge to reduce the amount of time taken to process.
+
+\subsection{General Idea}
+When classifying, you search for similarities or analogies in your given set of data. These provide a basis for a comparison among previously created optimiziations

### 2015-03-28 Paper Seminar/classification.tex
Edited the file. Inserted general idea, parameter description and algorithm.

@@
 
-\subsection{General Idea}
-When classifying, you search for similarities or analogies in your given set of data. These provide a basis for a comparison among previously created optimiziations
+\subsection{General Idea and Application}
+When classifying, you search for similarities or analogies in your given set of data.
+These provide a basis for a comparison among previously created optimizations.
+The former result being the most similar (or closest with respect to the comparison) can be taken as a starting point for the next iteration.
+
+In treatment planning, there are several steps to be executed when trying to classify.
+First, the body to be treated ("treated body") needs to be analyzed (i.e.
+measured in various variables).
+Second, these measurements build a formula providing a norm (like the euclidian norm), which then can be used to compare the bodies.
+Afterwards, the body with the smallest distance to the treated body is taken.
+If it has an optimization result stored ("stored body), this will be used during the next iteration.
+This aims at lower runtimes for at least an equal coverage.
+
+\subsection{Parameters and Influences}
+In table \ref{tbl:classification_parameters}, the various parameters measured in order to compare are displayed.
+Further on, their influence in the norm is shown.
+Additionally, in table \ref{tbl:classification_type_weights}, the weighted influence of the five different body types are presented.
+\begin{table}
+\centering
+\caption{Parameters and Influences in classification}
+\begin{tabular}[htbp]{l | c | r}
+\textbf{Parameter} & \textbf{Influence Type} & \textbf{Weight} \\
+volume size  & relative & 0.1\\
+volume center & relative & 0.2 \\
+distance tumor-center & absolute & 0.2\\
+closest distance & absolute & 0.5
+\end{tabular}
+\label{tbl:classification_parameters}
+\end{table}
+
+\begin{description}
+\item[Volume Size]~\\
+The plain count of all voxels respective to the different body types. As the body is given as an equidistant 3D-array, it's comparable without calculation of the real volume.
+\item[Volume Center]~\\ The center of the volume of each body type, with respect to the coordinates within the 3D-array.
+For calculation, all coordinates of the body type's voxels are summed up and divided by the volume size.
+\item[Distance Tumor-Center]~\\ The distance between the center of the tumor center and the center of the other body type volumes, calculated by a simple substraction.
+\item[Closest Distance]~\\ The closest distance between two volumes in units, corresponding to the 3D-array's distance inbetween the voxels.
+\end{description}
+
+The four parameters descripted above are calculated in a measuring algorithm, detailed further in section \ref{classification:algorithm}.
+They contain two types of influence onto the norm: textit{relative} and \textit{absolute}.
+\textit{relative} depicts a comparison with respect to the treated body's values of the same parameter.
+For example, the tumor volume size of the stored body will be divided by the tumor volume size of the treated body.
+In constrast, the \textit{absolute} comparison omits the division completely.
+This is due to division by 0, when the tumor overlaps with other body type volumes.
+\\
+
+They each share a weight parameter, which sums up to 1 in total.
+The values shown in table \ref{tbl:classification_parameters} represent a multiplication factor within the norm.
+The norm itself is a sum of the different parameters multiplied with their weight, corresponding to the smallest error between the stored and treated bodies.
+The weights correspond to a simple consideration about how much each parameter should influence the the classification.
+To allow for small to medium deviations within the three first parameters (volume size, volume center and distance tumor-center), they have been chosen small.
+This enables a few transformations (i.e.
+translation of single/all volumes) of the body volumes resulting in a similar classification.
+Therefore, the weight for closest distance has been chosen larger, to prevent extreme transformations to distort the result.
+\\
+
+Additionally to the weights of the various weights of the measurement parameters, it's possible to change the influence of the different body types (compare table \ref{tbl:classification_type_weights}).
+This leads to a more detailed and precise representation of the aims during the treatment planning.
+In this case, a high focus has been set upon not damaging the spine and the pancreas whilst treating the tumor properly.
+
+\begin{table}
+\centering
+\caption{Influences of Body Types}
+\begin{tabular}[htbp]{c | c | c | c | c}
+\textbf{normal} & \textbf{spine} & \textbf{liver} & \textbf{pancreas} & \textbf{tumor} \\
+1 & 5 & 0.5 & 5 & 3
+\end{tabular}
+\label{tbl:classification_type_weights}
+\end{table}
+
+\subsection{Algorithm and Performance}\label{classification:algorithm}
+The algorithm is shown in algorithm \ref{alg:classification}. It consists of a main loop, iterating through the stored body
+
+\begin{algorithm}[H]
+\label{alg:classification}
+\ForAll{voxels in body}
+	count body_type \;
+	sum up coordinates \;
+	\If{body_type == tumor}
+		store coordinate for later usage \;
+	\EndIf
+\EndFor
+volumeCenter = coordinates / body_type \;
+centerDistance = euclid_norm (tumorCenter, volumeCenter) \;
+\ForAll{voxels in body}
+	\ForAll{voxels in tumor stored}
+		calculated distance (voxel, tumor) \;
+		\If{distance $\textless$ closestDistance}
+			update closestDistance \;
+		\EndIf
+	\EndFor
+\EndFor
+\caption{Measurement algorithm in pseudo code}
+\end{algorithm}

### 2015-03-28 Paper Seminar/classification.tex
cleared up Elmos mess, including the algo and the tables

@@
 \caption{Parameters and Influences in classification}
-\begin{tabular}[htbp]{l | c | r}
-\textbf{Parameter} & \textbf{Influence Type} & \textbf{Weight} \\
-volume size  & relative & 0.1\\
-volume center & relative & 0.2 \\
-distance tumor-center & absolute & 0.2\\
+\begin{tabular}[htbp]{l | c | c}
+\textbf{Parameter} & \textbf{Influence Type} & \textbf{Weight} \\ \hline
+volume size  & relative & 0.1\\ \hline
+volume center & relative & 0.2 \\ \hline
+distance tumor-center & absolute & 0.2\\ \hline
 closest distance & absolute & 0.5
@@
 \begin{tabular}[htbp]{c | c | c | c | c}
-\textbf{normal} & \textbf{spine} & \textbf{liver} & \textbf{pancreas} & \textbf{tumor} \\
+\textbf{normal} & \textbf{spine} & \textbf{liver} & \textbf{pancreas} & \textbf{tumor} \\ \hline
 1 & 5 & 0.5 & 5 & 3
@@
 \label{alg:classification}
-\ForAll{voxels in body}
-	count body_type \;
-	sum up coordinates \;
-	\If{body_type == tumor}
-		store coordinate for later usage \;
-	\EndIf
-\EndFor
-volumeCenter = coordinates / body_type \;
-centerDistance = euclid_norm (tumorCenter, volumeCenter) \;
-\ForAll{voxels in body}
-	\ForAll{voxels in tumor stored}
-		calculated distance (voxel, tumor) \;
-		\If{distance $\textless$ closestDistance}
-			update closestDistance \;
-		\EndIf
-	\EndFor
-\EndFor
+\ForAll{$voxels\ in\ body$}{
+	$count\ body\_type;$ \\
+	$sum\ up\ coordinates;$\\
+	\If{$body\_type == tumor$}{
+		$store\ coordinate\ for\ later\ usage;$
+	}
+}
+$volumeCenter = coordinates / body\_type;$ \\
+$centerDistance = euclid\_norm (tumorCenter, volumeCenter);$ \\
+\ForAll{$voxels\ in\ body$}{
+	\ForAll{$voxels\ stored\ in\ tumor$}{
+		$calculated\ distance (voxel, tumor);$\\
+		\If{distance $\textless$ closestDistance}{
+			$update\ closestDistance;$
+		}
+	}
+}
 \caption{Measurement algorithm in pseudo code}

### 2015-03-28 Paper Seminar/classification.tex
Added evaluation and prospects, finished algorithm. Waiting for Thobi to fix the code listing.

--- 2015-03-28 Paper Seminar/classification.tex
@@
 The four parameters descripted above are calculated in a measuring algorithm, detailed further in section \ref{classification:algorithm}.
-They contain two types of influence onto the norm: textit{relative} and \textit{absolute}.
+They contain two types of influence onto the norm: \textit{relative} and \textit{absolute}.
 \textit{relative} depicts a comparison with respect to the treated body's values of the same parameter.
@@
 
-\subsection{Algorithm and Performance}\label{classification:algorithm}
-The algorithm is shown in algorithm \ref{alg:classification}. It consists of a main loop, iterating through the stored body
+\subsection{Algorithm}\label{classification:algorithm}
+The algorithm is shown in algorithm \ref{alg:classification}. It consists of a main loop, iterating through the stored body and performing the following measurements and additional tasks. \\
+First, the body type counter for the current voxel is incremented (line 2).
+Second, the x-, y- and z-coordinate of the voxel is added to total coordinate of the body type (line 3).
+This aims at the calculation of the volume center.
+If the body type is of \textit{tumor}, then the voxel coordinate is stored for later usage, i.e.
+the calculation of the closest distance parameter (line 4 \& 5).
+Afterwards, the parameter volume center and distance tumor-center are calculated by division and euclidian norm (line 6 \& 7).\\
+To calculate the closest distance between the volumes, it's necessary cross-compare the voxels.
+Therefore, another loop iterates through the complete body, and, within, iterates over all previously stored tumor coordinates.
+For each pair of voxel and tumor coordinate, the distance is calculated and, if less then before, the closest distance is updated (line 8-12).
 
-\begin{algorithm}[H]
-\label{alg:classification}
-\ForAll{voxels in body}
-	count body_type \;
-	sum up coordinates \;
-	\If{body_type == tumor}
-		store coordinate for later usage \;
-	\EndIf
-\EndFor
-volumeCenter = coordinates / body_type \;
-centerDistance = euclid_norm (tumorCenter, volumeCenter) \;
-\ForAll{voxels in body}
-	\ForAll{voxels in tumor stored}
-		calculated distance (voxel, tumor) \;
-		\If{distance $\textless$ closestDistance}
-			update closestDistance \;
-		\EndIf
-	\EndFor
-\EndFor
-\caption{Measurement algorithm in pseudo code}
-\end{algorithm}
+%\begin{algorithm}[H]
+%\label{alg:classification}
+%\ForAll{voxels in body}
+%	count body_type \;
+%	sum up coordinates \;
+%	\If{body_type == tumor}
+%		store coordinate for later usage \;
+%	\EndIf
+%\EndFor
+%volumeCenter = coordinates / body_type \;
+%centerDistance = euclid_norm (tumorCenter, volumeCenter) \;
+%\ForAll{voxels in body}
+%	\ForAll{voxels in tumor stored}
+%		calculated distance (voxel, tumor) \;
+%		\If{distance $\textless$ closestDistance}
+%			update closestDistance \;
+%		\EndIf
+%	\EndFor
+%\EndFor
+%\caption{Measurement algorithm in pseudo code}
+%\end{algorithm}
+
+\subsection{Performance evalution and prospects}
+Without the calculation of the closest distance in between the tumor volume and all other volumes, the performance is linear.
+The computing takes only a few seconds (2-4s), using a body similar to dimensions of 110x110x200.
+Alas, the computing time increases exponentially when performing the cross-comparison of each voxel in the body with all voxels of type \textit{tumor}.
+A first step of optimization has been taken when introducing a cache for the tumor voxel (already included in the algorithm above).
+Another step towards lower runtimes contains the implementation of a multi-threaded cross-comparison or completely multi-threaded algorithm.
+The multi-threaded cross-comparison has been implemented and it indicates a speed up by approximately the number of processor cores used, but has not beend thoroughly tested yet.
+\\
+Like the multi-threading, at the time of this paper, the classification and the results of it are not properly tested.
+Therefore, no scientific statement can be made.
+However, for linear programming no influence upon the runtime has been observed.
+When using the genetic algorithm or simulating annealing, only a small deviation of runtime has been observed.
+Unluckily it is not clear, whether the deviation is caused by the standard variation of the algorithms runtime, or by the influence of classification.\\
+Therefore, a major part of future work should contain the testing of the classification. Afterwards, a founded evaluation and adaptions according to the evaluation can be made.

### 2015-03-28 Paper Seminar/classification.tex

@@
 
-%\begin{algorithm}[H]
-%\label{alg:classification}
-%\ForAll{voxels in body}
-%	count body_type \;
-%	sum up coordinates \;
-%	\If{body_type == tumor}
-%		store coordinate for later usage \;
-%	\EndIf
-%\EndFor
-%volumeCenter = coordinates / body_type \;
-%centerDistance = euclid_norm (tumorCenter, volumeCenter) \;
-%\ForAll{voxels in body}
-%	\ForAll{voxels in tumor stored}
-%		calculated distance (voxel, tumor) \;
-%		\If{distance $\textless$ closestDistance}
-%			update closestDistance \;
-%		\EndIf
-%	\EndFor
-%\EndFor
-%\caption{Measurement algorithm in pseudo code}
-%\end{algorithm}
+\ForAll{$voxels\ in\ body$}{
+	$count\ body\_type;$ \\
+	$sum\ up\ coordinates;$\\
+	\If{$body\_type == tumor$}{
+		$store\ coordinate\ for\ later\ usage;$
+	}
+}
+$volumeCenter = coordinates / body\_type;$ \\
+$centerDistance = euclid\_norm (tumorCenter, volumeCenter);$ \\
+\ForAll{$voxels\ in\ body$}{
+	\ForAll{$voxels\ stored\ in\ tumor$}{
+		$calculated\ distance (voxel, tumor);$\\
+		\If{distance $\textless$ closestDistance}{
+			$update\ closestDistance;$
+		}
+	}
+}
+\caption{Measurement algorithm in pseudo code}
+\end{algorithm}
 

### 2015-03-29 Paper Seminar/isim_paper.tex
Fixed ref errors, did some beauty work (i.e. newpage after toc)
added bib stuff for LP

@@
 M.~E.~Schroeder \and
-D.~Golebiewski \and
+%D.~Golebiewski \and
 L.~Mordhorst}
@@
 \tableofcontents
+\newpage
 
@@
 \section{Linear Programming}
-In Linear Programming an objective function and constraints are used to compute the optimal solution to a minimization or maximization problem, as described in [?].
+In Linear Programming an objective function and constraints are used to compute the optimal solution to a minimization or maximization problem, as described in [\cite{matprog}].
 The problem consists of a certain number of variables and a feasible region, which is defined by the contraints.
@@
 
-For our problem we used CPLEX, an implementation of the simplex algorithm by IBM[?].
+For our problem we used CPLEX, an implementation of the simplex algorithm by IBM[\cite{cplex}].
 In simplex the feasible region is considered a polyhedron.
@@
 
+\newpage
 \bibliographystyle{plain}

### 2015-03-29 Paper Seminar/isim_paper.tex
Merge I guess.
Added introductary paragraph. Additional libtex source, therefore lib.bib updated.

@@
 \title{Treatment planning for HDR Brachytherapy}
-\date{March 30, 2015}
+\date{March 31st, 2015}
 \author{S.~Elm \and
@@
 M.~E.~Schroeder \and
-%D.~Golebiewski \and
-L.~Mordhorst}
+D.~Golebiewski \and
+L.~Mordhorst
+}
 
@@
 \newpage
+
+\section{Introduction}
+High dose rate brachytherapy (abbreviated: hdr brachytherapy) as a conformal radiation therapy allows a much higher cancer tissue ratio, compared to other treatment approaches.
+During hdr brachytherapy, radioactive sources (i.e.
+seeds) are placed within the patient's body, in, close to and around the cancer tissue.
+Then, the seeds eradiate into the surrounding tissues, remaining a pre-calculated time frame (dwell time).
+Therefore, seed placement and dwell times are the most important factors when analyzing brachytherapy treatments \cite{brachytherapy}.
+\\
+The focus of this article is set upon optimizing the dwell times for a fixed seed placement using computer aided algorithms, i.e.
+genetic algorithms and linear programming, as well as step-wise linear programming.
+The different algorithms will be described.
+Further on, they are analyzed and evaluated upon their adjustment possibilities (like parameters).
+Additionally, conclusions about the quality of the resulting treatment plans are drawn.
+Lastly, a possible approach to reduce runtim complexity and reuse of prior knowledge is presented.
 
@@
 \end{description}
-
-\section{Classification}
-\input{classification}
 
@@
 
+\section{Classification}
+\input{classification}
+
 \newpage

### 2015-03-29 Paper Seminar/isim_paper.tex
Rewritten body model, added picture for example body.

@@
 The different algorithms will be described.
-Further on, they are analyzed and evaluated upon their adjustment possibilities (like parameters).
+Further on, they are analyzed and evaluated upon their adjustment possibilities (like parameters), subject to different treatment goals.
 Additionally, conclusions about the quality of the resulting treatment plans are drawn.
@@
 \section{Model description}
-HDR brachytherapy consists of several radioactive sources being placed inside the patient's body by needles.
-These sources will eradiate into the surrounding tissues, treating tumor cells and other critical volumes.
-Therefore, it relies heavily on optimal algorithms to determine the dwell times for certain, predefined goals.
+As a data basis for the optimization, a simple body model is used.
+
+\subsection{Body modle}
+The body model is supplied as a three-dimensional array ("body array"), containing the type of the body cells, within a matlab file (.mat).
+The body array is set up as an equidistant point cloud of cells (1 unit $\widehat{=}$ 0.2cm), being transformed into 3D-array of voxels (compare algorithm \ref{model:voxel}).
+Besides the coordinate, a voxel contains several dosis variables (like extreme dosis, goal dosis) and the body type (normal tissue, spine, liver, pancreas and tumor tissue).
 \\
-The body is represented by an equidistant grid of voxels, defining the single body cell type.
-The seeds are randomly positioned within the tumor voxels.
-Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible.
-\\
-In the model the tumor is called the planning target volume (PTV) and the surrounding organs are called organs at risk (OAR).
-The different body types might also be referred as volume of interest (VOI).
+\begin{algorithm}[H]
+maxDosis: \textit{double} [cGy] \\
+minDosis: \textit{double} [cGy] \\
+goalDosis: \textit{double} [cGy] \\
+currentDosis: \textit{double} [cGy] \\
+coordinate: \textit{Coordinate} [x: units, y: units, z: units] \\
+bodyType: int
+\label{model:voxel}
+\caption{Properties of a voxel}
+\end{algorithm}
 
-\section{Evaluation criteria}
+An example body is displayed in figure \ref{model:body}.
+Here, the planning target volume (PTV, i.e.
+the tumor) and the other volumes of interest (VOI) are displayed.
+The VOIs might also be referred to as organs at risk (OAR).
+The ptv is in the middle of the figure (drawn red), surrounded by the liver (green).
+The yellow, cylindric volume behind the liver on the right represents the spine.
+At last, the pancreas is drawn as a black sphere (bottom right from the tumor).
+\begin{figure}
+\centering
+\includegraphics[width=.7\textwidth]{pictures/body}
+\caption{Example body with tumor}
+\label{model:body}
+\end{figure}
+
+Inside the body, the seeds are positioned randomly within the tumor voxels.
+When eradiating into the surrounding tissue, the seeds influence overlap and form a superposition.
+Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limite, where the influence gets negligible (see chapter \ref{dose_function}).
+
+\subsection{Evaluation criteria}
 	Reflecting the multicriteria nature of the problem, several measurements are considered for treatment evaluation.
@@
 
-\section{Dose-Function}
+\section{Dose-Function}\label{dose_function}
 The dose-distribution within the patient's body is the accumulation of all the radiation-seeds.

### 2015-03-29 Paper Seminar/isim_paper.tex
found one mistake

@@
 
-\subsection{Body modle}
+\subsection{Body model}
 The body model is supplied as a three-dimensional array ("body array"), containing the type of the body cells, within a matlab file (.mat).

### 2015-03-29 Paper Seminar/isim_paper.tex
Grammar and spelling correction, slight formulation beauty work.

@@
 \begin{algorithm}[H]
-maxDosis: \textit{double} [cGy] \\
-minDosis: \textit{double} [cGy] \\
-goalDosis: \textit{double} [cGy] \\
-currentDosis: \textit{double} [cGy] \\
+maxDosis: \textit{double} [Gy] \\
+minDosis: \textit{double} [Gy] \\
+goalDosis: \textit{double} [Gy] \\
+currentDosis: \textit{double} [Gy] \\
 coordinate: \textit{Coordinate} [x: units, y: units, z: units] \\
@@
 				\input{pictures/EvalCrit_DVH.tikz} 
-				\caption{A cumulative Dose Volume Histogram}
+				\caption{A cumulative dose volume histogram}
 				\label{fig:evalCritDVH}
@@
 The dose-distribution within the patient's body is the accumulation of all the radiation-seeds.
-Therefore it is mandatory to know the radiation intensity of each seed.
-Equation \eqref{eq:dosefunction} shows the general approach to approximate the dose intensity function and \eqref{eq:dosefunction2} shows the used implementation of that equation.
+Therefore, it is mandatory to know the radiation intensity of each seed.
+Equation \eqref{eq:dosefunction} shows the general approach to approximate the dose intensity function, whilst \eqref{eq:dosefunction2} is showing the used implementation of that equation.
 \begin{equation}
@@
 \\
-The following section will describe the used paramters:
+The following section will describe the used parameters:
 \begin{description}
 \item[Air-Kerma-Strength $S_{k}$]~\\
-The $S_{k}$ describes the \textbf{k}inetic \textbf{e}nergy \textbf{r}eleased per unit \textbf{ma}ss called kerma. It is measured in U. Where $1\ U =1 \ \frac{cGy \cdot m^{2}}{h} $   
+The $S_{k}$ describes the \textbf{k}inetic \textbf{e}nergy \textbf{r}eleased per unit \textbf{ma}ss called kerma. It is measured in $U$ ($1\ U =1 \ \frac{cGy \cdot m^{2}}{h} $)   
 \item[Dose Rate Constant $\Lambda$]~\\
@@
 \item[Dose Function $g_{L}(r)$]~\\
-The dose function describes the dose intensity dependent on the radial distance.
+The dose function describes the dose intensity depending on the radial distance.
 
 \item[Two dimensional anisotropy function $ F(r_{0},\theta_{0})$]~\\
-The anisotropy resolves issues due to the fact, that the dose distribution is'nt rotation invariant. Therefore the dose intensity is different for different angles.
+The anisotropy resolves issues due to the fact that the dose distribution isn't rotation invariant. Therefore, the dose intensity is different for different angles.
 
 \item[Geometry Function $G_{L}(r_{0},\theta_{0})$ ]~\\
-Due to scatter and attenuation an estimation of the dose distribution with point sources is less precise then an estimation with a line source.
-Therefore a line source model is used.
-\ref{fig:geometry} gives an overview of the geometry of the model.
+Due to scatter and attenuation, an estimation of the dose distribution with point sources is less precise then an estimation with a line source.
+Therefore, a line source model is used.
+Figure \ref{fig:geometry} gives an overview of the geometry of the model.
 
@@
 \end{figure}
-
 \end{description}
-\newpage
 
 Figure \ref{fig:doseplot} shows the plotted dose function.
-It is easily seen that the curve has a realy steep slope.
-Hence it should be discussed for which radii the given function should be evaluated.
+It is easily seen that the curve has a significant steep slope.
+Hence, it should be discussed, for which radii the given function should be evaluated.
 
@@
 
-Genetic algorithm  or evolutionary algorithm \cite{genpro} describes a search procedure which belongs to the category local search\cite{aima}.
+Genetic algorithm  or evolutionary algorithm \cite{genpro} describes a search procedure which belongs to the category of local searches\cite{aima}.
 Local search algorithms are used when the path to your solution is irrelavant (unlike e.g.
@@
 \begin{description}
-\item[Individual] ~\\
+\item[Individual]~\\
 \label{indiv}
@@
 It is forced to reproduce itself with other individuals so the solution might improve.
-The key here is to find an efficient encoding for your problem.
+The key is to find an efficient encoding for your problem.
 \item[Population]~\\
 \label{pop}
-The sum of all individuals is called population. The population provides the gene pool. Every individual shall reproduce it's gene with another one out of the population. 
-\item[Fitness-Function] ~\\
+The collection of all individuals is called population. The population provides the gene pool. Every individual shall reproduce it's gene with another one out of the population. 
+\item[Fitness-Function]~\\
 \label{fifu}
-The fitnees-function is a heuristic which is used to evaluate the fitness value of an individual.
-The fitness value indicates how good the solution is, which the individual contains.
-Further the fitness value is used to select the best individuals for reproduction.
+The fitness-function is a heuristic which is used to evaluate the fitness value of an individual.
+The fitness value indicates the quality of a solution contained by the individual.
+Further on, the fitness value is used to select the best individuals for reproduction.
 \end{description}
@@
 \item[1. Elitism]~\\
-In the elitism step a specific number of the best individuals are taken into the next iteration without further reproduction or other actions. 
+In the elitism step, a specific number of the best individuals are taken into the next iteration without further reproduction or other actions. 
 \item[2. Selection]~\\
-During the selection phase each individual gets selected with a probabilty, that is proportional to it's fitnessvalue.
-$P_{selection_{i}}= \tfrac{fitness_{i}}{\sum_{j}^I fitness_{j}}$ Where $I$ is the Population size and $i$ the specific index for a individual.
+During the selection phase, each individual gets selected with a probabilty which is proportional to its fitnessvalue.\\
+$P_{selection_{i}}= \tfrac{fitness_{i}}{\sum_{j}^I fitness_{j}}$, where $I$ is the population size and $i$ the specific index for an individual.
 \item[3. Crossover]~\\
-In the crossover step the previously selected individuals are taken pairwise to exchange their solution by using the one point crossover.
-One point crossover cuts the gene in two parts at a randomly chosen point and exchanges one half of each's gene.
+During the crossover step the previously selected individuals are taken pairwise to exchange their solution by using the one point crossover.
+One point crossover cuts the gene in two parts at a randomly chosen point and swaps one half of each gene.
 
 \item[4. Mutation]~\\
-In the end each individual changes randomly some parts of it's solution by a given probability.
+In the end, each individual changes randomly some parts of its solution by a given probability.
 \end{description}
@@
 
-The individual is ipmlemented as a $n^{th}$ dimensional vector of the dwell times, where $n$ denotes the number of seeds. Equation \eqref{eq:individual} shows one individual for $n  $ Seeds.
+The individual is implemented as a $n$-dimensional vector of dwell times, where $n$ denotes the number of seeds. Equation \eqref{eq:individual} shows one individual for $n $ seeds.
  \begin{equation}
@@
 
- The population is, like in \ref{pop} said, the accumulation of all individuals. $k$ denotes the number of individuals.
+ The population is the accumulation of all individuals (see \ref{pop}. $k$ denotes the number of individuals.
 \begin{equation}
@@
 
- The Fitness-Function calculates the squared differnce between the goaldose and the current dose within the whole body.
+ The fitness-function calculates the squared difference between the goal dose and the current dose within the whole body.
  \begin{equation}
@@
  \end{equation}
- For further adjustment and prioritisation the fitness-function is extended by an weighting factor, which holds a specifig weight for any bodytype (tumor, liver, spine,...).
+ For further adjustment and prioritisation the fitness-function is extended by a weight factor, which holds a specific weight for any body type (i.e. tumor, liver, spine).
  \begin{equation}
@@
 \label{subsubsec:optimisation}
-The fitness-function as described in equation \eqref{eq:fifu} is shown in algorithm \ref{alg:fifu}.
+The calculation of the fitness-function, as described in equation \eqref{eq:fifu}, is shown in algorithm \ref{alg:fifu}.
 It is easily seen that it iterates over every voxel in the body.
-This results in a very high computational complexity and high runtime.
-To improve the performance of the algorithm three different optimisation approaches are introduced.
+This results in a very high computational complexity and runtime.
+To improve the performance of the algorithm, three different optimisation approaches are introduced.
 \\
@@
 ~\\
-This strategy evaluates not every single voxel in the body. Only every $n^{th}$ voxel is evaluated. Algortithm \ref{alg:scalefif} shows how this is implemented. \\
+This strategy evaluates not every single voxel in the body. Only every n$th$ voxel is evaluated. Algorithm \ref{alg:scalefif} shows the implementation of scaling. \\
 
@@
 \label{para:treatment range}
-~\\
-Like the dosefunction chapter has shown, the distribution of the radiation decreases exponential with the distance.
-This effect could be used to reduce the number of voxel, that has to be evaluated, because the influence of voxels that are relatively far away is negligibly small.
+As it is shown in chapter \ref{dose_function}, the distribution of the radiation decreases exponentially with increasing distance.
+This effect could be used to reduce the number of voxel being evaluated, as the influence of voxels relatively far away is negligibly small.
 Algorithm \ref{alg:scalefif} shows how that information is used.
-The Region of Interest (roi) denotes the treatment range.
+The region of interest (roi) denotes the treatment range.
 \\
@@
   	}
-\caption{Optimised Fitness-Function}
+\caption{Optimised fitness-function}
 \end{algorithm}
@@
 ~\\ 
-The multithreading approach seperates the 
-three-dimensional
-body into two-dimensional planes. Each plane is then
-processed in an own thread. Algorithm \ref{alg:multithread} shows that approach.\\
+The multithreading approach seperates the three-dimensional body into two-dimensional planes.
+Each plane is then processed in an own thread.
+Algorithm \ref{alg:multithread} demonstrates this approach.\\
 
@@
 
-\subsubsection{Stop Criterion}
+\subsubsection{Stoping Criterion}
 \label{subsubsec: stopcrit}
 One of the challenges for the genetic algorithm is to decide when and how the algorithm should terminate.
-It is difficult to detmine how close the current iteration is to the optimal solution.
+It is difficult to determine how close the current iteration is to the optimal solution.
 The approach used in this implementation calculates the difference between the current fitness-value and the fitness-value of the previous iteration.
-If the absolute value of this differnece is less or equal to a small boundary value a counter is incremented.
+If the absolute value of this difference is less or equal to a small boundary value, a counter is incremented.
 If this counter is incremented five times in a row, the algorithm terminates.
 \\
-This procedure shall ensure that small fluctuations, that can accure due to the mutations within the individuals, will have no effect on the termination. \\
+This procedure shall ensure that small fluctuations, that can occur due to mutations within the individuals, will have no effect on the termination. \\
 
@@
 \subsubsection{Parameters}
-The implemented algorithm has a set of parameters, which is used to adjust it's outcome. The usage and the influence of those values are discussed in the following.
+The implemented algorithm has a set of parameters, which is used to adjust its outcome. The usage and the influence of those values are described and discussed in the following.
 
 \paragraph{Weights}~\\
-The fitness-function as described in \ref{fifu} uses individual weighting factors for each body type.  But every tissue type can be weighted with an individual coefficent.
+The fitness-function as described in \ref{fifu} uses individual weighting factors for each body type.  Nonetheless, every tissue type can be weighted with an individual coefficent.
 
 \paragraph{Probabilities} ~\\ 
-There are two important probabilites within the algorithm.
+There are two important probabilities within the algorithm.
 The first one is the crossover rate.
@@
 The fitness-function for the given optimisation problem has very high computational complexity.
-Hence two scaling parameters are introduced to achieve shorter runtimes.
+Hence, two scaling parameters are introduced to achieve shorter runtimes.
 The challenge here is to find a compromise between good accuracy for the optimisation and practicable runtimes.\\ The first parameter is the treatment range.
 It chooses the range around the PTV which shall be evaluated by the algorithm.
-The second parameter is a simple scaling value which defines that just every  $n^{th}$ voxel is evaluated.
-Within the treatment range of course.
+The second parameter is a simple scaling value which defines that just every  n$th$ voxel is evaluated within the treatment range.
 
@@
 The results have shown, that a high mutation rate leads to a very low tumor coverage.
-Furthermore the algorithm often doesn't converges, because the changes of the individuals accure too often.
+Furthermore, the algorithm often doesn't converge, because the changes of the individuals occur too often.
 A stable state becomes less likely.
@@
 \end{tabular}
-\caption{Mutation rate and coverage for different mutation-  \& crossover rates}
+\caption{Mutation rate and coverage for different mutation \& crossover rates}
 
@@
 The objective function, modeled with \textit{c},  is to be minimized, while the constraints, modeled with \textit{A} and \textit{b}, are satisfied.
-The used algorithm yields a vector x for the optimal solution.
+The used algorithm yields a vector $x$ for the optimal solution.
 
@@
 For our problem we used CPLEX, an implementation of the simplex algorithm by IBM[\cite{cplex}].
-In simplex the feasible region is considered a polyhedron.
+In simplex, the feasible region is considered a polyhedron.
 The optimal solution must be in one of the vertices of the polyhedron.
 Simplex finds a start vertex and approaches the optimal solution by moving from vertex to vertex on an edge, until there is no better solution.
-The constraints allow us a good modelling of the tumor, since we are able to define a constraint with certain limits for each body type.
+The constraints allow us a good modeling of the tumor, since we are able to define a constraint with certain limits for each body type.
 
@@
 This implementation checks every voxel of the body, if it is a tumor voxel or not.
-In case of a tumor voxel the influence of the radiation of every seed will be summed up, building a greater-equal constraint.
-For body voxels the influence of the radiation builds a term for the objective function.
+In case of a tumor voxel, the influence of the radiation of every seed will be accumulated, building a greater-equal constraint.
+For body voxels, the influence of the radiation builds a term for the objective function.
 
@@
 		\label{subsec:lpswIdea}
-		Based on an initial solution the planner can take a variable number of steps until an appropiate solution is determined.
+		Based on an initial solution, the planner can take a variable number of steps until an appropriate solution is determined.
 By relaxing bounds for one VOI, an improvement in another VOI might be achieved.
-Therefore, the method can be used to identiy tradeoffs for conflicting goals by specyfying dose bounds rather than weighting factors.
+Therefore, the method can be used to identify tradeoffs for conflicting goals by specifying dose bounds rather than weighting factors.
 Figure \ref{fig:lpswWorkflow} presents the basic workflow of this method.
@@
 While $\hat{t}$ and $\check{t}$ are valid for a whole VOI, $\hat{s}$ and $\check{s}$ correspond to a single voxel of a VOI.
-In an optimization step only one VOI is taken into account by setting the coefficient values to 1, specyfying the desired bounds and minimizing the objective function.
+In an optimization step, only one VOI is taken into account by setting the coefficient values to 1, specifying the desired bounds and minimizing the objective function.
 		\begin{equation}
@@
 	\subsubsection{Sampling}
-		To reduce computational complexity only a subset of voxels is taken into account for the linear program.
+		To reduce computational complexity, only a subset of voxels is taken into account for the linear program.
 A selection mechanism approximates the significance of voxels as a function of their distance to the closest seed.

### 2015-03-30 Paper Seminar/isim_paper.tex
Conformed usage of "optimization"

@@
 \newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1
+]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
 
@@
 
-\title{Treatment planning for HDR Brachytherapy}
-\date{March 31st, 2015}
-\author{S.~Elm \and
-T.~Karthe \and
-M.~E.~Schroeder \and
-D.~Golebiewski \and
-L.~Mordhorst
-}
-
-\maketitle
-\newpage
+\input{title.tex}
 
@@
 
-\subsubsection{Optimisation}
-\label{subsubsec:optimisation}
+\subsubsection{Optimization}
+\label{subsubsec:optimization}
 The calculation of the fitness-function, as described in equation \eqref{eq:fifu}, is shown in algorithm \ref{alg:fifu}.
@@
 This results in a very high computational complexity and runtime.
-To improve the performance of the algorithm, three different optimisation approaches are introduced.
+To improve the performance of the algorithm, three different optimization approaches are introduced.
 \\
@@
 \paragraph{Scaling and Accuracy}~\\
-The fitness-function for the given optimisation problem has very high computational complexity.
+The fitness-function for the given optimization problem has very high computational complexity.
 Hence, two scaling parameters are introduced to achieve shorter runtimes.
-The challenge here is to find a compromise between good accuracy for the optimisation and practicable runtimes.\\ The first parameter is the treatment range.
+The challenge here is to find a compromise between good accuracy for the optimization and practicable runtimes.\\ The first parameter is the treatment range.
 It chooses the range around the PTV which shall be evaluated by the algorithm.

### 2015-03-31 Paper Seminar/classification.tex
Fixed ref errors, did some beauty work (i.e. newpage after toc)
Rechtschreibung geprüft...

@@
 Afterwards, the body with the smallest distance to the treated body is taken.
-If it has an optimization result stored ("stored body), this will be used during the next iteration.
+If it has an optimization result stored ("stored body"), this will be used during the next iteration.
 This aims at lower runtimes for at least an equal coverage.
@@
 
-The four parameters descripted above are calculated in a measuring algorithm, detailed further in section \ref{classification:algorithm}.
+The four parameters described above are calculated in a measuring algorithm, detailed further in section \ref{classification:algorithm}.
 They contain two types of influence onto the norm: \textit{relative} and \textit{absolute}.
@@
 The norm itself is a sum of the different parameters multiplied with their weight, corresponding to the smallest error between the stored and treated bodies.
-The weights correspond to a simple consideration about how much each parameter should influence the the classification.
+The weights correspond to a simple consideration about how much each parameter should influence the classification.
 To allow for small to medium deviations within the three first parameters (volume size, volume center and distance tumor-center), they have been chosen small.
@@
 Additionally to the weights of the various weights of the measurement parameters, it's possible to change the influence of the different body types (compare table \ref{tbl:classification_type_weights}).
-This leads to a more detailed and precise representation of the aims during the treatment planning.
-In this case, a high focus has been set upon not damaging the spine and the pancreas whilst treating the tumor properly.
+This leads to a more detailed and precise representation of the goals during the treatment planning.
+In this case, a high focus has been set upon not damaging the spine and the pancreas, whilst treating the tumor properly.
 
@@
 To calculate the closest distance between the volumes, it's necessary cross-compare the voxels.
-Therefore, another loop iterates through the complete body, and, within, iterates over all previously stored tumor coordinates.
+Therefore, another loop iterates through the complete body and within iterates over all previously stored tumor coordinates.
 For each pair of voxel and tumor coordinate, the distance is calculated and, if less then before, the closest distance is updated (line 8-12).
 
+\begin{algorithm}
 \ForAll{$voxels\ in\ body$}{
@@
 \caption{Measurement algorithm in pseudo code}
+\label{alg:classification}
 \end{algorithm}
 
-\subsection{Performance evalution and prospects}
+\subsection{Performance evaluation and prospects}
 Without the calculation of the closest distance in between the tumor volume and all other volumes, the performance is linear.
-The computing takes only a few seconds (2-4s), using a body similar to dimensions of 110x110x200.
+The computation takes only a few seconds (2-4s), using a body similar to dimensions of 110x110x200.
 Alas, the computing time increases exponentially when performing the cross-comparison of each voxel in the body with all voxels of type \textit{tumor}.
@@
 Another step towards lower runtimes contains the implementation of a multi-threaded cross-comparison or completely multi-threaded algorithm.
-The multi-threaded cross-comparison has been implemented and it indicates a speed up by approximately the number of processor cores used, but has not beend thoroughly tested yet.
+The multi-threaded cross-comparison has been implemented and it indicates a speed up by approximately the number of processor cores used, but has not been thoroughly tested yet.
 \\

### 2015-03-31 Paper Seminar/isim_paper.tex
Rechtschreibung geprüft...

@@
 Additionally, conclusions about the quality of the resulting treatment plans are drawn.
-Lastly, a possible approach to reduce runtim complexity and reuse of prior knowledge is presented.
+Lastly, a possible approach to reduce runtime complexity and reuse of prior knowledge is presented.
 
@@
 The body array is set up as an equidistant point cloud of cells (1 unit $\widehat{=}$ 0.2cm), being transformed into 3D-array of voxels (compare algorithm \ref{model:voxel}).
-Besides the coordinate, a voxel contains several dosis variables (like extreme dosis, goal dosis) and the body type (normal tissue, spine, liver, pancreas and tumor tissue).
+Besides the coordinate, a voxel contains several dose variables (like extreme dose, goal dose) and the body type (normal tissue, spine, liver, pancreas and tumor tissue).
 \\
@@
 The VOIs might also be referred to as organs at risk (OAR).
-The ptv is in the middle of the figure (drawn red), surrounded by the liver (green).
+The PTV is in the middle of the figure (drawn red), surrounded by the liver (green).
 The yellow, cylindric volume behind the liver on the right represents the spine.
@@
 When eradiating into the surrounding tissue, the seeds influence overlap and form a superposition.
-Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limite, where the influence gets negligible (see chapter \ref{dose_function}).
+Hence, the dosis in a voxel is calculated as a sum of each seed by evaluating the dose-function up to a fixed limit, where the influence gets negligible (see chapter \ref{dose_function}).
 
@@
 \item[Dose Rate Constant $\Lambda$]~\\
-The dose rate constant is defined as the ratio of the dose rate at reference point $ P(r_{0},\theta_{0})  $ and the the air-kerma-strength $S_{k}$.
+The dose rate constant is defined as the ratio of the dose rate at reference point $ P(r_{0},\theta_{0})  $ and the air-kerma-strength $S_{k}$.
 Therefore $ \Lambda \ = \ \frac{\dot{D}(r_{0},\theta_{0})}{S_{k}} $.
@@
 
- The population is the accumulation of all individuals (see \ref{pop}. $k$ denotes the number of individuals.
+ The population is the accumulation of all individuals (see \ref{pop}) $k$ denotes the number of individuals.
 \begin{equation}
@@
 Algorithm \ref{alg:scalefif} shows how that information is used.
-The region of interest (roi) denotes the treatment range.
+The region of interest (ROI) denotes the treatment range.
 \\
@@
 \paragraph{Weights}
-It seems hard to examine a specific sets of weights, that is applicable to any type of patients.
+It seems hard to examine a specific sets of weights, that is applicable to any type of patient.
 The experiment has shown, that a higher weighted treatment has a much higher runtime and the convergency decreases.
@@
 \section{Linear Programming}
-In Linear Programming an objective function and constraints are used to compute the optimal solution to a minimization or maximization problem, as described in [\cite{matprog}].
+In linear programming an objective function and constraints are used to compute the optimal solution to a minimization or maximization problem, as described in [\cite{matprog}].
 The problem consists of a certain number of variables and a feasible region, which is defined by the contraints.

### 2015-04-08 Paper Seminar/isim_paper.tex
Added SA to Paper Seminar

@@
 The dose distribution of a radioactive seed shows that the dose at a radius of 10 cm is negligible small. So any treatment range higher than 10 cm wouldn't lead to better results. 
+\section{Simulated Annealing}
+\label{sec:SAIntro}
+
+Simulated Annealing \cite{genpro} describes a Monte Carlo (deterministic runtime) approach to minimization of multivariate functions.
+The idea of Simulated Annealing has been first proposed in a paper published by Metropolis et al.
+in 1953.
+The algorithm in said paper simulated the cooling of solids in a heat bath.
+This cooling process is known as annealing.
+
+The idea of the Simulated Annealing algorithm is to find the global minimum of a cost function by randomly selecting a valid solution from the set of viable solutions and determining its cost and keeping solutions which are more optimal than the previous iteration's solution.
+Following the annealing analogy, a less optimal solution can be accepted following a certain transition probability.
+This serves to prevent falling into local minima.
+\\
+The Simulated Annealing algorithm is comprised of
+
+\begin{description}
+\item[State]~\\
+\label{state}
+A state is any viable solution from the set of viable solutions for a problem. It can take be of any type.
+\item[Cost Function]~\\
+\label{cofu}
+The cost function is at the center of the Simulated Annealing algorithm.
+It determines the quality of a computed state with respect to certain criteria which embody the cost function.
+Since the cost is what is minmized (or maximized) during the iterative process, the cost function is the function whose global minimum is sought by Simulated Annealing.
+\\
+that maps from the solution space to $\mathbb{R}$ 
+\item[Metropolis Function, Metropolis Loop]~\\
+\label{metro}
+The \emph{Metropolis Function} is the function that determines how states of higher cost than the previous iteration's candidate solution state are handled.
+The \emph{Metropolis Loop} is the main loop which determines the cost of a candidate solution, compares it to the current best solution, accepts or discards it according to the Metropolis Function, computes a new candidate state and adjusts the annealing schedule.
+\item[Annealing Schedule]~\\
+\label{anneal}
+As Simulated Annealing is modeled on the annealing of solids, the annealing schedule emulates the influence of temperature on the mobility of molecular structures in solids.
+The further the virtual temperature drops, the less likely a less optimal state is accepted as the current best.
+The speed of the temperature decrease is called the \emph{annealing schedule}.
+\end{description}
+
+Simulated Annealing is an iterative algorithm. Each iteration consists of four steps, which are described in the following:
+
+\begin{description}
+\item[1. Next candidate state generation step]~\\
+This step consists of one operation: a random generation of a state vector or any other data structure to hold the values determining a solution to the optimization problem.
+\item[2. Cost calculation step]~\\
+This step calculates the cost of the randomly selected solution and stores it as a numerical value for efficent processing. The cost function is called with the candidate state as the argument.
+\item[3. Decision step]~\\
+The decision step determines if the currently best solution should be overwritten.
+If the cost of the candidate solution is better than the current best, the candidate solution becomes the current best and the candidate cost becomes the current best as well.
+\\ If the the candidate cost is lower than the current best, a random number between 0 and 1 (excluding those) is drawn and if that number lies below the value of the Metropolis Function then the current best cost and state are updated regardless.
+\\ In any other case, the candidate solution is rejected.
+\item[4. Annealing schedule adaptation step]~\\
+In this last step of an iteration, the virtual temperature is updated to a new value according to a predefined annealing schedule that determines the rate at wich the annealing temperature is meant to fall.
+This temperature is used in the next iteration to determine the value of the Metropolis Function.
+\end{description}
+
+\subsection{Implementation}
+\label{subsec:impl}
+The state is defined as the $n$-dimensional vector of all current dwelltimes, with $n$ denoting the number of seeds. Equation \eqref{eq:SAstate} shows a state vector for $n$ seeds.
+ \begin{equation}
+ \label{eq:SAstate}
+ State \ \ s := \begin{pmatrix}
+ t_{1} \\ \vdots \\ t_{n} 	
+\end{pmatrix}   
+\end{equation}
+
+The cost-function calculates the sum of squared differences between the goal dose and the current dose within the whole body, since the incurred radiation dose needs to be determined to evaluate the 
+ \begin{equation}
+ f(state) = \sum_{body}(GD-CD)^2
+ \end{equation}
+ To allow for tuning of the algorithm, the cost-function is extended by a weight factor for each subclass of voxel (i.e. tumor, liver, spine).
+ \begin{equation}
+ \label{eq:cofu}
+ f(state) = \sum_{body}w_{bodytype}(GD-CD)^2
+ \end{equation}
+
+\subsubsection{Optimization}
+\label{subsubsec:SAoptimization}
+The implementation of the cost-function, as described in equation \eqref{eq:cofu}, is shown in algorithm \ref{alg:cofu}.
+It iterates over every voxel in the body.
+This is computationally inefficient.
+To improve the performance of the cost calculation, an extension to the dose calculation process is introduced.
+\\
+
+\begin{algorithm}[H]
+\label{alg:cofu}
+ $diffcost=0$\;
+ $tempintensity=0$\;
+ \For{x=0; x $\textless$ body.xDim; x++}{
+ 	\For{y=0; y $\textless$ body.yDim; y++}{
+ 		\For{z=0; z $\textless$ body.zDim; z++}{
+ 			\For{j=0; j $\textless$ numberOfSeeds; j++} {
+				 	$tempintensity\ \ = tempintensity +  dosebyseed(x,y,z,j) $\;		
+ 			}
+  		}
+  		$diffcost += {w_{bodytype}\cdot(GoalDose-CurrentDose)^2}$\;
+  		% $diffcost = \sum_{body}{w_{bodytype}\cdot(GoalDose-CurrentDose)^2}$\;
+  	}
+  }
+
+ \Return{$\sqrt{diffcost}$}\;
+ \caption{Calculation of the cost}
+
+ \end{algorithm}
+
+ ~\\ The following extension is based on the fact that the dosefunction, save for the distance( and the angles in future implementations) , remains constant.
+This is leveraged in precomputing the value of the dosefunction and saving it in memory.
+The cost-function algorithm changes as follows: \\
+
+\begin{algorithm}[H]
+\label{alg:cofu2}
+ $diffcost=0$\;
+ $tempintensity=0$\;
+ \For{x=0; x $\textless$ body.xDim; x++}{
+ 	\For{y=0; y $\textless$ body.yDim; y++}{
+ 		\For{z=0; z $\textless$ body.zDim; z++}{
+ 			\For{j=0; j $\textless$ numberOfSeeds; j++} {
+				 	$tempintensity +=  efficientdosebyseed(x,y,z,j) $\;		
+ 			}
+  		}
+  		$diffcost += {w_{bodytype}\cdot(GoalDose-CurrentDose)^2}$\;
+  		% $diffcost = \sum_{body}{w_{bodytype}\cdot(GoalDose-CurrentDose)^2}$\;
+  	}
+  }
+
+ \Return{$\sqrt{diffcost}$}\;
+ \caption{Calculation of the cost by an efficient dosefunction}
+
+ \end{algorithm}
+
+\subsubsection{Stopping Criterion}
+\label{subsubsec: SAstopcrit}	
+An integral Part of the Simulated Annealing algorithm is its known stopping condition: the algorithm stops once the "temperature" has reached zero.
+This is determined by the so-called "annealing schedule", which determines the drop in temperature.
+Since it is a function with the current iteration as its only input, it can be fashioned in any way desired.
+\\
+
+\begin{algorithm}[H]
+	$T_{i+1} = T_{i} \cdot \frac{1}{i}$~\\
+	$...$ ~\\
+	if $T<=0$ do break;
+
+\caption{Stopping Criterion}
+\end{algorithm}
+
+\newpage
+\subsubsection{Parameters}
+Simulated Annealing, as introduced, has a set of parameters which can be tuned to improve performance. The relevant parameters and their effects are discussed here.
+
+\paragraph{Weights}~\\
+The cost-function as described in \ref{cofu} uses individual weighting factors for each body type.
+The estimation of these weights is the at the core of influencing the therapy outcome since it directly penalizes radiation accumulation for specific areas of the irradiated volume.
+It can be taken even further by using gradients instead of fixed weights to penalize the core structures of a certain organ more than the outside layers.
+This of course requires verification with a medical professional.
+
+\paragraph{Transition Probability} ~\\ 
+The transition probability of the Metropolis Function \ref{metro} has been deduced from how a thermodynamic system, minimizing its temperature, moves from its current state to a following state.
+The optimality of this transition depends on the shape of the cost-function \ref{cofu} and thus depends on each patient's case and cannot be generalized for a truly optimal outcome.
+Since its main use is to prevent the falling into local minima, multiple runs combined with evolutionary algorithms might be an approach worth investigating.
+
+\paragraph{Annealing Schedule} ~\\ 
+The annealing schedule determines how quickly the likelihood of accepting a less optimal state drops.
+In the current implementation so-called Cauchy annealing is used.
+Tuning this parameter influences both computation speed and risk of local minimization.
+A truly adaptive annealing schedule that takes all previous cost-function computations into account would be a very good starting point.
+
+\subsubsection{Results}
+
+\paragraph{Weights in cost-function}
+As mentioned before, the weights are preset for any computation.
+If the outcome of a computation is not satisfactory, a modification of the weights and recomputation is necessary.
+In the tests run, weighting by tissue type has had a influence on runtime since the difference need to be multiplied by a scalar only but, as mentioned, recomputation is necessary for improving the outcome.
+\\
+
+%\newpage
+\paragraph{Precomputed Dose-function} 
+The use of a precomputed dose-function has reduced computation time by 70.8 \%  on average while keeping the results consistent.
 

### 2015-04-08 Paper Seminar/isim_paper.tex
korrektur

@@
 \label{state}
-A state is any viable solution from the set of viable solutions for a problem. It can take be of any type.
+A state is any viable solution from the set of viable solutions for a problem. It can be of any type.
 \item[Cost Function]~\\
@@
 It determines the quality of a computed state with respect to certain criteria which embody the cost function.
-Since the cost is what is minmized (or maximized) during the iterative process, the cost function is the function whose global minimum is sought by Simulated Annealing.
+Since the cost is what is minimized (or maximized) during the iterative process, the cost function is the function whose global minimum is sought by Simulated Annealing.
 \\
-that maps from the solution space to $\mathbb{R}$ 
 \item[Metropolis Function, Metropolis Loop]~\\
@@
 \item[2. Cost calculation step]~\\
-This step calculates the cost of the randomly selected solution and stores it as a numerical value for efficent processing. The cost function is called with the candidate state as the argument.
+This step calculates the cost of the randomly selected solution and stores it as a numerical value for efficient processing. The cost function is called with the candidate state as the argument.
 \item[3. Decision step]~\\
@@
 If the cost of the candidate solution is better than the current best, the candidate solution becomes the current best and the candidate cost becomes the current best as well.
-\\ If the the candidate cost is lower than the current best, a random number between 0 and 1 (excluding those) is drawn and if that number lies below the value of the Metropolis Function then the current best cost and state are updated regardless.
+\\ If the candidate cost is lower than the current best, a random number between 0 and 1 (excluding those) is drawn and if that number lies below the value of the Metropolis Function then the current best cost and state are updated regardless.
 \\ In any other case, the candidate solution is rejected.
 \item[4. Annealing schedule adaptation step]~\\
-In this last step of an iteration, the virtual temperature is updated to a new value according to a predefined annealing schedule that determines the rate at wich the annealing temperature is meant to fall.
+In this last step of an iteration, the virtual temperature is updated to a new value according to a predefined annealing schedule that determines the rate at which the annealing temperature is meant to fall.
 This temperature is used in the next iteration to determine the value of the Metropolis Function.
@@
 \label{subsec:impl}
-The state is defined as the $n$-dimensional vector of all current dwelltimes, with $n$ denoting the number of seeds. Equation \eqref{eq:SAstate} shows a state vector for $n$ seeds.
+The state is defined as the $n$-dimentional vector of all current dwelltimes, with $n$ denoting the number of seeds. Equation \eqref{eq:SAstate} shows a state vector for $n$ seeds.
  \begin{equation}
@@
 
-The cost-function calculates the sum of squared differences between the goal dose and the current dose within the whole body, since the incurred radiation dose needs to be determined to evaluate the 
+The cost-function calculates the sum of squared differences between the goal dose and the current dose within the whole body, since the incurred radiation dose needs to be determined to evaluate the cost function.
  \begin{equation}
@@
 \label{subsubsec: SAstopcrit}	
-An integral Part of the Simulated Annealing algorithm is its known stopping condition: the algorithm stops once the "temperature" has reached zero.
+An integral part of the Simulated Annealing algorithm is its known stopping condition: the algorithm stops once the "temperature" has reached zero.
 This is determined by the so-called "annealing schedule", which determines the drop in temperature.
